### 2022-03-09 01_Species_abundance.Rmd
Initial commit

---
title: "01_Vegetation_cover"
author: "Annegreet Veeken"
date: "24-11-2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# load libraries
library(tidyverse)
library(readxl)
library(Taxonstand)
```

## To do

## QGIS

Import Scotland land cover HLCM2020_SCOTLAND_TIFF

-   Set layer style with .qml Import field site coordinate

-   Reproject to british national grid

Split vector layer based on Zone

-   Zone_A\_coordinates.shp = zone A coordinates

-   Zone_B\_coordinate.shp = zone B veg plot coordinates

Create buffer zone of 100 m and 1000 m

-   Zone_B\_buffer.shp

-   Zone_C\_buffer.shp

Use Zone histogram to compute pixels in zone attributed to certain land
cover type Export as Veg_cover_zoneC.csv and Veg_cover_zoneB.csv

Use Sample raster values to compute landcover types for vegetation
plots. Export as Veg_type_zoneB_sites.csv and Veg_type_zoneA_sites.csv

## Prepare data
Vegetation cover data.
```{r load and prepare veg cover data}
# load data
zone_b <- read.csv("Data/Veg_cover_zoneB.csv")
zone_c <- read.csv("Data/Veg_cover_zoneC.csv")

# EUNIS land cover classifications
veg_type_labels <- c(HISTO_1 = "C1: Water",
                     HISTO_2 = "D1: Raised and blanket bogs",
                     HISTO_3 = "D2: Valley mires, poor fens and transition mires",
                     HISTO_4 = "D4: Base-rich fens and calcareous spring mires", 
                     HISTO_5 = "E1: Dry grasslands",
                     HISTO_6 = "E2: Mesic grasslands",
                     HISTO_7 = "E3: Seasonally wet and wet grasslands",
                     HISTO_8 = "E4: Alpine and subalpine grasslands",
                     HISTO_9 = "E5: Woodland fringes and clearings and tall forb stands",
                     HISTO_10 = "F2: Arctic, alpine and subalpine scrub",
                     HISTO_11 = "F3: Temperate and mediterranean-montane scrub",
                     HISTO_12 = "F4: Temperate shrub heathland",
                     HISTO_13 = "F9: Riverine and fen scrubs",
                     HISTO_14 = "G1: Broadleaved deciduous woodland",
                     HISTO_15 = "G3: Coniferous woodland",
                     HISTO_16 = "G4: Mixed deciduous and coniferous woodland",
                     HISTO_17 = "G5: Line of trees, amall anthropogenic woodlands, recently felled woodland, early-stage woodland and coppice",
                     HISTO_18 = "H2: Screes",
                     HISTO_19 = "Inland cliffs, rock pavements and outcrops",
                     HISTO_20 = "I1: Arable land and market gardens",
                     HISTO_21 = "J: Builtup (and Bare before post)",
                     HISTO_23 = "O: Bare (after post)")

# make data tidy
zone_b <- zone_b %>% 
  select(sitename = name, lat, lon, contains("HISTO")) %>% 
  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
  # rename veg_types
  mutate(veg_type = recode(veg_type, !!!veg_type_labels),
  # rename sitenames       
         sitename = str_remove(sitename, "-A")) %>% 
  # filter land cover type that are not vegation
  filter(!veg_type %in% c("C1: Water", 
                         "J: Builtup (and Bare before post)",
                         "O: Bare (after post)")) %>% 
  # calculate veg_type percentage
  group_by(sitename) %>% 
  # calculate percentages
  mutate(veg_cov_b = value/sum(value)) %>%
  group_by(sitename, veg_type) %>%
  summarise(veg_cov_b = sum(veg_cov_b, na.rm = TRUE))

# make data tidy
zone_c <- zone_c %>% 
  select(sitename = name, lat, lon, contains("HISTO")) %>% 
  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
  # rename veg_types
  mutate(veg_type = recode(veg_type, !!!veg_type_labels),
  # rename sitenames       
         sitename = str_remove(sitename, "-A")) %>% 
  # filter land cover type that are not vegation
  filter(!veg_type %in% c("C1: Water", 
                         "J: Builtup (and Bare before post)",
                         "O: Bare (after post)")) %>% 
  # calculate veg_type percentage
  group_by(sitename) %>% 
  # calculate percentages
  mutate(veg_cov_c = value/sum(value)) %>%
  group_by(sitename, veg_type) %>%
  summarise(veg_cov_c = sum(veg_cov_c, na.rm = TRUE))

veg_type <- left_join(zone_b, zone_c, by = c("sitename", "veg_type")) 
```

```{r load and prepare vegetation data}
# Load in vegetation type data for the plots
zone_b_site <- read.csv("Data/20220106-Veg_type_zoneB_sites.csv")

# Make tidy
zone_b_site <- zone_b_site %>% 
  # create seperate site and plot column
  separate(name, c("sitename", "plot")) %>% 
  # recode veg_type
  mutate(veg_type = as.factor(rvalue_1)) %>% 
  select(sitename, plot, lat, lon, veg_type) %>% 
  mutate(veg_type = recode(veg_type, 
                           "1" = "C1: Water",
                           "2" = "D1: Raised and blanket bogs",
                           "3" = "D2: Valley mires, poor fens and transition mires",
                           "4" = "D4: Base-rich fens and calcareous spring mires", 
                           "5" = "E1: Dry grasslands",
                           "6" = "E2: Mesic grasslands",
                           "7" = "E3: Seasonally wet and wet grasslands",
                           "8" = "E4: Alpine and subalpine grasslands",
                           "9" = "E5: Woodland fringes and clearings and tall forb stands",
                           "10"= "F2: Arctic, alpine and subalpine scrub",
                           "11" = "F3: Temperate and mediterranean-montane scrub",
                           "12"= "F4: Temperate shrub heathland",
                           "13" = "F9: Riverine and fen scrubs",
                           "14"= "G1: Broadleaved deciduous woodland",
                           "15" = "G3: Coniferous woodland",
                           "16" = "G4: Mixed deciduous and coniferous woodland",
                           "17" = "G5: Line of trees, amall anthropogenic woodlands, recently felled woodland, early-stage woodland and coppice",
                           "18" = "H2: Screes",
                           "19" = "Inland cliffs, rock pavements and outcrops",
                           "20" = "I1: Arable land and market gardens",
                           "21" = "J: Builtup (and Bare before post)",
                           "23" = "O: Bare (after post)"))

# load in vegetation data
spec_b_raw <- read_xlsx("Data/species_zone_B.xlsx")

# make tidy
spec_b <- spec_b_raw %>% 
  pivot_longer(contains("plot"), names_to = "plot", values_to = "abun",
               values_drop_na = TRUE) %>% 
  mutate(plot = str_remove(plot, "plot_ ")) %>% 
  select(sitename = Site, plot, species = Species, abun) %>% 
  # join with vegetation type data
  left_join(zone_b_site, by = c("sitename", "plot"))
```

## Zone B & C

```{r calculate vegetation cover}
veg_cover_bc <- spec_b %>% 
  # remove moss
  filter(!species %in% c("Moss", "Sphagnum", "Polytrichum commune",
                         "Sphagnum squarrosum")) %>% 
  # calculate avarage species cover per vegetation type using the plots belonging to that vegetation type
  group_by(sitename, veg_type, species) %>% 
  summarise(abun = mean(abun)) %>% 
  # join with vegetation cover data
  full_join(veg_type, by = c("sitename", "veg_type")) %>% 
  # weight species abundance by veg cover
  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c) %>% 
  # drop unneeded columns
  select(sitename, veg_type, species, veg_cov_b, veg_cov_c, spec_abun_b,
         spec_abun_c)

# print table with missing species data for vegetation types
missing_vegtype <- veg_cover_bc %>% 
  filter(is.na(species)) %>% 
  group_by(sitename, veg_type) %>% 
  summarise(cov_b = round(sum(veg_cov_b), 2), cov_c = round(sum(veg_cov_c), 2)) %>%
  filter(!cov_b == 0|!cov_c == 0)

# for missing data, use species data from other plots
missing <- spec_b %>% 
  # remove moss
  filter(!species %in% c("Moss", "Sphagnum", "Polytrichum commune",
                         "Sphagnum squarrosum")) %>% 
  filter(veg_type %in% missing_vegtype$veg_type) %>% 
  group_by(veg_type, species) %>% 
  summarise(abun = mean(abun))

veg_cover_bc_na <- veg_cover_bc %>% 
  filter(is.na(species)) %>%
  select(-species) %>% 
  left_join(missing, by = "veg_type") %>% 
  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c) %>% 
  select(sitename, veg_type, species, veg_cov_b, veg_cov_c, spec_abun_b,
         spec_abun_c)

veg_cover_bc <-  veg_cover_bc %>% 
  filter(!is.na(species)) %>% 
  bind_rows(., veg_cover_bc_na) %>% 
  arrange(sitename, veg_type, species)
```

## Zone A

```{r}
# load data
spec_a_raw <- read_xlsx("Data/Species_zone_A.xlsx", skip = 1)

# Make tidy
spec_a <- spec_a_raw %>% 
  pivot_longer(3:19, names_to = "plot", values_to = "abun") %>% 
  rename(sitename = Site, species = Species) %>% 
  # remove moss
  filter(!species %in% c("Moss", "Sphagnum", "Polytrichum commune",
                         "Sphagnum squarrosum")) %>% 
  # create direction and distance column
  mutate(direction = case_when(str_detect(plot, "N") ~ "North",
                               str_detect(plot, "S") ~ "South",
                               str_detect(plot, "E") ~ "East",
                               str_detect(plot, "W") ~ "West",
                               str_detect(plot, "NE") ~ "Northeast",
                               str_detect(plot, "NW") ~ "Northwest",
                               str_detect(plot, "SW") ~ "Southwest",
                               str_detect(plot, "SE") ~ "Southeast",
                               plot == "Center" ~ "Center"),
         distance = case_when(plot %in% c("N...4", "E...5", "S...6","W...7") ~ 
                                "1.5-3 meter",
                              plot %in% c("N...8", "E...9", "S...10", "W...11")  ~ 
                                "3-6 meter",
                              plot %in% c("N...12","NE","SE","NW","SW", "E...14", "S...16", "W...18") ~ 
                                "6-10 meter",
                              plot == "Center" ~ "0 meter")) %>% 
  # create unique id for plot in site
  group_by(sitename, direction, distance) %>% 
  mutate(plot = cur_group_id())

```

## Standardization species names
```{r standardize species names}
harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx")

spec_a <- spec_a %>% 
  left_join(harm, by = "species")
veg_cover_bc <- veg_cover_bc %>% 
  left_join(harm, by = "species")

saveRDS(veg_cover_bc, "RDS_files/01_Species_abundance_bc.rds")
saveRDS(spec_a, "RDS_files/01_Species_abundance_a.rds")
```

### 2022-03-09 03_CWM_estimates.Rmd
Initial commit

---
title: "Estimating CWM"
author: "Annegreet Veeken"

output: html_document
---

This script estimates CWM values from the pollen data and from the species measurements

```{r load packages}
library(tidyverse)
library(rjags) 
library(runjags)
library(mcmc)
library(coda)
# library(xlsx)
library(Hmisc) # capitalize function
library(furrr) 
```
# Questions
- use all trait data or per site
- use all flowering/not-flowering

# To do 
- run Plantheight for pollen
- use beta distribution

# Pollen based CWM

## Load data

```{r load gapfilled data, echo=FALSE}
## Species per pollen type list
spec <- readRDS("RDS_files/02_PollenType_species.rds") %>% as_tibble()
spec <- spec %>% 
  distinct(stand.spec, pollentaxon) # remove alternative names

# check for duplicate values (species assigned to two or more pollen type) 
duplicats <- spec %>% 
  filter(duplicated(stand.spec)|
           duplicated(stand.spec, fromLast = TRUE))

## Gap-filled trait data
gapdata <- readRDS("RDS_files/Gap_filled_traits_2020.rds") 
# fix capitalization in  some species names in trait data
gapdata <- gapdata %>% 
  mutate(Species = tolower(Species) %>% capitalize(.))
# fix some species names
spec$stand.spec[spec$stand.spec == "Salix salamonii"] <- 
  "Salix x salamonii"  
spec$stand.spec[spec$stand.spec == "Rabelera holostea"] <- 
  "Stellaria holostea"  
spec$stand.spec[spec$stand.spec == "Doronicum excelsum"] <- 
  "Doronicum plantagineum"

# check for missing species in the trait data
misspec <- 
  spec$stand.spec[!spec$stand.spec %in% gapdata$Species]
misspec <- tibble(miss.spec = misspec, 
                  note = c("Hybrid", "Not available, but close relative S.fragilus is",
                           "not a species", "Not available", "Not available, garden plant", 
                           "Not available, garden plant", "Not available, garden plant",
                           "Hybrid", "Hybrid","Not available", "Not available, garden plant",
                           "Not available", "Not available","Not available", "Not available",
                           "Not available, garden plant", "Hybrid", "Hybrid","Not available",
                           "Not available"))
# only the relevant species  
gapdata <- gapdata %>% 
  filter(Species %in% spec$stand.spec)

# merge trait data with spec/pol list
dfTRAIT <- gapdata %>% 
  inner_join(spec, by = c("Species" = "stand.spec")) %>% 
  arrange(pollentaxon) # sort alphabetically

# Pollen data
dfPOL <- readRDS("RDS_files/01_pollen_data_Scot.rds")
pollentaxa <- dfPOL %>% 
  pull(pollentaxon) %>% 
  unique()

traits <- c("PlantHeight", "SLA", "LA", "LeafC", "LeafN", "LDMC")
```

## Normal traits
```{r cmw model log}
dfPOL <- dfPOL %>% 
  ungroup() %>% 
  select(sitename, pollentaxon, percent) %>% 
  # filter out unindentified
  filter(!pollentaxon == "unindentified") %>% 
  pivot_wider(names_from = pollentaxon, values_from = percent)

cwmlogfunc <- function(selectedtrait){
Trait <- dfTRAIT %>%
  filter(pollentaxon %in% pollentaxa) %>%
  pull(selectedtrait)

Tax <- dfTRAIT %>%
  filter(pollentaxon %in% pollentaxa) %>%
  pull(pollentaxon) %>% 
  as.factor()

N <- length(Trait)

Ab <- dfPOL %>%
  dplyr::select(-sitename) %>%
  as.matrix()
Ntax <- ncol(Ab)
Nsites <- nrow(Ab)

# for setting priors taxon level
Mean <- mean(Trait)
SD <- sd(Trait)

cat("model {
for (i in 1:Nsites){
cwm[i] ~ dnorm(mean.tax[r[i]], tau.tax[r[i]])
r[i] ~ dcat(Ab[i, 1:Ntax])
}

for (i in 1:N) {
      Trait[i] ~ dlnorm(mean.tax[Tax[i]], tau.tax[Tax[i]])
}

for (l in 1:Ntax){
# vague priors taxon level
    mean.tax[l] ~ dnorm(0, 10^-4)
    tau.tax[l] ~ dgamma(0.001, 0.001)
    sd.tax[l] <- sqrt(1/tau.tax[l])
}

#data# N, Nsites, Trait, Tax, Ntax, Ab
#monitor# cwm
#inits# mean.tax, tau.tax
}", file = "CWMmodel_mix.txt")

# initial values
mean.tax <- list(chain1 = rep(min(Trait),Ntax), chain2 = rep(max(Trait),Ntax))
tau.tax <- list(chain1 = rep(1/(SD*0.1)^2,Ntax), chain2 = rep(1/(SD*10)^2, Ntax))

results <- run.jags("CWMmodel_mix.txt", n.chains = 2)

plot(results, file = paste0("Convergence_diagnostics/Conv_pollen_", selectedtrait,".rds"))

res <- as.matrix(summary(results))
saveRDS(res, 
        paste0("RDS_files/03_CWM_estimates_pollen_", selectedtrait,
               ".rds"))
}

normaltraits <- c("SLA","LA","LeafC","LeafN")
plan(multisession, workers = 2)
furrr::future_map(normaltraits, ~cwmlogfunc(selectedtrait = .))
```

## Beta traits

```{r ldmc all taxa}
Trait <- dfTRAIT %>%
  filter(pollentaxon %in% pollentaxa) %>%
  pull(LDMC)

Tax <- dfTRAIT %>%
  filter(pollentaxon %in% pollentaxa) %>%
  pull(pollentaxon) %>% 
  as.factor()

N <- length(Trait)

Ab <- dfPOL %>%
  dplyr::select(-sitename) %>%
  as.matrix()
Ntax <- ncol(Ab)
Nsites <- nrow(Ab)

# for setting priors taxon level
Mean <- mean(Trait)
SD <- sd(Trait)

cat("model {
for (i in 1:Nyr){
cwm[i] ~ dbeta(alpha.tax[r[i]], beta.tax[r[i]])
r[i] ~ dcat(Ab[i, 1:Ntax])
}

for (i in 1:N) {
      Trait[i] ~ dbeta(alpha.tax[Tax[i]], beta.tax[Tax[i]])
}
for (l in 1:Ntax){
# vague priors taxon level
    alpha.tax[l] <- mean.tax[l] * sd.tax[l] 
    beta.tax[l] <- (1 - mean.tax[l]) * sd.tax[l]
    mean.tax[l] ~ dnorm(0, 10^-4) 
    sd.tax[l]  ~ dgamma(0.001,0.001)
}

#data# N, Nyr, Trait, Tax, Ntax, Ab
#monitor# cwm
#inits# mean.tax, sd.tax
}", file = "CWMmodel_beta_mix.txt")

# initial values
mean.tax <- list(chain1 = rep(min(Trait),Ntax), chain2 = rep(max(Trait),Ntax))
sd.tax <- list(chain1 = rep(1/(SD*0.1)^2,Ntax), chain2 = rep(1/(SD*10)^2, Ntax))

results <- run.jags("CWMmodel_beta_mix.txt", n.chains = 2)

plot(results, vars = params, file = "Convergence_diagnostics/Conv_pollen_LDMC.pdf")

res <- as.matrix(summary(results))

saveRDS(res, "RDS_files/03_CWM_pollen_LDMC.rds")
```

```{r plant height all taxa}
# Plant height (not in gapfilled data)
TRYdata <- read.csv("Data/24082020TRY_long_cleaned.csv")
TRYdata <- TRYdata %>% filter(AccSpeciesName %in% spec$stand.spec)

TRYdata <- TRYdata %>%
  # convert cm to m
  mutate(StdValue.m = ifelse(UnitName == "cm", StdValue/100, StdValue)) %>% 
  filter(CleanTraitName == "plant_height_generative") %>% 
  # join with pollen list
  left_join(spec, by = c("AccSpeciesName" = "stand.spec"))

Trait <- TRYdata %>%
  filter(pollentaxon %in% pollentaxa) %>%
  pull(StdValue.m)

Tax <- TRYdata %>%
  filter(pollentaxon %in% pollentaxa) %>%
  filter(CleanTraitName == "plant_height_generative") %>%
  pull(pollentaxon) %>% 
  as.factor()

N <- length(Trait)

Ab <- dfPOL %>%
  dplyr::select(-sitename) %>%
  as.matrix()
Ntax <- ncol(Ab)
Nsites <- nrow(Ab)

# for setting priors taxon level
Mean <- mean(Trait)
SD <- sd(Trait)

# for setting priors taxon level
Mean <- mean(Trait)
SD <- sd(Trait)

# initial values
mean.tax <- list(chain1 = rep(min(Trait),Ntax), chain2 = rep(max(Trait),Ntax))
tau.tax <- list(chain1 = rep(1/(SD*0.1)^2,Ntax), chain2 = rep(1/(SD*10)^2, Ntax))

results <- run.jags("CWMmodel_mix.txt", n.chains = 2)

plot(results, vars = params, file = 
       "Convergence_diag_unbinned_PlantHeight.pdf")

res <- as.matrix(summary(results))

saveRDS(res, "RDS_files/03_CWM_pollen_PlantHeight.rds")
```

# Species based CWM

```{r load data}
# Trait data
plantheight <- readRDS("RDS_files/01_PlantHeight.rds") %>% 
  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
  ungroup()
ldmc <- readRDS("RDS_files/01_LDMC.rds") %>% 
  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
  ungroup()
sla <- readRDS("RDS_files/01_SLA.rds") %>% 
  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
  ungroup()
la <- readRDS("RDS_files/01_LA.rds") %>% 
  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
  ungroup()

lTrait <- list(PlantHeight = plantheight,
               LDMC = ldmc,
               SLA = sla,
               LA = la)

# Species data
dfABUN_a <- readRDS("RDS_files/01_Species_abundance_a.rds")
dfABUN_bc <- readRDS("RDS_files/01_Species_abundance_bc.rds")
```

```{r cwm zone A}
taxa <- dfABUN_a %>% 
  pull(stand.spec) %>% 
  unique()
selectedtrait <- "SLA"

cwm_func <- function(selectedtrait){
# missing trait data
traittax <- lTrait %>%
  pluck(selectedtrait) %>%
  pull(stand.spec) %>% 
  unique
missingtax <- taxa[!taxa %in% traittax]

# filter species with too little observations (2 or less)
nobs <- lTrait %>%
  pluck(selectedtrait) %>% 
  group_by(stand.spec) %>% 
  summarise(n = n()) %>% 
  filter(n <= 2) %>% 
  pull(stand.spec)

Trait <- lTrait %>%
  pluck(selectedtrait) %>%
  filter(!stand.spec %in% nobs) %>% 
  filter(stand.spec %in% taxa) %>%
  pull(., selectedtrait)

Tax <- lTrait %>%
  pluck(selectedtrait) %>%
  filter(!stand.spec %in% nobs) %>% 
  filter(stand.spec %in% taxa) %>%
  pull(stand.spec) %>% 
  as.factor()

Site <- lTrait %>%
  pluck(selectedtrait) %>%
  filter(!stand.spec %in% nobs) %>% 
  filter(stand.spec %in% taxa) %>%
  pull(sitename) %>% 
  as.factor()

N <- length(Trait)

Ab <- dfABUN_a %>%
  # filter species with missing/too little trait data
  filter(!stand.spec %in% missingtax) %>% 
  filter(!stand.spec %in% nobs) %>% 
  # calculate zone level species abundance
  group_by(sitename, stand.spec) %>% 
  summarise(abun = sum(abun)) %>% 
  group_by(sitename) %>% 
  mutate(specsum = sum(abun, na.rm = TRUE)) %>% 
  group_by(sitename, stand.spec) %>% 
  mutate(abun = abun / specsum * 100) %>% 
  dplyr::select(-specsum) %>% 
  # convert to wide format
  pivot_wider(names_from = stand.spec, values_from = abun,
              values_fill = 0) %>% 
  ungroup() %>% 
  dplyr::select(-sitename) %>%
  as.matrix()

Ntax <- ncol(Ab)
Nsites <- nrow(Ab)

# for setting priors taxon level
Mean <- mean(Trait)
SD <- sd(Trait)

mod <- "model {
for (siteID in 1:Nsites){
      cwm[siteID] ~ dnorm(mean.tax[siteID,r[siteID]], tau.tax[siteID,r[siteID]])
      r[siteID] ~ dcat(Ab[siteID, 1:Ntax])

for (j in 1:Nsites){
for (i in 1:N) {
    Trait[j,i] ~ dlnorm(mean.tax[Site[j],Tax[i]], tau.tax[Site[j],Tax[i]])
  }
}
for(j in 1:Nsites){
for (taxID in 1:Ntax){
# vague priors taxon level
    mean.tax[j,taxID] ~ dnorm(0, 10^-4)
    tau.tax[j,taxID] ~ dgamma(0.001, 0.001)
  }
} 
}
#data# N, Nsites, Trait, Tax, Ntax, Ab,Site
#monitor# mean.tax, tau.tax, cwm
#inits# mean.tax, tau.tax
}"

# initial values
mean.tax <- list(chain1 = rep(min(Trait), Ntax), 
                 chain2 = rep(max(Trait), Ntax))
tau.tax <- list(chain1 = rep(1/(SD*0.1)^2,Ntax), 
                chain2 = rep(1/(SD*10)^2, Ntax))

results <- run.jags(mod, n.chains = 2)

# plot(results)

res <- as.matrix(summary(results))
saveRDS(res, paste0("RDS_files/03_CWM_estimates_speciesA_", selectedtrait,".rds"))
}

cwm_func(selectedtrait)
# plan(multisession)
# furrr::future_map(c("SLA","LA","PlantHeight","LDMC"),
#                   ~cwm_func(.), .options = furrr_options(seed = TRUE))

```

```{r cwm zone B}
taxa <- dfABUN_bc %>% 
  pull(stand.spec) %>% 
  unique()

cwm_func <- function(selectedtrait){
# missing trait data
traittax <- lTrait %>%
  pluck(selectedtrait) %>%
  pull(stand.spec) %>% 
  unique
missingtax <- taxa[!taxa %in% traittax]

# filter species with too little observations (2 or less)
nobs <- lTrait %>%
  pluck(selectedtrait) %>% 
  group_by(stand.spec) %>% 
  summarise(n = n()) %>% 
  filter(n <= 2) %>% 
  pull(stand.spec)

Trait <- lTrait %>%
  pluck(selectedtrait) %>%
  filter(!stand.spec %in% nobs) %>% 
  filter(stand.spec %in% taxa) %>%
  pull(selectedtrait)

Tax <- lTrait %>%
  pluck(selectedtrait) %>%
  filter(!stand.spec %in% nobs) %>% 
  filter(stand.spec %in% taxa) %>%
  pull(stand.spec) %>% 
  as.factor()

Site <- lTrait %>%
  pluck(selectedtrait) %>%
  filter(!stand.spec %in% nobs) %>% 
  filter(stand.spec %in% taxa) %>%
  pull(sitename) %>% 
  as.factor()

N <- length(Trait)

Ab <- dfABUN_bc %>% 
  # filter species with missing/too little trait data
  filter(!stand.spec %in% missingtax) %>% 
  filter(!stand.spec %in% nobs) %>% 
  # calculate zone level species abundance
  group_by(sitename, stand.spec) %>% 
  summarise(abun = sum(spec_abun_b)) %>% 
  group_by(sitename) %>% 
  mutate(specsum = sum(abun, na.rm = TRUE)) %>% 
  group_by(sitename, stand.spec) %>% 
  mutate(abun = abun / specsum * 100) %>% 
  select(-specsum) %>% 
  # convert to wide format
  pivot_wider(names_from = stand.spec, values_from = abun,
              values_fill = 0) %>% 
  ungroup() %>% 
  dplyr::select(-sitename) %>%
  as.matrix()

Ntax <- ncol(Ab)
Nsites <- nrow(Ab)

# for setting priors taxon level
Mean <- mean(Trait)
SD <- sd(Trait)

mod <- "model {
for (siteID in 1:Nsites){
      cwm[siteID] ~ dnorm(mean.tax[r[siteID]], tau.tax[r[siteID]])
      r[siteID] ~ dcat(Ab[siteID, 1:Ntax])
    }

for (i in 1:N) {
      Trait[i] ~ dlnorm(mean.tax[Tax[i]], tau.tax[Tax[i]])
}

for (taxID in 1:Ntax){
# vague priors taxon level
    mean.tax[taxID] ~ dnorm(0, 10^-4)
    tau.tax[taxID] ~ dgamma(0.001, 0.001)
}

#data# N, Nsites, Trait, Tax, Ntax, Ab,Site
#monitor# mean.tax, tau.tax, cwm
#inits# mean.tax, tau.tax
}"

# initial values
mean.tax <- list(chain1 = rep(min(Trait), Ntax), 
                 chain2 = rep(max(Trait), Ntax))
tau.tax <- list(chain1 = rep(1/(SD*0.1)^2,Ntax), 
                chain2 = rep(1/(SD*10)^2, Ntax))

results <- run.jags(mod, n.chains = 2)

# plot(results)

res <- as.matrix(summary(results))
saveRDS(res, paste0("RDS_files/03_CWM_estimates_speciesB_",selectedtrait,".rds"))
}

plan(multisession)
furrr::future_map(c("PlantHeight", "SLA", "LDMC", "LA"),
                  ~cwm_func(.), .options = furrr_options(seed = TRUE))

```

```{r cwm zone C}
taxa <- dfABUN_bc %>% 
  pull(stand.spec) %>% 
  unique()

cwm_func <- function(selectedtrait){
# missing trait data
traittax <- lTrait %>%
  pluck(selectedtrait) %>%
  pull(stand.spec) %>% 
  unique
missingtax <- taxa[!taxa %in% traittax]

# filter species with too little observations (2 or less)
nobs <- lTrait %>%
  pluck(selectedtrait) %>% 
  group_by(stand.spec) %>% 
  summarise(n = n()) %>% 
  filter(n <= 2) %>% 
  pull(stand.spec)

Trait <- lTrait %>%
  pluck(selectedtrait) %>%
  filter(!stand.spec %in% nobs) %>% 
  filter(stand.spec %in% taxa) %>%
  pull(selectedtrait)

Tax <- lTrait %>%
  pluck(selectedtrait) %>%
  filter(!stand.spec %in% nobs) %>% 
  filter(stand.spec %in% taxa) %>%
  pull(stand.spec) %>% 
  as.factor()

Site <- lTrait %>%
  pluck(selectedtrait) %>%
  filter(!stand.spec %in% nobs) %>% 
  filter(stand.spec %in% taxa) %>%
  pull(sitename) %>% 
  as.factor()

N <- length(Trait)

Ab <- dfABUN_bc %>% 
  # filter species with missing/too little trait data
  filter(!stand.spec %in% missingtax) %>% 
  filter(!stand.spec %in% nobs) %>% 
  # calculate zone level species abundance
  group_by(sitename, stand.spec) %>% 
  summarise(abun = sum(spec_abun_c)) %>% 
  group_by(sitename) %>% 
  mutate(specsum = sum(abun, na.rm = TRUE)) %>% 
  group_by(sitename, stand.spec) %>% 
  mutate(abun = abun / specsum * 100) %>% 
  select(-specsum) %>% 
  # convert to wide format
  pivot_wider(names_from = stand.spec, values_from = abun,
              values_fill = 0) %>% 
  ungroup() %>% 
  dplyr::select(-sitename) %>%
  as.matrix()

Ntax <- ncol(Ab)
Nsites <- nrow(Ab)

# for setting priors taxon level
Mean <- mean(Trait)
SD <- sd(Trait)

mod <- "model {
for (siteID in 1:Nsites){
      cwm[siteID] ~ dnorm(mean.tax[r[siteID]], tau.tax[r[siteID]])
      r[siteID] ~ dcat(Ab[siteID, 1:Ntax])
    }

for (i in 1:N) {
      Trait[i] ~ dlnorm(mean.tax[Tax[i]], tau.tax[Tax[i]])
}

for (taxID in 1:Ntax){
# vague priors taxon level
    mean.tax[taxID] ~ dnorm(0, 10^-4)
    tau.tax[taxID] ~ dgamma(0.001, 0.001)
}

#data# N, Nsites, Trait, Tax, Ntax, Ab,Site
#monitor# mean.tax, tau.tax, cwm
#inits# mean.tax, tau.tax
}"

# initial values
mean.tax <- list(chain1 = rep(min(Trait), Ntax), 
                 chain2 = rep(max(Trait), Ntax))
tau.tax <- list(chain1 = rep(1/(SD*0.1)^2,Ntax), 
                chain2 = rep(1/(SD*10)^2, Ntax))

results <- run.jags(mod, n.chains = 2)

# plot(results)

res <- as.matrix(summary(results))
saveRDS(res, paste0("RDS_files/03_CWM_estimates_speciesC_",selectedtrait,".rds"))
}

plan(multisession)
furrr::future_map(c("PlantHeight", "SLA", "LDMC", "LA"),
                  ~cwm_func(.), .options = furrr_options(seed = TRUE))

```

```{r compile estimates}
# all taxa
files <- 
  list.files("RDS_files/") %>% 
  str_subset(pattern = "03_CWM_estimates_")

traitname <- files %>% 
  str_remove(., "03_CWM_estimates_pollen_") %>% 
  str_remove(., "03_CWM_estimates_speciesA_") %>% 
  str_remove(., "03_CWM_estimates_speciesB_") %>% 
  str_remove(., "03_CWM_estimates_speciesC_") %>% 
  str_remove(., ".rds")
traitname
zone <- files %>% 
  str_remove(., "03_CWM_estimates_") %>% 
  str_remove(., "_LA.rds") %>% 
  str_remove(., "_LeafC.rds") %>% 
  str_remove(., "_SLA.rds") %>%
  str_remove(., "_PlantHeight.rds") %>%
  str_remove(., "_LDMC.rds") %>%
  str_remove(., "_LeafN.rds")
zone

lab <- files %>% 
  str_remove(., "03_CWM_estimates_") %>% 
  str_remove(., ".rds")

folderpath.fun <- function(x)
{paste("RDS_files/", x, sep = "/")}

dfCWM <- files %>% 
  folderpath.fun(.) %>% 
  purrr::map2(lab, ~readRDS(.) %>% 
               as.data.frame() %>% 
               rownames_to_column("parameter") %>% 
               filter(., str_detect(parameter, "cwm\\[")) %>% 
               dplyr::select(Mean, SD) %>%  
               mutate(label = .y,
                      sitename = seq(1:nrow(.)))) %>% 
  bind_rows() %>% 
  separate(col = label, into = c("zone", "trait"), sep = "_")

saveRDS(dfCWM,"RDS_files/03_Posterior_CWM.rds")
```

### 2022-03-09 01_Prepare_pollen_data.R

@@ -1,0 +1,95 @@
+library(tidyverse)
+library(readxl)
+
+# Scotland ----
+pollen_raw <- read.csv("Data/Pollen_count.csv", skip = 1)
+ppe <- read_xlsx("Data/RPP_Dataset_v1_Table_3_4.xlsx")
+ppe <- tibble(ppe.taxon = ppe$`Target taxon`,
+              PPE = ppe$`RPP v1 (Northern Hemisphere)`)
+
+
+# make tidy
+dfPOL <- pollen_raw %>% 
+  # filter empty rows
+  filter(!Key.index == "---") %>% 
+  filter(!str_detect(Sample.name,"---")) %>% 
+  # select relevant columns
+  select(pollentaxon = Sample.name, contains("C0")) %>% 
+  # convert to long format
+  pivot_longer(cols = contains("C0"), names_to = "sitename",
+               values_to = "count") %>% 
+  # correct miss named site name in polycounter
+  filter(!sitename == "C014") %>% 
+  mutate(sitename = recode(sitename, C014.1 = "C014"),
+         # convert count to numeric       
+         count = as.numeric(count)) %>% 
+  #correct pollentaxon names
+  mutate(pollentaxon = 
+           # remove "-type" to facilitate GBIF search
+           str_remove(pollentaxon, "-type") %>% 
+           # Typos
+           recode(Caryophyllaccea = "Caryophyllaceae",
+                  Lilaceae = "Liliaceae")
+         )
+# join with ppe data
+dfPOL <- dfPOL %>% 
+  mutate(ppe.taxon = case_when(pollentaxon %in% ppe$ppe.taxon ~ pollentaxon,
+                               pollentaxon == "Juniperus communis" ~ "Juniperus",
+                               pollentaxon == "Ericales (tetrad)" ~ "Ericales",
+                               pollentaxon == "Carpinus betulus" ~ "Carpinus",
+                               pollentaxon == "Plantago" ~ "Plantaginaceae",
+                               pollentaxon %in% c("Pteridophyte", "Geraniaceae", "Viola","Pedicularis palustris",
+                                                  "Malvaceae","Tsuga" ) ~ "Not applicable"))
+# calculate pollen percentage
+dfPOL <- dfPOL %>%
+  filter(!pollentaxon == "unindentified") %>% 
+  left_join(ppe, by = "ppe.taxon") %>% 
[…]

### 2022-03-09 01_Prepare_trait_data.R

@@ -1,0 +1,96 @@
+## ---------------------------
+##
+## Script name: Prepare trait data
+##
+## Purpose of script: Cleaning and tidying, and species nomenclature harmonisation
+##
+## Author: Annegreet Veeken
+##
+## Last modified: 08-02-2022
+## ---------------------------
+##
+## Notes:
+##   
+##
+## ---------------------------
+
+# Load libraries
+library(tidyverse)
+library(readxl)
+library(Taxonstand)
+
+## Load data ----
+plantheight_raw <- read_xlsx("Data/Plant_height.xlsx")
+sla_ldmc_raw <- read_xlsx("Data/LDMC_SLA_dataset.xlsx")
+harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx")
+
+## Plant height ----
+# make plant height data tidy
+plantheight <- plantheight_raw %>% 
+  mutate_at(vars(matches("Height")), as.numeric) %>% 
+  # convert to long format
+  pivot_longer(cols = contains("Height"), names_to = "measurement", 
+               values_to = "PlantHeight") %>% 
+  # tidy up measurement column
+  mutate(measurement = str_remove(measurement, "Height_")) %>% 
+  # remove empty rows
+  filter(!is.na(PlantHeight)) %>%
+  # create flowering column based on string
+  separate_rows(Flowering) %>%
+  mutate(flowering = 
+           ifelse(Flowering == measurement, "Flowering","Not-flowering")) %>% 
+  # harmonise column names
+  select(sitename = Site, species = Species, flowering,
+         measurementID = measurement, PlantHeight) %>% 
+  # get rid of the duplicate rows introduced by separate_rows (measurement id avoids deletion of actual duplicates)
+  distinct(sitename, species, measurementID, PlantHeight, .keep_all = TRUE) %>% 
+  # merge with species harmonisation list
+  left_join(harm, by = "species") # check if all are species are available in list
+
+saveRDS(plantheight, "RDS_files/01_PlantHeight.rds")
+
+## SLA ----
+sla <- sla_ldmc_raw %>% 
+  # Harmonize column names
+  select(sitename = Site, species = Species,
+          SLA = `SLA (mm^2 mg−1)`) %>% 
+  # create measurement ID
+  group_by(sitename, species) %>% 
+  mutate(measurementID = row_number()) %>% 
+  # sort
+  arrange(sitename, species, measurementID) %>% 
[…]

### 2022-03-09 02_PollenType_Species_GBIF.R

@@ -1,0 +1,189 @@
+# This script creates a conversion table for pollen type to species based on the current 
+# distribution of terrestrial species belonging to the taxon in the UK
+# Annegreet Veeken
+
+# to do
+# -replace if else by case_when
+
+# libraries
+library(rgbif)
+library(tidyverse)
+library(Taxonstand)
+library(CoordinateCleaner)
+library(readxl)
+
+# Load in pollen data
+dfPOL <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
+
+# extract taxon column for GBIF search
+taxa <- dfPOL %>% 
+  pull(pollentaxon) %>% 
+  unique() 
+taxa <- taxa[!taxa == "unindentified"]
+
+# divide in genera, families and subfamily, to facilitate GBIF search
+sp <- taxa %>% str_subset("[:space:]") 
+sp <- sp[!sp == "Ericales (tetrad)"]
+gn <- taxa %>% str_subset("^(?!.*eae)") %>% 
+  str_subset("[:space:]", negate = TRUE)
+gn <- gn[!gn == "Pteridophyte"]
+fam <- taxa %>% str_subset("ceae")
+ord <-  "Ericales"
+class <- "Polypodiopsida"
+
+# find gbif id's
+# for species
+splist <-
+  purrr::map(sp, ~ name_suggest(., rank = 'SPECIES', limit = 100))
+spkeys <- splist %>% 
+  unlist(recursive = FALSE) %>% 
+  keep(names(.) == "data") %>% 
+  compact() %>% 
+  purrr::map(., ~filter(., !is.na(canonicalName))) %>% 
+  bind_rows() %>% 
+  # only include keys for exact taxon matches, not partial matches (possibly evaluate limit setting in name_suggest)
+  filter(canonicalName %in% sp) 
+
+# for genera
+gnlist <-
+  purrr::map(gn, ~ name_suggest(., rank = 'GENUS', limit = 100))
+# check this list for missing returns
+
+gnkeys <- gnlist %>% 
+  unlist(recursive = FALSE) %>% 
+  keep(names(.) == "data") %>% 
+  compact() %>% 
+  purrr::map(., ~filter(., !is.na(canonicalName))) %>% 
+  bind_rows() %>% 
+  # only include keys for exact taxon matches, not partial matches (possibly evaluate limit setting in name_suggest)
+  filter(canonicalName %in% gn) 
+
+# for family
[…]

### 2022-03-09 04_Correlations.R

@@ -1,0 +1,54 @@
+## ---------------------------
+##
+## Script name: Correlation 
+##
+## Purpose of script: Calculating correlation between CWM species based and pollen based
+##
+## Author: Annegreet Veeken
+##
+## Date created: 08-02-2022
+## ---------------------------
+##
+## Notes:
+##   - re-run when leafC and leafN are available
+##
+## ---------------------------
+
+## Load libraries
+library(tidyverse)
+library(rjags) 
+library(runjags)
+library(mcmc)
+library(coda)
+
+## Load data and prepare
+dfCWM <- readRDS("RDS_files/03_Posterior_CWM.rds") %>% 
+  as_tibble() %>% 
+  filter(!trait %in% c("LeafC", "LeafN")) # not yet available
+
+# Make list per trait, make tidy
+lCWM <- dfCWM %>% 
+  
+  split(f = as.factor(.$trait)) %>% 
+  purrr::map(., ~pivot_wider(., id_cols = sitename, names_from = zone ,
+                             values_from = Mean)) %>% 
+  keep(~ncol(.) == 5) 
+traitnames <- names(lCWM)
+
+## Plots ----
+plot_func <- function(df,traitname){
+p <- ggplot(df, aes(x = pollen, y = speciesA)) +
+  geom_point(col = "darkorange") +
+  geom_point(data = df, aes(x = pollen, y = speciesB), col = "darkorchid") +
+  geom_point(data = df, aes(x = pollen, y = speciesC), col = "cyan4") +
+  scale_x_continuous("Pollen-based CWM", limits = c(min(df[2:5]),max(df[2:5]))) +
+  scale_y_continuous("Vegetation CWM", limits = c(min(df[2:5]),max(df[2:5]))) +
+  ggtitle(traitname) +
+  theme_bw()
+}
+
+pall <- purrr::map2(lCWM,traitnames, ~plot_func(df = .x, traitname = .y))
+#
+
+## Regression bayesian----
+

### 2022-03-09 Harmonization_table.R
First commit

@@ -1,0 +1,31 @@
+# install.packages(c("tidyverse", "xlsx"))
+library(tidyverse)
+library(xlsx)
+
+# Accepted pollen types
+p_vars <- read.csv("Data/EPD data tables/P_VARS.csv", sep = ";")
+# Ecological group names
+p_group <- read.csv("Data/EPD data tables/P_GROUP.csv", sep = ";")
+# Group codes
+group_code <- read.csv("Data/EPD data tables/GROUP_CODES.csv", sep = ";")
+
+# Filter for accepted pollen types names
+acc_varname <- p_vars %>% 
+  select(AccVar., VarName) %>% 
+  distinct(AccVar., .keep_all = TRUE) %>%
+  rename(AccVarName = VarName) %>% 
+# add originale pollen types names
+  left_join(p_vars, by = "AccVar.") %>% 
+  select(AccVar. , AccVarName, Var., VarName) 
+
+# Add ecological groups
+ecol_group <- p_group %>% 
+  left_join(acc_varname, by = "Var.") %>% 
+  left_join(group_code, by = "GroupId") %>% 
+  # select relevant columns and rename
+  select(VarID = Var., GroupID = GroupId, AccVarID = AccVar.,
+        AccVarName, VarName, GroupName)
+
+# write to excel
+write.xlsx(ecol_group, file = "Data/EPD data tables/EPD_ecological_groups.xlsx",
+           row.names = FALSE)

### 2022-03-09 Write_bib_file.R

@@ -1,0 +1,14 @@
+library(bibtex)
+library(tidyverse)
+
+# Compile .bib file multiple references
+ref_files <- list.files("References/")
+
+ref_list <- list()
+
+for(i in 1:length(ref_files)){
+  ref <- read.bib(paste0("References/", ref_files[i]))
+  ref_list <- append(ref_list, ref)
+}
+
+write.bib(ref_list, file ="Validation_study_references")

### 2022-07-13 01_Prepare_trait_data.R
add cn data

@@ -6,22 +6,23 @@
 ##
 ## Author: Annegreet Veeken
 ##
-## Last modified: 08-02-2022
+## Last modified: 11-03-2022
 ## ---------------------------
 ##
 ## Notes:
-##   
+##   - make plots of traits per pollen taxa
+##   - check units of the trait data and the gapfilled data
 ##
 ## ---------------------------
 
 # Load libraries
 library(tidyverse)
 library(readxl)
-library(Taxonstand)
 
 ## Load data ----
 plantheight_raw <- read_xlsx("Data/Plant_height.xlsx")
 sla_ldmc_raw <- read_xlsx("Data/LDMC_SLA_dataset.xlsx")
+cn_raw <- read_xlsx("Data/CN_dataset_150.xlsx")
 harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx")
 
 ## Plant height ----
@@ -71,7 +72,8 @@ ldmc <- sla_ldmc_raw %>%
   distinct(sitename, species, .keep_all = TRUE) %>%  # ldmc has one measurement per species per site
   # create measurement ID
   group_by(sitename, species) %>% 
-  mutate(measurementID = row_number()) %>% 
+  mutate(measurementID = row_number(),
+         LDMC = LDMC/1000) %>% # convert from mg/g to g/g 
   # sort
   arrange(sitename, species, measurementID) %>% 
   # merge with species harmonisation list
@@ -94,3 +96,42 @@ la <- sla_ldmc_raw %>%
 
 saveRDS(la, "RDS_files/01_LA.rds")
     
+## C ----
+# Create lookup table to correct abbreviated species names
+ab_lookup <- tibble(new.name = sla_ldmc_raw$Species) %>% 
+  distinct %>% 
+  separate(., new.name, into = c("genus", "spec"), sep = " ", remove = FALSE) %>% 
+  filter(!is.na(spec)) %>% 
+  mutate(spec = str_to_lower(spec)) %>% 
+  mutate(old.name = paste0(str_extract(genus,pattern = "[:alpha:]"), ".", spec)) %>% 
+  select(new.name, old.name)
+regex_pattern <- setNames(ab_lookup$new.name, paste0("\\b", ab_lookup$old.name, "\\b"))
+cn <- cn_raw %>% 
+  # correct names
+  mutate(Name = str_replace_all(Name, regex_pattern)) %>% 
+  # Harmonize column names
[…]

### 2022-07-14 01_Swiss_data_compilation.R
initial commit

@@ -1,0 +1,141 @@
+## ---------------------------
+##
+## Script name: Swiss data compilation
+##
+## Purpose of script: Compile swiss trait and species information
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-06-30
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   - missing plot numbers
+##   - missing trait data
+##
+## ---------------------------
+
+# Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+if(!require(LCVP)) install.packages("LCVP")
+if(!require(lcvplants)) install.packages("lcvplants")
+
+# Load data
+bdm_meta <- read_csv("Data/bdm_metadata.csv") # contains plot codes, X/Y coordinates (on the Swiss grid), elevation, some basic climate data (MAT, TAP) and a habitat classification
+bdm_sp <- read_csv("Data/bdm_spp_comp.csv") # contains the relative abundances of all species (columns) in each plot/subplot (rows)
+bdm_sp_codes <- read_csv("Data/bdm_spp_codes.csv") # contains the species codes for species IDs in the traits and species composition datasets
+bdm_traits <- read_csv("Data/bdm_traits.csv") #  contains the dry mass, leaf area, SLA and height of plants in each plot/subplot 
+bdm_pollen <- readRDS("RDS_files/01_Pollen_data_Swiss.rds")
+
+# extract plot ids 
+plot_id <- bdm_pollen %>% 
+  pull(sitename) %>% 
+  str_remove(pattern = "X") %>% 
+  unique()
+
+# check sites 
+bdm_meta <- bdm_meta %>% 
+  filter(aID_STAO %in% plot_id)
+
+# Prepare sp code for merge with trait and abundance data
+bdm_sp_codes <- bdm_sp_codes %>% 
+  # create 1 species column
+  mutate(binom = case_when(!is.na(Species) ~ paste(Genus,Species),
+                           !is.na(Genus) ~ paste(Genus, "sp."),
+                           !is.na(Family) ~ paste(Family))
+         ) %>% 
+  # correct spelling mistakes
+  mutate(binom = recode(binom, 
+                        `Trifolum reppens` = "Trifolium repens",
+                        `Cragaetus laevigata` = "Crataegus laevigata",
+                        `Helix hedera` = "Hedera helix",
+                        `Leontondon helveticus` = "Leontodon helveticus",
[…]

### 2022-07-14 02_PollenType_Species_GBIF.R
update code to include swiss data

@@ -1,34 +1,56 @@
-# This script creates a conversion table for pollen type to species based on the current 
-# distribution of terrestrial species belonging to the taxon in the UK
-# Annegreet Veeken
+## ---------------------------
+##
+## Script name: 02_PollenType_Species_GBIF
+##
+## Purpose of script: This script creates a conversion table for pollen type to 
+## species based on the current distribution of terrestrial species belonging to 
+## the taxon in the UK/Switzerland
+##
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-06-30
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+## 
+## - area of species selection
+## - evaluate limit setting occ_search
+##   
+## ---------------------------
 
-# to do
-# -replace if else by case_when
 
-# libraries
-library(rgbif)
-library(tidyverse)
-library(Taxonstand)
-library(CoordinateCleaner)
-library(readxl)
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+if(!require(rgbif)) install.packages("rgbif")
+if(!require(LCVP)) devtools::install_github("idiv-biodiversity/LCVP")
+if(!require(lcvplants)) devtools::install_github("idiv-biodiversity/lcvplants")
+if(!require(CoordinateCleaner)) install.packages("CoordinateCleaner")
+
 
 # Load in pollen data
-dfPOL <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
+dfPOL_scot <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
+dfPOL_swiss <- readRDS("RDS_files/01_Pollen_data_Swiss.rds")
+dfPOL <- bind_rows(dfPOL_scot, dfPOL_swiss)
 
 # extract taxon column for GBIF search
 taxa <- dfPOL %>% 
   pull(pollentaxon) %>% 
   unique() 
 taxa <- taxa[!taxa == "unindentified"]
+taxa <- c(taxa, "Oxyria", "Rumex")
 
 # divide in genera, families and subfamily, to facilitate GBIF search
 sp <- taxa %>% str_subset("[:space:]") 
-sp <- sp[!sp == "Ericales (tetrad)"]
+sp <- sp[!sp == "Ericales (tetrad)"] # replaced by relevant tetrad ericales Ericaceae and empetraceae
[…]

### 2022-07-24 03_CWM_estimates_pollen.R
Split CWM script in vegetation and pollen specific script

@@ -1,0 +1,369 @@
+## ---------------------------
+##
+## Script name: 03_CWM_estimates_pollen.R
+##
+## Purpose of script: Calculate CWM values of pollen data
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-07-18
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   
+##
+## ---------------------------
+
+
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+if(!require(rjags))  install.packages("rjags")
+if(!require(runjags)) install.packages("runjags")
+if(!require(mcmc)) install.packages("mcmc")
+if(!require(coda)) install.packages("coda")
+if(!require(Hmisc)  )install.packages("Hmisc") 
+if(!require(furrr))  install.packages("furrr")
+if(!require(LCVP)) install.packages("LCVP")
+if(!require(lcvplants)) install.packages("lcvplants")
+
+## Load and prepare data -----
+# Trait data ----
+dfTRAIT <- readRDS("RDS_files/02_Gapfilled_traits.rds")
+
+# Pollen data ----
+dfPOL_Scot <- readRDS("RDS_files/01_pollen_data_Scot.rds")
+dfPOL_Swiss <- readRDS("RDS_files/01_pollen_data_Swiss.rds")
+dfPOL <- bind_rows("Scotland" = dfPOL_Scot, "Switzerland" =
+                     dfPOL_Swiss, .id = "country") %>% 
+  ungroup()
+
+selectedcountry <- "Scotland"
+selectedtrait <- "LA"
+selectedabun <- str_subset(colnames(dfPOL), "percent")[57]
+# Calculate CWM ----
+cwmlogfunc <- function(selectedtrait, selectedcountry, selectedabun){
+  pol <- dfPOL %>% 
+    # select country
+    filter(country == selectedcountry) %>%
+    dplyr::select(sitename, pollentaxon, pollen_abun = all_of(selectedabun)) %>% 
+    # filter out unidentified
+    filter(!pollentaxon == "unindentified",
+           !is.na(pollen_abun)) %>% 
+    pivot_wider(names_from = pollentaxon, values_from = pollen_abun) %>% 
+    # filter empty columns (taxon not in that country)
+    discard(~all(is.na(.))) 
+  
+  # save sitenames
+  sitenames <- pol %>% 
+    pull(sitename)
+  # save pollen taxa
[…]

### 2022-07-24 03_CWM_estimates_vegetation.R

@@ -1,0 +1,579 @@
+## ---------------------------
+##
+## Script name:03_CWM_estimates_vegetation.R
+##
+## Purpose of script: Calculating CWM values of the vegetation
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-07-18
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   
+##
+## ---------------------------
+
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+if(!require(rjags))  install.packages("rjags")
+if(!require(runjags)) install.packages("runjags")
+if(!require(mcmc)) install.packages("mcmc")
+if(!require(coda)) install.packages("coda")
+if(!require(Hmisc)  )install.packages("Hmisc") 
+if(!require(furrr))  install.packages("furrr")
+if(!require(LCVP)) install.packages("LCVP")
+if(!require(lcvplants)) install.packages("lcvplants")
+
+set.seed(1)
+
+## Load and prepare data -----
+# Trait data ----
+# Scotland ----
+plantheight <- readRDS("RDS_files/01_PlantHeight.rds") %>% 
+  mutate(sitename = as.factor(sitename), species = as.factor(species),
+         PlantHeight = PlantHeight) %>% # plantheight in centimeter
+  ungroup()
+ldmc <- readRDS("RDS_files/01_LDMC.rds") %>% 
+  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
+  ungroup()
+sla <- readRDS("RDS_files/01_SLA.rds") %>% 
+  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
+  ungroup()
+la <- readRDS("RDS_files/01_LA.rds") %>% 
+  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
+  ungroup()
+c <- readRDS("RDS_files/01_CN.rds") %>% 
+  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
+  ungroup() %>% 
+  rename(LeafC = C)
+n <- readRDS("RDS_files/01_CN.rds") %>% 
+  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
+  ungroup() %>% 
+  rename(LeafN = N)
+
+lTrait <- list(PlantHeight = plantheight,
+               LDMC = ldmc,
+               SLA = sla,
[…]

### 2022-07-24 01_Prepare_pollen_data.R
object rename

@@ -29,8 +29,8 @@ set.seed(1)
 # Scotland ----
 pollen_raw <- read.csv("Data/Pollen_count.csv", skip = 1)
 ppe <- read_xlsx("Data/Githumbi_2022_RPP.xlsx")
-pollenmode <- readRDS("RDS_files/01_Pollination_mode.rds")
-  
+polmode <- readRDS("RDS_files/01_Pollination_mode.rds")
+
 # make tidy
 dfPOL <- pollen_raw %>% 
   # filter empty rows
@@ -73,6 +73,16 @@ dfPOL_nc <- dfPOL %>%
   group_by(sitename, pollentaxon) %>%
   summarise(percent = sum(percent)) 
 
+# make function to continue draw from rnorm till RPP is positive (otherwise negative percentages)
+norm_positive <- function(RPP, RPP_sd){
+  success <- FALSE
+  while(!success){
+    x <- rnorm(1, RPP, RPP_sd)
+    success <- x > 0
+    }
+  return(x)
+}
+
 n <- 100 # number of draws from RPP distribution
 dfPOL_cor <- dfPOL %>%
   # join with ppe data 
@@ -83,7 +93,7 @@ dfPOL_cor <- dfPOL %>%
   mutate(mean = count/RPP) %>%  
   # correction with random draw from RPP distribution
   rowwise() %>%
-  mutate(col_rnorm = list(setNames(count/rnorm(n, RPP, RPP_sd), paste0("draw", 1:n)))) %>%
+  mutate(col_rnorm = list(setNames(count/replicate(n,norm_positive(RPP, RPP_sd)), paste0("draw",1:n)))) %>%
   unnest_wider(col_rnorm) %>% 
   # calculate percentages
   group_by(sitename) %>% 
@@ -95,7 +105,7 @@ dfPOL_cor <- dfPOL %>%
 dfPOL <- dfPOL_nc %>% 
   left_join(dfPOL_cor, by = c("sitename", "pollentaxon")) %>% 
   # add PFT and pollination mode
-  left_join(pollenmode, by = "pollentaxon") %>% 
+  left_join(polmode, by = "pollentaxon") %>% 
   ungroup()
 
 saveRDS(dfPOL, "RDS_files/01_Pollen_data_Scot.rds")
@@ -137,7 +147,6 @@ dfPOL_nc <- dfPOL %>%
   group_by(sitename, pollentaxon) %>%
[…]

### 2022-07-24 01_Prepare_trait_data.R
select to dplyr::select

@@ -23,7 +23,8 @@ library(readxl)
 plantheight_raw <- read_xlsx("Data/Plant_height.xlsx")
 sla_ldmc_raw <- read_xlsx("Data/LDMC_SLA_dataset.xlsx")
 cn_raw <- read_xlsx("Data/CN_dataset_150.xlsx")
-harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx")
+harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx") %>% 
+  rename(family = fam)
 
 ## Plant height ----
 # make plant height data tidy
@@ -41,7 +42,7 @@ plantheight <- plantheight_raw %>%
   mutate(flowering = 
            ifelse(Flowering == measurement, "Flowering","Not-flowering")) %>% 
   # harmonise column names
-  select(sitename = Site, species = Species, flowering,
+  dplyr::select(sitename = Site, species = Species, flowering,
          measurementID = measurement, PlantHeight) %>% 
   # get rid of the duplicate rows introduced by separate_rows (measurement id avoids deletion of actual duplicates)
   distinct(sitename, species, measurementID, PlantHeight, .keep_all = TRUE) %>% 
@@ -53,7 +54,7 @@ saveRDS(plantheight, "RDS_files/01_PlantHeight.rds")
 ## SLA ----
 sla <- sla_ldmc_raw %>% 
   # Harmonize column names
-  select(sitename = Site, species = Species,
+  dplyr::select(sitename = Site, species = Species,
           SLA = `SLA (mm^2 mg−1)`) %>% 
   # create measurement ID
   group_by(sitename, species) %>% 
@@ -67,7 +68,7 @@ saveRDS(sla, "RDS_files/01_SLA.rds")
 
 ## LDMC ----
 ldmc <- sla_ldmc_raw %>% 
-  select(sitename = Site, species = Species,
+  dplyr::select(sitename = Site, species = Species,
          LDMC = "LDMC (mg g–1)") %>% 
   distinct(sitename, species, .keep_all = TRUE) %>%  # ldmc has one measurement per species per site
   # create measurement ID
@@ -83,7 +84,7 @@ saveRDS(ldmc, "RDS_files/01_LDMC.rds")
 
 ## LA ----
 la <- sla_ldmc_raw %>% 
-  select(sitename = Site, species = Species,
+  dplyr::select(sitename = Site, species = Species,
[…]

### 2022-07-24 01_Species_abundance.Rmd
tidy script
select to dplyr::select

@@
 ## Prepare data
+
 Vegetation cover data.
+
 ```{r load and prepare veg cover data}
@@
 zone_b <- zone_b %>% 
-  select(sitename = name, lat, lon, contains("HISTO")) %>% 
+  dplyr::select(sitename = name, lat, lon, contains("HISTO")) %>% 
   pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
@@
 zone_c <- zone_c %>% 
-  select(sitename = name, lat, lon, contains("HISTO")) %>% 
+  dplyr::select(sitename = name, lat, lon, contains("HISTO")) %>% 
   pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
@@
   mutate(veg_type = as.factor(rvalue_1)) %>% 
-  select(sitename, plot, lat, lon, veg_type) %>% 
+  dplyr::select(sitename, plot, lat, lon, veg_type) %>% 
   mutate(veg_type = recode(veg_type, 
@@
   mutate(plot = str_remove(plot, "plot_ ")) %>% 
-  select(sitename = Site, plot, species = Species, abun) %>% 
+  dplyr::select(sitename = Site, plot, species = Species, abun) %>% 
   # join with vegetation type data
@@
   # drop unneeded columns
-  select(sitename, veg_type, species, veg_cov_b, veg_cov_c, spec_abun_b,
-         spec_abun_c)
+  dplyr::select(sitename, veg_type, species, veg_cov_b, veg_cov_c,
+                spec_abun_b,spec_abun_c)
 
@@
   filter(is.na(species)) %>%
-  select(-species) %>% 
+  dplyr::select(-species) %>% 
   left_join(missing, by = "veg_type") %>% 
   mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c) %>% 
-  select(sitename, veg_type, species, veg_cov_b, veg_cov_c, spec_abun_b,
+  dplyr::select(sitename, veg_type, species, veg_cov_b, veg_cov_c, spec_abun_b,
          spec_abun_c)
@@
 ## Standardization species names
+
 ```{r standardize species names}
 harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx") 
-
 spec_a <- spec_a %>% 

### 2022-07-24 04_Correlation_bayesian.R
add script

@@ -1,0 +1,135 @@
+## ---------------------------
+##
+## Script name: 
+##
+## Purpose of script:
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-07-24
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   
+##
+## ---------------------------
+
+
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+if(!require(rjags)) install.packages("rjags")
+if(!require(runjags)) install.packages("runjags")
+if(!require(mvtnorm)) install.packages("mvtnorm") #to generate correlated data with rmvnorm.
+if(!require(car)) install.packages("car") #To plot the estimated bivariate normal distribution.
+
+# collate data
+traits <- c("LA","SLA","PlantHeight")
+# all taxa
+files <- 
+  list.files("RDS_files/") %>% 
+  str_subset(pattern = "03_CWM_estimates_") %>% 
+  str_subset(pattern = "adjusted", negate = TRUE) %>% 
+  str_subset(paste(traits, collapse = "|"))
+
+lab <- files %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
+  str_subset(., "multivariate", negate = TRUE) %>% 
+  str_remove(., "03_CWM_estimates_") %>% 
+  str_remove(., ".rds")
+
+folderpath.fun <- function(x)
+{paste("RDS_files/", x, sep = "/")}
+
+dfCWM <- files %>% 
+  folderpath.fun(.) %>% 
+  purrr::map2(lab, ~readRDS(.) %>% 
+                as.data.frame() %>% 
+                rownames_to_column("parameter") %>% 
+                filter(., str_detect(parameter, "cwm\\[")) %>% 
+                dplyr::select(Mean, SD) %>%  
+                mutate(label = .y,
+                       id = seq(1:nrow(.))
+                )
+  ) %>% 
+  bind_rows() %>% 
+  mutate(trait = str_extract(label, pattern = paste(traits, collapse = "|")) %>% 
+           str_remove_all("_"),
+         zone = str_extract(label, pattern = "pollen|zoneA|zoneB|zoneC"),
+         growthform = case_when(str_detect(label, pattern = "grassherbfernshrubtreeNA") ~ "all",
+                                str_detect(label, pattern = "grassherbfern") ~ "herb",
[…]

### 2022-07-24 01_Swiss_data_compilation.R
remove excess space

@@ -116,6 +116,7 @@ bdm_sp <- bdm_sp %>%
 
 saveRDS(bdm_sp, "RDS_files/01_Species_abundance_Swiss.rds")
 
+
 ## ---- Traits
 bdm_traits <- bdm_traits %>% 
   left_join(bdm_sp_codes, by = c("species_code" = "Species_code")) %>% 

### 2022-07-24 02_Prepare_trait_data_gapfilled.R
add script

@@ -1,0 +1,319 @@
+## ---------------------------
+##
+## Script name: 02_prepare_trait_data_gapfilled.R
+##
+## Purpose of script: Prepare gapfilled data for CWM calculation
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-07-19
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   
+##
+## ---------------------------
+
+
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+
+## Add pollen type to gapdata
+# Species per pollen type list
+spec <- readRDS("RDS_files/02_PollenType_species.rds")
+spec <- spec  %>% 
+  # missing species
+  add_row(family = "Plantaginaceae", genus = "Digitalis", species = "Digitalis purpurea",
+          count = NA, stand.spec = "Digitalis purpurea", country = "Scotland", pollentaxon = NA) %>% 
+  add_row(family = "Plantaginaceae", genus = "Digitalis", species = "Digitalis purpurea",
+          count = NA, stand.spec = "Digitalis purpurea", country = "Switzerland", pollentaxon = NA) %>% 
+  add_row(family = "Primulaceae", genus = "Lysimachia", species = "Lysimachia nemorum",
+          count = NA, stand.spec = "Lysimachia nemorum", country = "Scotland", pollentaxon = NA) %>% 
+  add_row(family = "Primulaceae", genus = "Lysimachia", species = "Lysimachia nemorum",
+          count = NA, stand.spec = "Lysimachia nemorum", country = "Switzerland", pollentaxon = NA) %>% 
+  add_row(family = "Juncaceae", genus = "Luzula", species = "Luzula sylvatica",
+          count = NA, stand.spec = "Luzula sylvatica", country = "Scotland", pollentaxon = NA) %>% 
+  add_row(family = "Juncaceae", genus = "Luzula", species = "Luzula sylvatica",
+          count = NA, stand.spec = "Luzula sylvatica", country = "Switzerland", pollentaxon = NA) %>% 
+  add_row(family = "Brassicaceae", genus = "Cardamine", species = "Cardamine pratensis",
+          count = NA, stand.spec = "Cardamine pratensis", country = "Scotland") %>% 
+  add_row(family = "Brassicaceae", genus = "Cardamine", species = "Cardamine pratensis",
[…]

### 2022-07-24 03_Pollen_diagrams.R

@@ -1,0 +1,175 @@
+## ---------------------------
+##
+## Script name: Pollen diagrams
+##
+## Purpose of script: Plot pollen diagrams
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-03-24
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   
+##
+## ---------------------------
+
+
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+if(!require(rioja)) install.packages("rioja") # plot pollen diagrams
+if(!require(ggrepel)) install.packages("ggrepel") # repel point labels
+if(!require(readxl)) install.packages("readxl") # repel point labels
+
+# - compare representation of pollen in vegetation
+# - classify by pollination mode
+# - classify by woody/non-woody
+# - compare with and without correction factors
+
+
+## Prepare data ----
+pollen <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
+polmode <- read_xlsx("Data/Pollination_modes_Reitalu_2019.xlsx")
+polspec <- readRDS("RDS_files/02_PollenType_species.rds") %>% 
+  select(pollentaxon, stand.spec)
+
+# Add pollination mode and plant functional type
+# Harmonize nomenclature
+polmode <- polmode %>% 
+  mutate(pollentaxon = recode(pollentaxon,  "Apiaceae undiff." = "Apiaceae",
+                              "Caryophyllaceae undiff." = "Caryophyllaceae",
+                              "Cyperaceae undiff." = "Cyperaceae",
+                              "Ericacea-type" = "Ericales (tetrad)",
+                              "Fraxinus excelsior" = "Fraxinus",
+                              "Geranium" = "Geraniaceae",
+                              "Juniperus" = "Juniperus communis",
+                              "Pedicularis" = "Pedicularis palustris",
+                              "Ranunculuceae undiff." = "Ranunculaceae",
+                              "Rosaceae undiff." = "Rosaceae",
+                              "Viola palustris-type" = "Viola")) %>% 
[…]

### 2022-07-24 03_Vegatation_pollen_comparison.R

@@ -1,0 +1,142 @@
+## ---------------------------
+##
+## Script name: Vegetation_pollen_comparison
+##
+## Purpose of script: Compare representation of species in the vegetation in the pollen
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-03-21
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   - 
+##     
+## ---------------------------
+
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+if(!require(rioja)) install.packages("rioja") # plot pollen diagrams
+if(!require(ggrepel)) install.packages("ggrepel") # repel point labels
+if(!require(readxl)) install.packages("readxl") # repel point labels
+
+# - compare representation of pollen in vegetation
+# - classify by pollination mode
+# - classify by woody/non-woody
+# - compare with and without correction factors
+
+
+## Prepare data ----
+dfPOL_Scot <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
+dfPOL_Swiss <- readRDS("RDS_files/01_Pollen_data_Swiss.rds")
+dfPOL <- bind_rows(dfPOL_Scot, dfPOL_Swiss)
+
+polspec <- readRDS("RDS_files/02_PollenType_species.rds") %>% 
+  dplyr::select(pollentaxon, stand.spec) %>% as_tibble()
+dfABUN_a <- readRDS("RDS_files/01_Species_abundance_a.rds")
+dfABUN_bc <- readRDS("RDS_files/01_Species_abundance_bc.rds") %>% ungroup
+bdm_abun <- readRDS("RDS_files/01_Species_abundance_Swiss.rds") %>% 
+  left_join(polspec, by = c("binom" = "stand.spec"))
+
+# Summarize vegetation data by pollen type
+zoneA <- dfABUN_a %>%
+  # select plots that are comparable with Switzerland
+  filter(distance %in% c("0 meter", "1.5-3 meter")) %>% 
+  # average species abundance on the zone level
+  group_by(sitename,pollentaxon) %>% 
+  summarise(abun = mean(abun)/100) %>%  
+  mutate(zone = "zoneA",
+         country = "Scotland")
+zoneA_swiss <- bdm_abun %>% 
+  group_by(plot, pollentaxon) %>% 
+  summarise(abun = mean(abun)) %>% 
+  rename(sitename = plot)%>%  
[…]

### 2022-07-24 04_Correlations.R

@@ -1,6 +1,6 @@
 ## ---------------------------
 ##
-## Script name: Correlation 
+## Script name: Correlations 
 ##
 ## Purpose of script: Calculating correlation between CWM species based and pollen based
 ##
@@ -20,35 +20,119 @@ library(rjags)
 library(runjags)
 library(mcmc)
 library(coda)
+library(gridExtra)
 
 ## Load data and prepare
-dfCWM <- readRDS("RDS_files/03_Posterior_CWM.rds") %>% 
-  as_tibble() %>% 
-  filter(!trait %in% c("LeafC", "LeafN")) # not yet available
+dfCWM <- readRDS("RDS_files/03_Posterior_CWM.rds") 
 
-# Make list per trait, make tidy
-lCWM <- dfCWM %>% 
-  
-  split(f = as.factor(.$trait)) %>% 
-  purrr::map(., ~pivot_wider(., id_cols = sitename, names_from = zone ,
-                             values_from = Mean)) %>% 
-  keep(~ncol(.) == 5) 
-traitnames <- names(lCWM)
+# Make df per trait, make tidy
+lCWM <- dfCWM %>%
+  dplyr::select(-id) %>% 
+  group_by(trait) %>% 
+  group_split(.keep = TRUE) %>% 
+  purrr::map(., ~pivot_wider(., id_cols = c(trait, sitename, country), names_from = zone ,
+                             values_from = Mean))
+
+
+
+# dfCWM <- readRDS("RDS_files/03_Posterior_CWM.rds") %>%
+#   as_tibble() %>%
+#   pivot_wider(names_from = zone, values_from = c(Mean,SD))
+# 
+# dfCWM %>% 
+#   filter(trait == "SLA") %>% 
+#   ggplot(aes(x = exp(Mean_pollencorrected), y = Mean_zoneA)) +
+#   geom_point(col = "darkorange") +
+#   geom_smooth(method = "lm", col = "darkorange") 
+# 
+# 
+# ggplot(dfCWM, aes(x = Mean_pollencorrected, y = Mean_zoneB)) +
+#   geom_point(col = "darkorchid") +
+#   geom_smooth(method = "lm", col = "darkorchid")+
+#   facet_wrap(~trait)
+# 
+# ggplot(dfCWM, aes(x = Mean_pollencorrected, y = Mean_speciesC)) +
+#   geom_point(col = "cyan4") +
+#   geom_smooth(method = "lm", col = "cyan4")+
[…]

### 2022-07-26 02_PollenType_Species_GBIF.R
fix species-pollentype table

@@ -17,7 +17,6 @@
 ##
 ## Notes:
 ## 
-## - area of species selection
 ## - evaluate limit setting occ_search
 ##   
 ## ---------------------------
@@ -112,7 +111,7 @@ countries <- c("United Kingdom", "Switzerland")
 countrykeys <- isocodes %>% filter(name %in% countries) %>% pull(code)
 
 ## create function  that downloads, cleans, saves and creates conversion table for pollen type to species
-
+if(0){
 gbifsaveclean <- function(taxonkey, countrykey){
   # perform gbif search
   occsearch <- occ_search(taxonKey = taxonkey,
@@ -191,6 +190,9 @@ spec_ch <- purrr::map(taxonkeys,
 
 spec <- bind_rows("Scotland" = spec_gb, "Switzerland" = spec_ch, .id = "country")
 saveRDS(spec, "RDS_files/02_gbif_raw.rds")
+} else{
+spec <- read_rds("RDS_files/02_gbif_raw.rds")
+}
 
 # combine to make table, add pollen taxon
 specpol <- spec %>% 
@@ -221,6 +223,144 @@ specpol <- specpol %>%
                               .$spec == "Rabelera holostea",
                               values = "Stellaria holostea")) %>% 
   # remove na's (after checking for synonyms manually, all cultivars)
-  filter(!is.na(stand.spec))  
+  filter(!is.na(stand.spec)) %>% 
+  as_tibble()
+
+# some species recorded in the field not retrieved by gbif
+dfABUN_a <- readRDS("RDS_files/01_Species_abundance_a.rds")
+dfABUN_bc <- readRDS("RDS_files/01_Species_abundance_bc.rds") %>% ungroup
+bdm_abun <- readRDS("RDS_files/01_Species_abundance_Swiss.rds") 
+
+sp_extra <- c(dfABUN_a$stand.spec, dfABUN_bc$stand.spec, bdm_abun$stand.spec) %>% 
+  unique %>% 
+  .[!. %in% unique(specpol$stand.spec)]
+
+lcvp_out <- lcvp_search(str_subset(sp_extra, " |sp.")) %>% as_tibble()
+lcvp <- lcvp_out %>% 
+  mutate(genus = word(Output.Taxon, 1),
+         stand.spec = word(Output.Taxon, 1,2)) %>% 
+  dplyr::select(family = Family, genus , species = Search, stand.spec) %>% 
+  drop_na() %>% 
[…]

### 2022-07-26 02_Prepare_trait_data_gapfilled.R

@@ -24,206 +24,6 @@ if(!require(tidyverse)) install.packages("tidyverse")
 ## Add pollen type to gapdata
 # Species per pollen type list
 spec <- readRDS("RDS_files/02_PollenType_species.rds")
-spec <- spec  %>% 
-  # missing species
-  add_row(family = "Plantaginaceae", genus = "Digitalis", species = "Digitalis purpurea",
-          count = NA, stand.spec = "Digitalis purpurea", country = "Scotland", pollentaxon = NA) %>% 
-  add_row(family = "Plantaginaceae", genus = "Digitalis", species = "Digitalis purpurea",
-          count = NA, stand.spec = "Digitalis purpurea", country = "Switzerland", pollentaxon = NA) %>% 
-  add_row(family = "Primulaceae", genus = "Lysimachia", species = "Lysimachia nemorum",
-          count = NA, stand.spec = "Lysimachia nemorum", country = "Scotland", pollentaxon = NA) %>% 
-  add_row(family = "Primulaceae", genus = "Lysimachia", species = "Lysimachia nemorum",
-          count = NA, stand.spec = "Lysimachia nemorum", country = "Switzerland", pollentaxon = NA) %>% 
-  add_row(family = "Juncaceae", genus = "Luzula", species = "Luzula sylvatica",
-          count = NA, stand.spec = "Luzula sylvatica", country = "Scotland", pollentaxon = NA) %>% 
-  add_row(family = "Juncaceae", genus = "Luzula", species = "Luzula sylvatica",
-          count = NA, stand.spec = "Luzula sylvatica", country = "Switzerland", pollentaxon = NA) %>% 
-  add_row(family = "Brassicaceae", genus = "Cardamine", species = "Cardamine pratensis",
-          count = NA, stand.spec = "Cardamine pratensis", country = "Scotland") %>% 
-  add_row(family = "Brassicaceae", genus = "Cardamine", species = "Cardamine pratensis",
-          count = NA, stand.spec = "Cardamine pratensis", country = "Switzerland") %>% 
-  add_row(family = "Ophioglossaceae", genus = "Botrychium", species = "Botrychium lunaria",
-          count = NA, stand.spec = "Botrychium lunaria", country = "Scotland") %>% 
-  add_row(family = "Ophioglossaceae", genus = "Botrychium", species = "Botrychium lunaria",
[…]

### 2022-07-26 03_Vegetation_pollen_comparison.R
fix spelling mistake

@@ -23,50 +23,63 @@ if(!require(rioja)) install.packages("rioja") # plot pollen diagrams
 if(!require(ggrepel)) install.packages("ggrepel") # repel point labels
 if(!require(readxl)) install.packages("readxl") # repel point labels
 
-# - compare representation of pollen in vegetation
-# - classify by pollination mode
-# - classify by woody/non-woody
-# - compare with and without correction factors
-
-
 ## Prepare data ----
 dfPOL_Scot <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
 dfPOL_Swiss <- readRDS("RDS_files/01_Pollen_data_Swiss.rds")
 dfPOL <- bind_rows(dfPOL_Scot, dfPOL_Swiss)
-
+polmode <- readRDS("RDS_files/01_Pollination_mode.rds")
 polspec <- readRDS("RDS_files/02_PollenType_species.rds") %>% 
-  dplyr::select(pollentaxon, stand.spec) %>% as_tibble()
+  dplyr::select(pollentaxon, stand.spec) %>% distinct() %>% as_tibble()
 dfABUN_a <- readRDS("RDS_files/01_Species_abundance_a.rds")
 dfABUN_bc <- readRDS("RDS_files/01_Species_abundance_bc.rds") %>% ungroup
-bdm_abun <- readRDS("RDS_files/01_Species_abundance_Swiss.rds") %>% 
-  left_join(polspec, by = c("binom" = "stand.spec"))
+bdm_abun <- readRDS("RDS_files/01_Species_abundance_Swiss.rds") 
 
 # Summarize vegetation data by pollen type
 zoneA <- dfABUN_a %>%
+  # join with pollen to species table
+  left_join(polspec, by = "stand.spec") %>% 
+  mutate(pollentaxon = if_else(is.na(pollentaxon), "Not in pollen data", pollentaxon)) %>% 
   # select plots that are comparable with Switzerland
   filter(distance %in% c("0 meter", "1.5-3 meter")) %>% 
-  # average species abundance on the zone level
-  group_by(sitename,pollentaxon) %>% 
-  summarise(abun = mean(abun)/100) %>%  
+  # calculate species abundance on the plot level
+  group_by(sitename, pollentaxon) %>% 
+  summarise(abun = sum(abun)) %>% 
+  mutate(abun = abun/sum(abun)) %>% 
[…]

### 2022-07-27 Manuscript.Rmd
add rmd of manuscript

---
output:
  word_document:
    reference_docx: Manuscript-Markdown-template.docx
    fig_caption: yes
bibliography: Validation_study_references.bib
---

```{r setup, include=FALSE}
# Document options
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      collapse = TRUE,
                      results = "hide")

source("Write_bib_file.R") # Compile library for reference list

# Libraries
library(bibtex)
library(knitr)
```

# Testing the pollen - functional diversity relationship

-   pre-print?

## Introduction

-say something about shifts in plant composition in the Anthropocene and ecosystem processes- In the past two decades, plant functional traits have become an important tool in ecology and biogeography for the study of consequences of changing plant composition on ecosystem functioning and the effect of environment factors on plant composition [@Hevia_2017]. Recently, considerable interest in the trait-based approach has arisen in palaeoecology as well [@Brussel_2018; @Blaus_2020; @Carvalho_2019; @Reitalu_2015; @Sande_2019]. Applying the trait-based approach could offer a new way of interpreting pollen data and could increase our understanding of the drivers of vegetation change on timescales of centuries to millennial. For instance, @Sande_2019 combined functional trait and pollen data to study the effect of climate and human disturbances on plant trait composition in the Amazon on a 7000-year time scale.

Although palaeoecological studies that have applied the trait-based approach show promising results, it is a relatively new approach of which the intricacies are underexplored. Pollen quantity does not directly represent past plant abundance, taxa can become over- or underrepresented in the pollen record because of differences in pollen production and dispersal. -something about pollination syndrome and plant function type - To improve our understanding of the relationship between the trait composition of the vegetation and trait compositions reconstructed from pollen records, we here use surface pollen samples. Surface pollen assemblages approximate vegetation composition in recent years and are commonly used in palaeoecology to verify pollen-vegetation relationships [@Birks_2016]. Here, we use surface samples of 16 Scottish and 27 Swiss sites to test the relationship between pollen trait composition and vegetation trait composition.

To correct for these differences pollen productivity factors can be use, these correction factors down-weight the contribution of dominant taxa. Furthermore, comparing between functional groups based on growth form or polination syndrome could give insights in to the representativeness of the pollen assemblage to the vegetation.

Perhaps the most challenging in reliably reconstructing past trait is to attribute functional trait data to the taxa found in pollen records, which can rarely be resolved to the species level (Birks 2020). Trait variation within taxa can be substantial (REF). It can be expected that this especially problematic in the case of large, abundandt plant families that cannot easily be resolved to a lower taxonomic level, such as the Poaceae family. Also small families where the trait variation is high could be problematic, such as is the case with the Fabaceae family where the height variation can be ...-fold (for instance Ulex shrub height and clover). Vegetation types with a large proportion of taxa with large trait variation might therefore be less reliably reconstructed. One way of dealing with the uncertainty caused by trait variation in pollen taxa is by using Bayesian models. Bayesian models allow you to do this and this.

With this study, we aim to test the validity of using of pollen records for plant trait composition reconstructions. More specifically, we will test the effect of vegetation composition and trait choice on the representation of pollen assemblages for functional diversity reconstructions. We examine the effectivity of pollen production factors to improve the pollen - functional diversity relationship. Lastly, we test the representation of plant functional groups based on growth form and pollination syndrome in the functional composition reconstruction.

-   resolution of the pollen record - calculate CWM using different taxonomic levels of the vegetation?

-   trait choice

-   assignment of species to pollen taxa

-   representation of pft's - pollination syndrome

## Method

### Study design

Pollen data, plant species composition data and trait data were collected in 16 sampling locations in June 2021 in the Cairngorms National Park, Scotland, United Kingdom. Study design mostly followed the protocol of [@Bunting_2013]. Sampling locations were at least 1 km from each other to avoid pollen source areas from overlapping with each other. Sites were selected to represent four vegetation types in the area: grassland, heath land, deciduous forest and coniferous forest. Every sampling location consisted of three nested concentric rings of 3m, 100m and 1000m in radius. The inner two rings were surveyed in the field and for the outer ring vegetation cover was estimated using QGIS and the Scottish Habitat and Land cover map of 2020 . To determine the composition of the modern pollen assemblage at every site, surface samples of moss were collected at the center of each sampling location. Moss was collected from multiple location within a 1m radius from the center. Only living tissue was sampled.

An addition 27 pollen samples were collected from sites in the Swiss Biodiversity Monitoring network, that were selected based on an elevation gradient (?). Only the inner ring was surveyed in these sites

### Vegetation surveys

In the inner ring (0-10m) 17 plots of 0.5 m were surveyed: one in the center, 3 in every wind direction between 3, 6 and 10 meters, and one in the NE, NW, SE and SW direction (See design in Appendix). In the middle ring (10-100m) between 4 and 6 plots of 10 by 10? meter were randomly selected. Vegetation abundance was estimated using percentage cover.

### Trait sampling

-   Plant height, SLA and LA measured in the field

### Assigning species to pollen taxa

-   Species list of Switzerland and Scotland were used

-   ...

### Preparation of pollen data

-   Pollen were extracted from the moss using standardized pollen preparation protocol (REF)

-   Pollen were identified using light microscope and the key of (REF)

-   pollen correction factors- new data of Githumbi 2022,

    -   draws from RPP distribution
    -   no ppe data - removed in corrected pollen data
    -   X percent of data was removed because missing PPE
        -   Pteridophytes

        -   Populus in Switzerland

### Calculation of CWM

-   Bayesian model

    -   univariate instead multivariate

-   pollen cwm - gapfilled data from try

-   vegetation cwm - data collected in the field

    -   Try data was used for species traits that were not recorded in the field or when less than two observation was made

### Statistical analysis

#### Correction factor

-   draws from dnorm(RPP, RPP_sd), discarded draw when negative

## Results

### Effect of correction factors

### Effect of pollination mode and plant functional group

### Effect of taxonomic resolution

###

Orange = zone A (0-10m) Purple = zone B (10-100m) Greenish = zone C (100-1000m)

```{r,  fig.cap="CWM of vegetation and pollen-based CWM"}
source("04_Correlations.R")
```

## Appendix

### Pollen representation

```{r fig.height=6, fig.width=10, fig.cap="Scatter plot of vegetation basal area % against pollen abundance %"}
source("03_Vegetation_pollen_comparison.R") # pollen plot
pol_veg_plot
```

```{r fig.height=6, fig.width=10, fig.cap="Scatter plot of vegetation basal area % against adjusted pollen abundance %"}

source("03_Vegetation_pollen_comparison.R") # pollen plot
pol_veg_plot_adj

```

### Study design

![Design of Swiss trait sampling and vegetation surveys](Other/BDM_study_design.png)

![Design of Scottish vegetation survey in zone A](Other/Scotland_field_design.png)

### Trait values

#### gapfilled data

```{r}
source("02_Prepare_trait_data_gapfilled.R")
kable(sum_trait)
```

```{r}
source("01_Prepare_trait_data.R", encoding = 'UTF-8', echo=TRUE)
kable(sum_ph)
kable(sum_la)
kable(sum_sla)
```

## References

### 2022-07-27 01_Prepare_trait_data.R
add summary table for trait measurements

@@ -121,3 +121,27 @@ cn <- cn_raw %>%
   left_join(harm, by = "species") %>% 
   dplyr::select(-measurementID)
 saveRDS(cn, "RDS_files/01_CN.rds")
+
+## SUmmary table per pollentaxon
+spec <- readRDS("RDS_files/02_PollenType_species.rds")
+
+sum_ph <- plantheight %>% 
+  left_join(spec, by = c("species" = "stand.spec")) %>% 
+  group_by(pollentaxon) %>% 
+  summarise(nspec = length(unique(stand.spec)),
+            nobs = n(),
+            PlantHeight = paste(round(mean(PlantHeight), 1), "+-", round(sd(PlantHeight), 1)))
+sum_la <- la %>% 
+  left_join(spec, by = c("species" = "stand.spec")) %>% 
+  group_by(pollentaxon) %>% 
+  summarise(nspec = length(unique(stand.spec)),
+            nobs = n(),
+            LA = paste(round(mean(LA), 1), "+-", round(sd(LA), 1)))
+sum_sla <- sla %>% 
+  left_join(spec, by = c("species" = "stand.spec")) %>% 
+  group_by(pollentaxon) %>% 
+  summarise(nspec = length(unique(stand.spec)),
+            nobs = n(),
+            PlantHeight = paste(round(mean(SLA), 1), "+-", round(sd(SLA), 1)))
+
+rm(list=setdiff(ls(), c("sum_ph", "sum_la", "sum_sla")))

### 2022-07-27 02_Prepare_trait_data_gapfilled.R
fix paste issue

@@ -110,9 +110,10 @@ sum_trait <- dfTRAIT %>%
   group_by(pollentaxon) %>% 
   summarise(nspec = length(unique(stand.spec)),
             nobs = n(),
-            SLA = paste(round(mean(SLA), 1), "±", round(sd(SLA), 2)),
-            LA = paste(round(mean(LA), 1), "±", round(sd(LA), 1)),
-            `Plant height` = paste(round(mean(PlantHeight), 1), "±", round(sd(PlantHeight), 1)))
+            SLA = paste(round(mean(SLA), 1), "+-", round(sd(SLA), 2)),
+            LA = paste(round(mean(LA), 1), "+-", round(sd(LA), 1)),
+            `Plant height` = paste(round(mean(PlantHeight), 1), "+-", round(sd(PlantHeight), 1))) %>% 
+  drop_na()
 
 saveRDS(dfTRAIT, "RDS_files/02_Gapfilled_traits.rds")
 

### 2022-07-27 03_CWM_estimates_vegetation.R
Tidy script

@@ -27,6 +27,7 @@ if(!require(Hmisc)  )install.packages("Hmisc")
 if(!require(furrr))  install.packages("furrr")
 if(!require(LCVP)) install.packages("LCVP")
 if(!require(lcvplants)) install.packages("lcvplants")
+if(!require(FD)) install.packages("FD")
 
 set.seed(1)
 
@@ -34,9 +35,8 @@ set.seed(1)
 # Trait data ----
 # Scotland ----
 plantheight <- readRDS("RDS_files/01_PlantHeight.rds") %>% 
-  mutate(sitename = as.factor(sitename), species = as.factor(species),
-         PlantHeight = PlantHeight) %>% # plantheight in centimeter
-  ungroup()
+  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
+  ungroup() 
 ldmc <- readRDS("RDS_files/01_LDMC.rds") %>% 
   mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
   ungroup()
@@ -99,16 +99,26 @@ zoneC <- dfABUN_bc %>%
             by = c("stand.spec" = "AccSpeciesName")) %>% 
   ungroup %>% 
   dplyr::select(sitename, stand.spec, genus, family = fam, abun = spec_abun_c, growthform, polmode)
+lABUN <- list("zoneA" = zoneA, "zoneB" = zoneB, "zoneC" = zoneC)
 
 selectedtrait <- "PlantHeight"
-abun <- zoneB
-selectedpolmode <- unique(dfPFT$polmode)
+selectedabun <- "zoneB"
+selectedpolmode <- unique(dfPFT$polmode)[2]
 selectedpft <- unique(dfPFT$growthform)
 selectedtaxres <- "stand.spec"
 
-# Calculate CWM Scotland----
-cwm_func <- function(abun, selectedtrait,selectedtaxres, selectedpolmode, selectedpft){
-  taxa <- abun %>% 
+trsh <- c("tree", "shrub")
+herb <- c("herb", "grass", "fern")
+allpft  <- c("tree","herb","grass","fern","shrub", NA)
+wind <- "wind"
+nowind <- "not wind"
+allpol <- c("wind", "not wind", NA)
+
+# Scotland----
+if(0){ 
+cwm_func <- function(selectedabun, selectedtrait,selectedtaxres, selectedpolmode, selectedpft){
+  taxa <- lABUN %>% 
+    pluck(selectedabun) %>% 
[…]

### 2022-07-27 03_CWM_estimates_pollen.R
tidy result writing

@@ -227,7 +227,15 @@ cwmlogfunc <- function(selectedtrait, selectedcountry, selectedabun,
   
   # summarize results
   res <- data.frame(summary(results),
-                   sitename = sitenames)
+                    sitename = sitenames, 
+                    trait = selectedtrait,
+                    country = selectedcountry,
+                    growthform = case_when(all(selectedpft == trsh) ~ "trsh",
+                                           all(selectedpft == herb) ~ "herb",
+                                           TRUE ~ "allpft"),
+                    pollination = case_when(all(selectedpolmode == wind) ~ "wind",
+                                            all(selectedpolmode == nowind) ~ "no wind",
+                                            TRUE ~ "allpol"))
   saveRDS(res, paste0("RDS_files/03_CWM_estimates_pollen_", 
                       selectedcountry,"_", selectedtrait, "_", 
                       deparse(substitute(selectedpft)), ".rds"))
@@ -338,7 +346,11 @@ cwmlogfunc <- function(selectedtrait, selectedcountry, selectedabun,
   
   # summarize results
   res <- data.frame(summary(results),
-                    sitename = sitenames)
+                    sitename = sitenames, 
+                    trait = selectedtrait,
+                    country = selectedcountry,
+                    pollination = selectedpolmode,
+                    growthform = "allpft")
   saveRDS(res, paste0("RDS_files/03_CWM_estimates_pollen_", 
                       selectedcountry,"_", selectedtrait, "_", 
                       deparse(substitute(selectedpolmode)), ".rds"))

### 2022-07-27 04_Correlation_bayesian.R
write multivariate jags model

@@ -72,64 +72,56 @@ dfCWM <- files %>%
                             zone == "pollen" ~ "pollentaxon",
                             TRUE ~ "stand.spec"),
          country = str_extract(label, pattern = "Scotland|Switzerland")
-  )
+  ) 
 
 
 Trait <- dfCWM %>%
-  filter(trait == "PlantHeight",
-         growthform == "all",
-         pollination == "all",
-         taxres %in% c("pollentaxon", "species")) %>%
-  filter(zone %in% c("pollen","zoneA"),
-         country == "Scotland") %>%
-  dplyr::select(zone, id, Mean) %>%
-  pivot_wider(id_cols = c("id", "zone"), names_from = zone, values_from = Mean) 
+  filter(label %in% c("pollen_Scotland_PlantHeight_percent",
+                      "zoneB_Scotland_PlantHeight_stand.spec_windNAnot wind_treeshrubNAherbgrassfern")) %>%
+  dplyr::select(label, id, Mean, SD) %>%
+  pivot_wider(names_from = label, values_from = c(Mean,SD)) %>% 
+  rename(pollen_mean = 2, zoneA_mean =3, pollen_sd=4, zoneA_sd=5)
 
-plot(Trait)
-N <- nrow(Trait)
-
-set.seed(31415)
-
-mu <- c(10, 30)
-sigma <- c(20, 40)
-rho <- -0.7
-cov_mat <- rbind(c(sigma[1]^2, sigma[1]*sigma[2]*rho),
-                 c(sigma[1]*sigma[2]*rho, sigma[2]^2))
-N <- 30
-Trait <- rmvnorm(N, mu, cov_mat)
-plot(x, xlim=c(-125, 125), ylim=c(-100, 150))
-
+CWM_mean <- Trait %>% 
+  dplyr::select(pollen_mean, zoneA_mean) %>% 
+  as.matrix()
+CWM_sd <- Trait %>% 
+  dplyr::select(pollen_sd, zoneA_sd) %>% 
+  as.matrix()
+N <- nrow(CWM_mean)
 
 # https://www.sumsar.net/blog/2013/08/bayesian-estimation-of-correlation/
 model_string <- "
   model {
     for(i in 1:N) {
-      Trait[i,1:2] ~ dmnorm(mu[], prec[ , ])
+      CWM_mean[i,1:2] ~ dmnorm(mu[], prec[,,i])
     }
 
[…]

### 2022-08-16 02_PollenType_Species_GBIF.R
replace occ_search by occ_download

@@ -17,7 +17,7 @@
 ##
 ## Notes:
 ## 
-## - evaluate limit setting occ_search
+## 
 ##   
 ## ---------------------------
 
@@ -110,58 +110,35 @@ taxonkeys <-  bind_rows(spkeys, gnkeys, famkeys, classkeys) %>%
 countries <- c("United Kingdom", "Switzerland")
 countrykeys <- isocodes %>% filter(name %in% countries) %>% pull(code)
 
-## create function  that downloads, cleans, saves and creates conversion table for pollen type to species
-if(0){
-gbifsaveclean <- function(taxonkey, countrykey){
-  # perform gbif search
-  occsearch <- occ_search(taxonKey = taxonkey,
-                        country = countrykey,
-                        limit = 1000,  # evaluate limit setting
-                        year = "1990, 2020",
-                        hasCoordinate = T)     
-  
-  # proceed to next steps of the function (cleaning) if the search retrieved data  
-  if(!is.null(occsearch$data)) {
-  # save raw data
-  #saveRDS(occsearch, paste("RDS_files/RawGBIF/GBIFoccurances_", countrykey, taxonkey, ".rds", sep = ""))
-  
-  # only select data element, discard meta
-  occdat <- occsearch$data 
-  
-  # clear empty elements (key didn't retrieve data)
-  occdat <- occdat %>% compact()
-  
-  # clean occurrences
-  occdat <- occdat %>% #purrr::map(. 
-    data.frame() %>%
-      cc_val(lon = "decimalLongitude", lat = "decimalLatitude") %>% # removes records with empty coordinates
-      cc_equ(lon = "decimalLongitude", lat = "decimalLatitude") %>% # removes records with equal latitude and longitude coordinates
-      cc_cap(lon = "decimalLongitude", lat = "decimalLatitude") %>% # removes records within a certain radius around country capitals
-      cc_cen(lon = "decimalLongitude", lat = "decimalLatitude") %>% # removes records within a radius around the geographic centroids of political countries and provinces.
-      cc_inst(lon = "decimalLongitude", lat = "decimalLatitude") %>% # removes records assigned to the location of zoos, botanical gardens, herbaria, universities and museums
[…]

### 2022-08-17 01_PFT_table_pollen.R
add script

@@ -1,0 +1,73 @@
+## ---------------------------
+##
+## Script name: Create_PFT_table
+##
+## Purpose of script: Create pollination mode and pft table for pollen data
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-07-14
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   - pollination mode comes from Reitalu (2019) J of Ecology
+##
+## ---------------------------
+
+
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+if(!require(readxl)) install.packages("readxl")
+
+## Load data
+dfPOL_Scot <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
+dfPOL_Swiss <- readRDS("RDS_files/01_Pollen_data_Swiss.rds")
+dfPOL <- bind_rows(dfPOL_Scot, dfPOL_Swiss)
+polmode <- read_xlsx("Data/Pollination_modes_Reitalu_2019.xlsx")
+
+pollentaxa <- dfPOL %>% 
+  pull(pollentaxon) %>% 
+  unique()
+
+# Harmonize nomenclature between Reitalu study and current
+polmode <- polmode %>% 
+  mutate(pollentaxon_reit = pollentaxon,
+         pollentaxon = dplyr::recode(pollentaxon,  
+                                     "Apiaceae undiff." = "Apiaceae",
+                                     "Caryophyllaceae undiff." = "Caryophyllaceae",
+                                     "Cyperaceae undiff." = "Cyperaceae",
+                                     "Ericacea-type" = "Ericales (tetrad)",
+                                     "Fraxinus excelsior" = "Fraxinus",
+                                     "Geranium" = "Geraniaceae",
+                                     "Pedicularis" = "Pedicularis palustris",
+                                     "Ranunculuceae undiff." = "Ranunculaceae",
+                                     "Rosaceae undiff." = "Rosaceae",
+                                     "Viola palustris-type" = "Viola",
+                                     "Corylus/Myrica" = "Corylus",
+                                     "Drosera rotundifolia-type" = "Drosera",
+                                     "Epilobium-type" = "Epilobium",
+                                     "Fabaceae undiff." = "Fabaceae",
+                                     "Juglans regia" = "Juglans",
+                                     "Populus tremula" = "Populus",
[…]

### 2022-08-17 01_Prepare_TRY_data.R

@@ -1,0 +1,330 @@
+## ---------------------------
+##
+## Script name: 01_Prepare_TRY_data
+##
+## Purpose of script: 
+## - Prepare downloaded TRY data for further analysis
+## - for trait data to reconstruct trait composition from pollen data
+## - for pollination mode and plant functional type to classify vegetation data
+## 
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-08-15
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+## - based on a script by Franziska Schrodt   
+##
+## ---------------------------
+
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+if(!require(data.table)) install.packages("data.table")
+
+## Traits ----
+trait_raw <-
+  fread(
+    "Data/22274_15082022203634/22274.txt",
+    sep = "\t",
+    data.table = FALSE,
+    stringsAsFactors = FALSE,
+    strip.white = TRUE,
+    drop = c("LastName", "FirstName", "Dataset", "Comment", 
+             "V28", "Reference", "ValueKindName", "Replicates"))
+
+#removing trait outliers based on error risk
+trait <- trait_raw %>%
+  dplyr::select(DatasetID,
+          DataID,
+          ObsDataID,
+          ObservationID,
+          AccSpeciesID,
+          AccSpeciesName,
+          TraitID,
+          OriglName,
+          TraitName,
+          OrigObsDataID,
+          OrigValueStr,
+          OrigUnitStr,
+          StdValue,
+          UnitName,
+          ErrorRisk
+          ) %>%
+  mutate(ErrorRisk2 = ifelse(is.na(ErrorRisk), 0, ErrorRisk)) %>%
+  filter(ErrorRisk2 < 8) 
+
+# Get meta data
+meta <- trait %>% 
+  filter(is.na(TraitID))
+# Filter out meta data
+trait <- trait %>% 
+  filter(!is.na(TraitID))
+
+#generate list of units for ALL TRY traits
+units <- trait %>%
+  select(OriglName,
+         OrigUnitStr, 
+         TraitName, 
+         UnitName) %>%
+  unique()
+
+# Subsetting out traits and naming them
+# if origlname or unit is blank but there are no duplicates we are keeping it. 
+trait2 <- trait %>%
+  # remove problem data where there are replicates for ObservationID
[…]

### 2022-08-17 01_Species_abundance.Rmd
fix standardization of species name

@@
 ```{r standardize species names}
+# load in generalized species list, spelling checked and harmonized with
+# LCVP
 harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx") 
+harm <- harm %>% 
+  mutate(stand.spec = str_remove(stand.spec, pattern = " sp."))
 spec_a <- spec_a %>% 

### 2022-08-17 01_Swiss_data_compilation.R
fix standardization of species names

@@ -59,7 +59,8 @@ bdm_sp_codes <- bdm_sp_codes %>%
                                alcxan = "alcvul", # Alchemilla vulgaris current code alcxan is not in species or trait data, alcvul is
                                soldal = "solalp" # Soldanella alpinia idem
                                ) # soraur,lonneg, rubfru, giumon missing from codes 
-         )
+         ) 
+
 ## Abundance ----
 # Join species codes with abundance data
 bdm_sp <- bdm_sp %>% 
@@ -77,7 +78,9 @@ spec <- bdm_sp %>%
   filter(!is.na(Species)) %>% 
   pull(binom) %>% 
   unique()
+
 lcvp <- lcvp_search(spec) # find synonyms
+
 synonyms <- lcvp %>% 
   dplyr::select(binom = Search, stand_spec = Output.Taxon) %>% 
   mutate(stand_spec = word(stand_spec, 1, 2)) # drop authorship
@@ -88,6 +91,8 @@ regex_pattern <-
 bdm_sp <- bdm_sp %>% 
   # replace by standardized name
   mutate(stand.spec = str_replace_all(binom, regex_pattern)) %>% 
+  # remove sp. epiphet
+  mutate(stand.spec = str_remove(stand.spec, pattern = " sp.")) %>% 
   # select relevant columns, rename
   dplyr::select(sitename = plot, subplot, species_code, family = Family, genus = Genus, 
                 stand.spec, PFT = Functional_group, abun) %>% 
@@ -124,16 +129,21 @@ bdm_traits <- bdm_traits %>%
   filter(plot %in% plot_id)
 bdm_traits <- bdm_traits %>% 
   # replace by standardized name
-  mutate(binom = str_replace_all(binom, regex_pattern)) %>% 
+  mutate(stand.spec = str_replace_all(binom, regex_pattern)) %>% 
   # code unidentified and no plant categories
-  mutate(binom = case_when(!is.na(binom) ~ binom,
-                           str_detect(species_code, pattern = "[:digit:]") ~ "Unindentified",
-                           species_code %in% c("NA", "lonneg","rubfru","giumon","soraur") ~ "Unindentified" # missing species code
-                           )
-         ) %>% 
+  mutate(stand.spec = case_when(!is.na(stand.spec) ~ stand.spec,
+                                species_code %in% c("FO","GR", "CR") ~ "Unindentified",
+                                species_code %in% c("litter", "moss", "deadwood",
+                                                    "dead wood", "dead trunk",
[…]

### 2022-08-17 Manuscript.Rmd
Add text intro and results

@@
 source("Write_bib_file.R") # Compile library for reference list
+source("01_Prepare_pollen_data.R")
 
@@
 
-# Testing the pollen - functional diversity relationship
+(working title)
 
--   pre-print?
+# Reconstructing plant trait composition from palaeoecological data: testing the validity of the use of pollen data for functional trait reconstruction
 
@@
 
--say something about shifts in plant composition in the Anthropocene and ecosystem processes- In the past two decades, plant functional traits have become an important tool in ecology and biogeography for the study of consequences of changing plant composition on ecosystem functioning and the effect of environment factors on plant composition [@Hevia_2017].
+In the past two decades, plant functional traits have become an important tool in ecology and biogeography for the study of consequences of changing plant composition on ecosystem functioning and the effect of environment factors on plant composition [@Hevia_2017].
 Recently, considerable interest in the trait-based approach has arisen in palaeoecology as well [@Brussel_2018; @Blaus_2020; @Carvalho_2019; @Reitalu_2015; @Sande_2019].
-Applying the trait-based approach could offer a new way of interpreting pollen data and could increase our understanding of the drivers of vegetation change on timescales of centuries to millennial.
-For instance, @Sande_2019 combined functional trait and pollen data to study the effect of climate and human disturbances on plant trait composition in the Amazon on a 7000-year time scale.
+Applying the trait-based approach could offer a new way of interpreting pollen data and increase our understanding of the drivers of vegetation change on timescales of centuries to millennial.
+For instance, @Sande_2019 combined functional trait and pollen data to study the effect of climate and human disturbances on vegetation composition in the Amazon on a 7000-year time scale.
 
 Although palaeoecological studies that have applied the trait-based approach show promising results, it is a relatively new approach of which the intricacies are underexplored.
+Little is known about how representative the pollen record is of plant trait composition.
+[-include what others have found-.]{.underline} To improve our understanding of the relationship between the trait composition of the vegetation and trait compositions reconstructed from pollen records, we here use surface pollen samples.
+Surface pollen assemblages approximate vegetation composition in recent years and are commonly used in palaeoecology to verify pollen-vegetation relationships [@Birks_2016].
+Here, we use surface samples of 16 Scottish and 26 Swiss sites to test the relationship between reconstructed trait composition from pollen data and vegetation trait composition.
+
 Pollen quantity does not directly represent past plant abundance, taxa can become over- or underrepresented in the pollen record because of differences in pollen production and dispersal.
--something about pollination syndrome and plant function type - To improve our understanding of the relationship between the trait composition of the vegetation and trait compositions reconstructed from pollen records, we here use surface pollen samples.
-Surface pollen assemblages approximate vegetation composition in recent years and are commonly used in palaeoecology to verify pollen-vegetation relationships [@Birks_2016].
-Here, we use surface samples of 16 Scottish and 27 Swiss sites to test the relationship between pollen trait composition and vegetation trait composition.
+[-something about pollination syndrome and plant function type]{.underline} - To correct for these differences pollen productivity factors can be use, these correction factors down-weight the contribution of dominant taxa.
+Furthermore, comparing between functional groups based on growth form or pollination syndrome could give insights in to the representativeness of the pollen assemblage to the vegetation.
+[-something about why pollination syndrome, distance from sampling site]{.underline} [and plant function type would matter to representation -]{.underline}
 
-To correct for these differences pollen productivity factors can be use, these correction factors down-weight the contribution of dominant taxa.
-Furthermore, comparing between functional groups based on growth form or polination syndrome could give insights in to the representativeness of the pollen assemblage to the vegetation.
-
-Perhaps the most challenging in reliably reconstructing past trait is to attribute functional trait data to the taxa found in pollen records, which can rarely be resolved to the species level (Birks 2020).
-Trait variation within taxa can be substantial (REF).
-It can be expected that this especially problematic in the case of large, abundandt plant families that cannot easily be resolved to a lower taxonomic level, such as the Poaceae family.
+Perhaps the most challenging in reliably reconstructing past trait is to attribute functional trait data to the taxa found in pollen records, which can rarely be resolved to the species level [@Birks_2020].
+Trait variation within taxa can be substantial [@Siefert_2015].
+It can be expected that this especially problematic in the case of large, abundant plant families that cannot easily be resolved to a lower taxonomic level, such as the Poaceae family.
 Also small families where the trait variation is high could be problematic, such as is the case with the Fabaceae family where the height variation can be ...-fold (for instance Ulex shrub height and clover).
@@
 One way of dealing with the uncertainty caused by trait variation in pollen taxa is by using Bayesian models.
-Bayesian models allow you to do this and this.
+Using a Bayesian approach for modelling trait composition from pollen data allows for the propagation of uncertainty in the taxon level trait reconstruction to the community level.
 
-With this study, we aim to test the validity of using of pollen records for plant trait composition reconstructions.
-More specifically, we will test the effect of vegetation composition and trait choice on the representation of pollen assemblages for functional diversity reconstructions.
+With this study, we aim to test the validity of using of pollen records for plant trait composition reconstruction.
+More specifically, we will test the effect of vegetation composition, taxonomic resolution and trait choice on the representativeness of pollen assemblages for functional diversity reconstructions.
 We examine the effectivity of pollen production factors to improve the pollen - functional diversity relationship.
 Lastly, we test the representation of plant functional groups based on growth form and pollination syndrome in the functional composition reconstruction.
-
--   resolution of the pollen record - calculate CWM using different taxonomic levels of the vegetation?
-
--   trait choice
-
--   assignment of species to pollen taxa
-
--   representation of pft's - pollination syndrome
 
@@
 Pollen data, plant species composition data and trait data were collected in 16 sampling locations in June 2021 in the Cairngorms National Park, Scotland, United Kingdom.
-Study design mostly followed the protocol of [@Bunting_2013].
+Study design mostly followed the protocol of @Bunting_2013 (Appendix).
 Sampling locations were at least 1 km from each other to avoid pollen source areas from overlapping with each other.
 Sites were selected to represent four vegetation types in the area: grassland, heath land, deciduous forest and coniferous forest.
-Every sampling location consisted of three nested concentric rings of 3m, 100m and 1000m in radius.
-The inner two rings were surveyed in the field and for the outer ring vegetation cover was estimated using QGIS and the Scottish Habitat and Land cover map of 2020 .
-To determine the composition of the modern pollen assemblage at every site, surface samples of moss were collected at the center of each sampling location.
-Moss was collected from multiple location within a 1m radius from the center.
+Every sampling location consisted of three nested concentric rings of 3.4m, 100m and 1000m in radius.
+The inner two rings were surveyed in the field and for the outer ring vegetation cover was estimated using QGIS and the Scottish Habitat and Land cover map of 2020.
+To determine the composition of the modern pollen assemblage at every site, surface samples of moss were collected at the centre of each sampling location.
+Moss was collected from multiple location within a 1m radius from the centre.
 Only living tissue was sampled.
 
-An addition 27 pollen samples were collected from sites in the Swiss Biodiversity Monitoring network, that were selected based on an elevation gradient (?).
-Only the inner ring was surveyed in these sites
+An additional 26 pollen samples were collected from sites in the Swiss Biodiversity Monitoring network in summer 2020, that were selected based on an elevation gradient (?).
+Only the inner ring (0-3.4m) was surveyed in these sites.
+[(Check with Tom the exact method in Switzerland)]{.underline}
 
@@
 
-In the inner ring (0-10m) 17 plots of 0.5 m were surveyed: one in the center, 3 in every wind direction between 3, 6 and 10 meters, and one in the NE, NW, SE and SW direction (See design in Appendix).
-In the middle ring (10-100m) between 4 and 6 plots of 10 by 10?
-meter were randomly selected.
+In the Scottish sites, the inner ring (0-3.4m) 17 plots of 0.5 m were surveyed: one in the center, 3 in every wind direction between 3, 6 and 10 meters, and one in the NE, NW, SE and SW direction (See design in Appendix).
+In the middle ring (10-100m) between 4 and 6 plots of 10 by 10 meter were selected on a stratified random design based on vegetation types that were identified in the fields.
 Vegetation abundance was estimated using percentage cover.
+All species names were standardized according to the Leipzig Plant Catalogue using the *lcvplants* and *lcvp* R packages [@freiberg2020lcvp].
+(Check with Tom the exact method in Switzerland)
 
-### Trait sampling
+### Trait collection
 
--   Plant height, SLA and LA measured in the field
+Trait data was collected in the middle ring.
+For determining plant height 10 flowering of individuals were selected at every site if present, non-flowering individuals were selected if flowering individuals were not available.
+Five leaf samples per species for specific leaf area (SLA), leaf area (LA), leaf dry matter content (LDMC), leaf carbon and leaf nitrogen were collected in sealable plastic bags.
+Fresh weight and surface area were measured within 72 hours of collection using a carat scale and flatbad scanner.
+Samples were dried at 70 C for 72 hours before measuring dry weight.
+For measuring leaf carbon and nitrogen samples using ICP-MS were crushed and X.
+
+For reconstructing trait composition from pollen data, previously gapfilled data from the TRY database *(will replace this by the raw TRY data as discussed)* was used [@kattge2020try].
+
+(Check with Tom the exact method in Switzerland)
 
@@
 
--   Species list of Switzerland and Scotland were used
-
--   ...
+For reconstructing plant trait composition from pollen data, assumptions need to be made about the species belonging to the pollen taxa in the data set.
+We used the *rgbif* package to download the current distribution of species belonging to each pollen taxon in Scotland and Switzerland from the Global Biodiversity Information Facility database (REF).
+Species distribution data was queried using the pollen taxonomic name.
 
@@
 
--   pollen correction factors- new data of Githumbi 2022,
+    -   pollen correction factors- new data of Githumbi 2022 [@Githumbi_2022],
 
-    -   draws from RPP distribution
-    -   no ppe data - removed in corrected pollen data
-    -   X percent of data was removed because missing PPE
-        -   Pteridophytes
+    let to, on average, exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish
+    pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%) and `round(mean(missing_ppe_swiss$percent_missing),2)*100`% of the Swiss pollen data (max `r round(max(missing_ppe_swiss$percent_missing),2)*100`%).
 
-        -   Populus in Switzerland
+    ### 
+    Calculation of community weighted mean
 
-### Calculation of CWM
+    To test the representativeness of pollen for plant trait composition reconstructions, the community weighted mean trait values was calculated for the pollen data and for the vegetation data in the three nested circles: zone A (0-3.4m) zone B (10-100 m) and zone C (100 - 100 m).
+The trait distribution on the species or the pollen taxonomic level was estimated assuming that the trait data follows a lognormal distribution for the
+    likelihood, so that$$trait_{i}^{observed} \sim { lognormal(\mu_{j},\sigma_j)}$$Where *i* represents each trait observation of a taxon (*j*) from the trait data.
+$\mu_{j}$ and $\sigma_{j}$ are the mean trait value and the standard deviation per taxon, respectively.
+Vague priors were used for the taxon mean and standard deviation estimation.
+We modeled the community weighted mean (CWM) as a mixture of normal distributions that were weighted by corrected pollen percentages, so that
 
--   Bayesian model
+    $$
+    CWM_{s} \sim \sum^{n}\pi_{j}Normal(\mu_{j},\sigma_{j})
+    $$
 
-    -   univariate instead multivariate
+    Where *n* is the total number of taxa, and $\pi_{j}$ the abundance of each pollen taxon/species at a given site.
+The *CWM~­~^mean^* and *CWM^sd^* were subsequently used in the estimation of the correlation between the trait composition of the vegetation and the pollen data.
+We used a multivariate normal distribution to estimate correlation coefficient *ρ* so that,
+    $$[CWM_{s}^{veg},CWM_{s}^{veg}] \sim MultivariateNormal([\mu_{veg},\mu_{veg}],\sum) $$
 
--   pollen cwm - gapfilled data from try
+    $$
+    \sum \sim \begin{pmatrix} \sigma_{veg}^{2} & \rho\sigma_{veg}\sigma_{pol} \\ \rho\sigma_{veg}\sigma_{pol} & \sigma_{pol}^{2} \end{pmatrix}
+    $$
 
--   vegetation cwm - data collected in the field
+    $$
+    \rho \sim Uniform(-1,1)
+    $$
 
-    -   Try data was used for species traits that were not recorded in the field or when less than two observation was made
-
-### Statistical analysis
-
-#### Correction factor
-
--   draws from dnorm(RPP, RPP_sd), discarded draw when negative
+    Correlations were calculated for all zones and three traits (currently): SLA, LA and plant
+    height.
+Furthermore, four different \"treatments\" were assigned to test what factors could influence representation of trait composition by pollen data: the use of correction factors, the separation of vegetation and pollen data in to plant functional type and pollination mode.
+Lastly, the vegetation was aggregated to genus level and family level to test the effect of taxonomic resolution on the reconstructions.
 
@@
 
-### Effect of pollination mode and plant functional group
+![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Pollen_uncorrected.png)
+
+![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Pollen_corrected.png)
+
+### Pollination mode
+
+![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Wind_pollinated.png)
+
+![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/NotWind_pollinated.png)
+
+### Plant functional group
+
+![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/TRSH.png)
+
+![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/HERB.png)
 
@@
 
-###
+![](Figures/veg_genus.png)
 
-Orange = zone A (0-10m) Purple = zone B (10-100m) Greenish = zone C (100-1000m)
-
-```{r,  fig.cap="CWM of vegetation and pollen-based CWM"}
-source("04_Correlations.R")
-```
-
-## Appendix
+## ![](Figures/veg-Family.png)Appendix
 
@@
 
-#### gapfilled data
-
 ```{r}

### 2022-08-19 04_Correlation_bayesian.R
write code for to run code for different treatments

@@ -24,104 +24,473 @@ if(!require(rjags)) install.packages("rjags")
 if(!require(runjags)) install.packages("runjags")
 if(!require(mvtnorm)) install.packages("mvtnorm") #to generate correlated data with rmvnorm.
 if(!require(car)) install.packages("car") #To plot the estimated bivariate normal distribution.
+library(gridExtra)
 
 # collate data
 traits <- c("LA","SLA","PlantHeight")
-# all taxa
-files <- 
-  list.files("RDS_files/") %>% 
-  str_subset(pattern = "03_CWM_estimates_") %>% 
-  str_subset(pattern = "adjusted", negate = TRUE) %>% 
-  str_subset(paste(traits, collapse = "|"))
 
-lab <- files %>% 
+pollen_files <- list.files("RDS_files/") %>% 
+  str_subset(pattern = "03_CWM_estimates_") %>% 
   str_subset(paste(traits, collapse = "|")) %>% 
-  str_subset(., "multivariate", negate = TRUE) %>% 
+  str_subset(pattern = "pollen") 
+  
+pollen_lab <- pollen_files %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
+  str_remove(., "03_CWM_estimates_") %>% 
+  str_remove(., ".rds")
+
+veg_files <- list.files("RDS_files/") %>% 
+  str_subset(pattern = "03_CWM_estimates_") %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
+  str_subset(pattern = "pollen", negate = TRUE)
+veg_lab <- veg_files %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
   str_remove(., "03_CWM_estimates_") %>% 
   str_remove(., ".rds")
 
 folderpath.fun <- function(x)
-{paste("RDS_files/", x, sep = "/")}
+  {paste("RDS_files/", x, sep = "/")}
 
-dfCWM <- files %>% 
+dfCWM_pol <- pollen_files %>% 
   folderpath.fun(.) %>% 
-  purrr::map2(lab, ~readRDS(.) %>% 
+  purrr::map2(pollen_lab, ~readRDS(.) %>% 
                 as.data.frame() %>% 
                 rownames_to_column("parameter") %>% 
-                filter(., str_detect(parameter, "cwm\\[")) %>% 
-                dplyr::select(Mean, SD) %>%  
+                filter(., str_detect(parameter, "cwm\\[")) %>%
+               
[…]

### 2022-08-19 01_Prepare_TRY_data.R
check missing species

@@ -23,11 +23,16 @@
 ## Load packages
 if(!require(tidyverse)) install.packages("tidyverse")
 if(!require(data.table)) install.packages("data.table")
+if(!require(LCVP)) install.packages("LCVP")
+if(!require(lcvplants)) install.packages("lcvplants")
+
+
+memory.limit(999999)
 
 ## Traits ----
 trait_raw <-
   fread(
-    "Data/22274_15082022203634/22274.txt",
+    "Data/TRY-request/22284_16082022183406/22284.txt",
     sep = "\t",
     data.table = FALSE,
     stringsAsFactors = FALSE,
@@ -209,28 +214,89 @@ d453 <- trait3 %>% #this dataset has 3 obs per plant, but no way to link leaves 
 trait4 <- trait3 %>%
   filter(!DatasetID == 453) %>%
   bind_rows(d453) %>% 
-  dplyr::select(DatasetID, OrigObsDataID, ObservationID, AccSpeciesName, CleanTraitName, StdValue) %>% 
-  pivot_wider(names_from = CleanTraitName, values_from = StdValue) %>% 
+  dplyr::select(ObsDataID, DatasetID, OrigObsDataID, ObservationID, AccSpeciesName, 
+                CleanTraitName, StdValue) %>% 
+  pivot_wider(id_cols = c(ObsDataID, DatasetID, OrigObsDataID,ObservationID, 
+                          AccSpeciesName),
+              names_from = CleanTraitName, values_from = StdValue) %>% 
   # remove duplicated values
   filter(!(duplicated(.[4:10]) & !is.na(OrigObsDataID)))
 
-saveRDS(trait4, "RDS_files/01_TRY_raw_traits.rds")
+# Save clean TRY data
+saveRDS(trait4, "RDS_files/01_Clean_TRY_data.R")
+# trait4 <- readRDS("RDS_files/01_Cleane_TRY_data.R")
 
-# Pollination and PFT ----
+# add pollen to species translation table ----
+spec <- readRDS("RDS_files/02_PollenType_species.rds") %>% 
+  distinct()
+
+# check for missing species in the trait data
+trait_sp <- trait4$AccSpeciesName %>% unique
+pol_sp <- spec$stand.spec %>% unique()
+
+misspec <-
+  pol_sp[!pol_sp %in% trait_sp] %>%
+  sort %>% unique
+
+# improve matching by standardizing species names
[…]

### 2022-08-19 01_Species_abundance.Rmd
fix pattern matching issue

--- 2022-07-24 01_Species_abundance.Rmd
@@
 ```{r standardize species names}
+# load in generalized species list, spelling checked and harmonized with
+# LCVP
 harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx") 
+harm <- harm %>% 
+  mutate(stand.spec = str_remove(stand.spec, pattern = " sp\\."))
 spec_a <- spec_a %>% 

### 2022-08-19 01_Prepare_trait_data_vegetation.R

@@ -24,7 +24,8 @@ plantheight_raw <- read_xlsx("Data/Plant_height.xlsx")
 sla_ldmc_raw <- read_xlsx("Data/LDMC_SLA_dataset.xlsx")
 cn_raw <- read_xlsx("Data/CN_dataset_150.xlsx")
 harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx") %>% 
-  rename(family = fam)
+  rename(family = fam) %>% 
+  mutate(stand.spec = str_remove(stand.spec, " sp."))
 
 ## Plant height ----
 # make plant height data tidy

### 2022-08-19 01_Swiss_data_compilation.R
fix pattern matching issue

@@ -69,7 +69,7 @@ bdm_sp <- bdm_sp %>%
   # remove 0 occurrences
   filter(!abun == 0) %>% 
   # join species codes
-  left_join(bdm_sp_codes, by = c("species_code"= "Species_code")) %>% 
+  left_join(bdm_sp_codes, by = c("species_code" = "Species_code")) %>% 
   # filter for relevant plots
   filter(plot %in% plot_id) 
 
@@ -92,7 +92,7 @@ bdm_sp <- bdm_sp %>%
   # replace by standardized name
   mutate(stand.spec = str_replace_all(binom, regex_pattern)) %>% 
   # remove sp. epiphet
-  mutate(stand.spec = str_remove(stand.spec, pattern = " sp.")) %>% 
+  mutate(stand.spec = str_remove(stand.spec, pattern = " sp\\.")) %>% 
   # select relevant columns, rename
   dplyr::select(sitename = plot, subplot, species_code, family = Family, genus = Genus, 
                 stand.spec, PFT = Functional_group, abun) %>% 

### 2022-08-19 02_PollenType_Species_GBIF.R
add missing species

@@ -27,8 +27,6 @@ if(!require(tidyverse)) install.packages("tidyverse")
 if(!require(rgbif)) install.packages("rgbif")
 if(!require(LCVP)) devtools::install_github("idiv-biodiversity/LCVP")
 if(!require(lcvplants)) devtools::install_github("idiv-biodiversity/lcvplants")
-if(!require(CoordinateCleaner)) install.packages("CoordinateCleaner")
-
 
 # Load in pollen data
 dfPOL_scot <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
@@ -121,13 +119,18 @@ gbifsaveclean <- function(taxonkey, countrykey){
     pred_in("basisOfRecord", "HUMAN_OBSERVATION"),
     pred_in("taxonKey", taxonkey),
     pred_in("country", countrykey),
-    pred_gte("year", 2012),
+    pred_gte("year", 2012), # observation from 2012 onwards
     pred("hasCoordinate", TRUE),
-    format = "SPECIES_LIST")
+    format = "SPECIES_LIST") # only download species list, not occurrence points
   
   occdata <- occ_download_wait(occsearch)
   
-  occdata <- occ_download_get(occsearch) %>% 
+  occdata <- occ_download_get(occsearch, path = "Data/GBIF-requests/",
+                              overwrite = TRUE) %>% 
+    occ_download_import()
+  # for already downloaded data
+  occdata <- occ_download_get(str_remove(list.files(path = "Data/GBIF-requests/")[1], ".zip"),
+                              overwrite = TRUE) %>% 
     occ_download_import()
   
   # create species per pollen taxon list
@@ -142,7 +145,8 @@ gbifsaveclean <- function(taxonkey, countrykey){
   
   # standardize species names according to the leipzig plant catalogue
   spec.stand <-
-    lcvp_search(spec$species,progress_bar = TRUE) 
+    lcvp_search(spec$species, progress_bar = TRUE) 
+  
   spec <-
     spec %>% mutate(stand.spec = word(spec.stand$Output.Taxon, 1,2) %>% # only keep binomial without authority 
                       str_remove("_x"),# x added for unknown reason,
@@ -186,24 +190,56 @@ duplicats <- specpol %>%
 specpol <- specpol %>% 
   # missing alternative name
   mutate(stand.spec = replace(.$stand.spec,
[…]

### 2022-09-08 01_PFT_table_vegetation.R
add missing species, create summary table

@@ -23,7 +23,7 @@ if(!require(data.table)) install.packages("data.table")
 
 # Pollination and PFT for vegetation----
 traits <- fread(
-  "Data/TRY-request/21989_20072022132411/21989.txt",
+  "Data/TRY-request/22386_31082022154852/22386.txt",
   sep = "\t",
   data.table = FALSE,
   stringsAsFactors = FALSE,
@@ -54,8 +54,8 @@ pft <- traits %>%
   # generalize pft values
   summarise(pft = paste(unique(OrigValueStr), collapse = ", ")) %>% 
   mutate(growthform = case_when(str_detect(pft, "tree|Tree") ~ 'tree',
-                                str_detect(pft, "shrub|Shrub") ~ 'shrub',
                                 str_detect(pft, "herb|Herb|Forb") ~ 'herb',
+                                str_detect(pft, "shrub|Shrub") ~ 'shrub',
                                 str_detect(pft, "grass|Grass|Graminoid|graminoid") ~ 'grass',
                                 str_detect(pft, "fern|Fern") ~ 'fern')) %>% 
   filter(!is.na(growthform)) %>% 
@@ -136,7 +136,11 @@ pft <- traits %>%
                              "Corylus avellana","Crataegus laevigata", 
                              "Dactylis glomerata", "Deschampsia flexuosa", 
                              "Juniperus communis", "Lolium perenne","Luzula sylvatica",
-                             "Orthilia secunda","Pteridium aquilinum", "Rubus caesius"),
+                             "Orthilia secunda","Pteridium aquilinum", "Rubus caesius",
+                             "Juncus bufonius", "Larix decidua", "Lathyrus palustris",
+                             "Galium odoratum","Galium verum", "Geranium robertianum",
+                             "Clematis vitalba", "Helianthemum salicifolium",
+                             "Rhinanthus glacialis", "Potentilla brevifolia"),
           growthform = c("grass","grass", "grass", "grass", "herb",
                          "grass","herb", "shrub", "herb", "herb",
                          "fern", "herb", "herb", "herb", "fern",
@@ -147,8 +151,15 @@ pft <- traits %>%
                          "herb", "grass", "grass", "grass",
                          "fern", "shrub", "shrub", "grass",
[…]

### 2022-09-08 01_Prepare_TRY_data.R
convert from m to cm

@@ -26,7 +26,7 @@ if(!require(data.table)) install.packages("data.table")
 if(!require(LCVP)) install.packages("LCVP")
 if(!require(lcvplants)) install.packages("lcvplants")
 
-
+# increase memory limit
 memory.limit(999999)
 
 ## Traits ----
@@ -220,11 +220,9 @@ trait4 <- trait3 %>%
                           AccSpeciesName),
               names_from = CleanTraitName, values_from = StdValue) %>% 
   # remove duplicated values
-  filter(!(duplicated(.[4:10]) & !is.na(OrigObsDataID)))
-
-# Save clean TRY data
-saveRDS(trait4, "RDS_files/01_Clean_TRY_data.R")
-# trait4 <- readRDS("RDS_files/01_Cleane_TRY_data.R")
+  filter(!(duplicated(.[4:10]) & !is.na(OrigObsDataID))) %>% 
+  # conver plantheight from m to cm
+  mutate(PlantHeight = PlantHeight * 100)
 
 # add pollen to species translation table ----
 spec <- readRDS("RDS_files/02_PollenType_species.rds") %>% 
@@ -243,10 +241,15 @@ misspec <-
 # saveRDS(lcvp, "RDS_files/01_TRY_species_standardization.rds")
 lcvp <- readRDS("RDS_files/01_TRY_species_standardization.rds")
 lcvp_stand <- lcvp %>% 
-  dplyr::select(Search, Output.Taxon) %>% 
-  mutate(stand.spec = word(Output.Taxon, 1,2))
+  dplyr::select(Search, Output.Taxon, family = Family) %>% 
+  mutate(stand.spec = word(Output.Taxon, 1,2),
+         genus = word(Output.Taxon, 1))
 trait5 <- trait4 %>% 
   left_join(lcvp_stand, by = c("AccSpeciesName" = "Search")) 
+
+# Save clean TRY data
+saveRDS(trait5, "RDS_files/01_Clean_TRY_data.rds")
+# trait5 <- readRDS("RDS_files/01_Clean_TRY_data.R")
 
 # check missing species again
 trait_sp <- trait5$stand.spec %>% unique

### 2022-09-08 01_Prepare_pollen_data.R
pollen id check

@@ -27,7 +27,7 @@ if(!require(readxl)) install.packages("readxl")
 set.seed(1)
 
 # Scotland ----
-pollen_raw <- read.csv("Data/Pollen_count.csv", skip = 1)
+pollen_raw <- read.csv("Data/Pollen_count_Scotland-checked.csv", skip = 1, sep = ";")
 ppe <- read_xlsx("Data/Githumbi_2022_RPP.xlsx")
 polmode <- readRDS("RDS_files/01_Pollination_mode.rds")
 
@@ -54,7 +54,8 @@ dfPOL <- pollen_raw %>%
            recode(Caryophyllaccea = "Caryophyllaceae",
                   Lilaceae = "Liliaceae",
                   "Juniperus communis" = "Juniperus")
-         )
+         ) %>% 
+  filter(!is.na(count))
 
 # percentage data with missing ppe's 
 missing_ppe_scot <- dfPOL %>%
@@ -112,7 +113,7 @@ saveRDS(dfPOL, "RDS_files/01_Pollen_data_Scot.rds")
 
 # Switzerland ----
 # read in pollen data
-dfPOL <- read.csv("Data/Pollen_count_Switserland.csv", skip = 1) %>% 
+dfPOL <- read.csv("Data/Pollen_count_Switserland-checked.csv", skip = 1, sep = ";") %>% 
   filter(!Key.index == "---") %>% # filter empty rows
   filter(!Sample.name == "---") %>% 
   pivot_longer(cols = c(starts_with("X"),"unknown"), # convert to long format

### 2022-09-08 01_Species_abundance.Rmd
add swiss data

@@
 library(readxl)
-library(Taxonstand)
+library(LCVP)
+library(lcvplants)
 ```
 
-## To do
+# To do
+
+# Scotland
 
@@
 spec_a_raw <- read_xlsx("Data/Species_zone_A.xlsx", skip = 1)
+canop_a_raw <- read_xlsx("Data/Canopy_zone_A.xlsx")
 
+canop_a <- canop_a_raw %>% 
+  pivot_longer(-sitename, names_to = "species", values_to = "abun") %>% 
+  mutate(direction = "Canopy", distance = "0-10 meter")
 # Make tidy
@@
                               plot == "Center" ~ "0 meter")) %>% 
+  bind_rows(canop_a) %>% 
   # create unique id for plot in site
@@
 ```
+
+# Switzerland
+
+## QGIS
+
+Land cover data was downloaded from:
+<https://www.bfs.admin.ch/bfs/en/home/services/geostat/swiss-federal-statistics-geodata/land-use-cover-suitability.html>
+
+To decrease file size, I only selected relevant columns: the one
+containing coordinates and land cover data using the following code:
+
+```{r eval = FALSE}
+# ag_b_00_03_37_area_csv <- read_delim(
+#     "C:/Users/lgxgv/OneDrive - The University of Nottingham/Data analysis/PhD/Validation-study/Data/Swiss_LC/ag-b-00.03-37-area-csv.csv",+delim = ";",
+#     escape_double = FALSE,
+#     trim_ws = TRUE
+#   )
+# 
+# LC <- ag_b_00_03_37_area_csv %>%
+#   dplyr::select(RELI, E_COORD = E, N_COORD = N, LC18_6, LC18_27)
+# 
+# write.csv(LC, file = "C:/Users/lgxgv/OneDrive - The University of Nottingham/Data
+#           analysis/PhD/Validation-study/Data/Swiss_LC/Swiss_LC18.csv")
+
+```
+
+In this file I found the CRS:
+<https://www.bfs.admin.ch/bfs/en/home/services/geostat/swiss-federal-statistics-geodata/land-use-cover-suitability/swiss-land-use-statistics.assetdetail.20104775.html>\
+"Les coordonnées « E_COORD » et « N_COORD » des hectares dans le système
+de coordonnées suisse MN95 (7 chiffres). Les coordonnées représentent le
+point d'échantillonnage correspondant (intersections des coordonnées
+hectométriques)."
+
+Import Swiss land cover (100m resolution) Swiss_LC18.csv
+
+-   Set CRS code to EPSG:4150 Rasterize this layer using the GDAL
+    rasterize (vector to raster) tool
+-   Set burnin to LC18_27
+-   Horizontal and vertical resolution to 100 m
+-   export and set resolution to 25 m for merge with forest cover data
+
+Import forest cover data in ArcGIS pro reproject to EPSG:4150 and export
+raster(folder called reproj)/ Import to QGIS set layer crs to EPSG:4150
+
+Import coordinates of pollen plots (Pollen_site_coord.csv).
+
+Create buffer zone of 100 m and 1000 m
+
+Use Zonal histogram to compute pixels in zone attributed to certain land
+cover type Export as Veg_cover_zoneC.csv and Veg_cover_zoneB.csv
+
+Import coordinate of all BDM sites Use Sample raster values to compute
+landcover types for vegetation plots. Export as Veg_type_plots_csv
+
+```{r}
+zone_b_raw <- read.csv("Data/Swiss_LC/FC_LC_cover_zoneB_swiss.csv")
+zone_c_raw <- read.csv("Data/Swiss_LC/FC_LC_cover_zoneC_swiss.csv")
+LC_labels <- read_xlsx("Data/Swiss_LC/be-b-00.03-37-area-var-v01.xlsx",
+                       sheet = "Codes LC_27")
+FC_labels <- read.csv("Data/Swiss_LC/Forest_cover_labels.csv") %>% 
+  rename(CODE = category)
+LC_labels <- LC_labels %>% 
+  dplyr::select(CODE, LABEL_DE) %>% 
+  bind_rows(FC_labels)
+
+zone_b <- zone_b_raw %>% 
+  # make tidy
+  dplyr::rename(sitename = aID_STAO,lon = X, lat = Y) %>% 
+  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
+  # rename veg_types
+  mutate(veg_type = str_remove(veg_type, "HISTO_") %>% as.numeric) %>% 
+  filter(!value == 0) %>% 
+  left_join(LC_labels, by = c("veg_type" = "CODE")) %>% 
+  # filter land cover type that are not vegation
+  filter(!LABEL_DE %in% c( "Lockergestein", "Wasser","Gebäude",
+                           "Versteinte Flächen","Befestigte Flächen",
+                           "Gemischte Kleinstrukturen")) %>%
+  # calculate veg_type percentage
+  group_by(sitename) %>% 
+  # calculate percentages
+  mutate(veg_cov = value/sum(value)) %>%
+  group_by(sitename, LABEL_DE) %>%
+  summarise(veg_cov = sum(veg_cov, na.rm = TRUE))
+
+zone_c <- zone_c_raw %>% 
+  # make tidy
+  dplyr::rename(sitename = aID_STAO,lon = X, lat = Y) %>% 
+  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
+  # rename veg_types
+  mutate(veg_type = str_remove(veg_type, "HISTO_") %>% as.numeric) %>% 
+  filter(!value == 0) %>% 
+  left_join(LC_labels, by = c("veg_type" = "CODE")) %>% 
+  # filter land cover type that are not vegation
+  filter(!LABEL_DE %in% c( "Lockergestein", "Wasser","Gebäude",
+                           "Versteinte Flächen","Befestigte Flächen",
+                           "Gemischte Kleinstrukturen")) %>%
+  # calculate veg_type percentage
+  group_by(sitename) %>% 
+  # calculate percentages
+  mutate(veg_cov = value/sum(value)) %>%
+  group_by(sitename, LABEL_DE) %>%
+  summarise(veg_cov = sum(veg_cov, na.rm = TRUE))
+
+# merge zone b and c
+veg_cover <- zone_b %>% 
+  full_join(zone_c, by = c("sitename", "LABEL_DE"), 
+            suffix = c("_b","_c")) %>% 
+  replace_na(list(veg_cov_b = 0))
+```
+
+```{r join veg type data and species data}
+# Veg type of BDM sites
+veg_type <- read.csv("Data/Swiss_LC/FC_LC_type_swiss.csv")
+
+# Join with labels
+veg_type <- veg_type %>% 
+  left_join(LC_labels, by = c("SAMPLE_1" = "CODE"))
+
+# Load in species data
+bdm_sp <- readRDS("RDS_files/01_Species_abundance_Swiss.rds")
+
+bdm_sp <- bdm_sp %>% 
+  # filter stuff that is not a plant
+  filter(!stand.spec == "No plants") %>% 
+  # filter unindentified
+  filter(!stand.spec == "Unindentified") %>% 
+  # join with vegetation type data
+  left_join(veg_type, by = c("sitename" = "aID_STAO")) 
+# create table with genus and family to join after summarizing
+genus_fam <- bdm_sp %>% 
+  ungroup() %>% 
+  dplyr::select(family, genus, stand.spec) %>% 
+  distinct
+```
+
+```{r}
+spec_cover <- bdm_sp %>% 
+  # calculate average species cover per vegetation type using the plots belonging to that vegetation type
+  group_by(sitename, LABEL_DE, stand.spec) %>% 
+  summarise(abun = mean(abun)) %>% 
+  # join with vegetation cover data
+  right_join(veg_cover, by = c("sitename", "LABEL_DE")) %>% 
+  # weight species abundance by veg cover
+  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c)
+
+# print table with missing species data for vegetation types
+missing_vegtype <- spec_cover %>% 
+  filter(is.na(stand.spec)) %>% 
+  group_by(sitename, LABEL_DE) %>% 
+  summarise(cov_b = round(sum(veg_cov_b), 2), 
+            cov_c = round(sum(veg_cov_c), 2)) %>%
+  filter(!cov_b == 0|!cov_c == 0)
+
+# for missing data, use species data from other plots
+missing <- bdm_sp %>% 
+  filter(LABEL_DE %in% missing_vegtype$LABEL_DE) %>% 
+  group_by(LABEL_DE, stand.spec) %>% 
+  summarise(abun = mean(abun))
+spec_cover_na <- spec_cover %>% 
+  filter(is.na(stand.spec)) %>%
+  dplyr::select(sitename, LABEL_DE, veg_cov_b, veg_cov_c) %>% 
+  left_join(missing, by = "LABEL_DE") %>% 
+  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c) 
+spec_cover_bc <-  spec_cover %>% 
+  filter(!is.na(stand.spec)) %>%
+  bind_rows(., spec_cover_na) %>% 
+  arrange(sitename, LABEL_DE, stand.spec) %>% 
+  dplyr::select(sitename, stand.spec, spec_abun_b, spec_abun_c) %>% 
+  ungroup() %>% 
+  left_join(genus_fam, by = "stand.spec")
+
+saveRDS(spec_cover_bc, "RDS_files/01_Species_abundance_zoneBC_swiss.rds")
+```

### 2022-09-08 02_CWM_estimates_vegetation.R
priors to the scale of the data

@@ -1,6 +1,6 @@
 ## ---------------------------
 ##
-## Script name:03_CWM_estimates_vegetation.R
+## Script name:02_CWM_estimates_vegetation.R
 ##
 ## Purpose of script: Calculating CWM values of the vegetation
 ##
@@ -32,23 +32,17 @@ set.seed(1)
 # Trait data ----
 # Scotland ----
 plantheight <- readRDS("RDS_files/01_PlantHeight.rds") %>% 
-  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
   ungroup() 
 ldmc <- readRDS("RDS_files/01_LDMC.rds") %>% 
-  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
   ungroup()
 sla <- readRDS("RDS_files/01_SLA.rds") %>% 
-  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
   ungroup()
 la <- readRDS("RDS_files/01_LA.rds") %>% 
-  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
   ungroup()
 c <- readRDS("RDS_files/01_CN.rds") %>% 
-  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
   ungroup() %>% 
   rename(LeafC = C)
 n <- readRDS("RDS_files/01_CN.rds") %>% 
-  mutate(sitename = as.factor(sitename), species = as.factor(species)) %>% 
   ungroup() %>% 
   rename(LeafN = N)
 
@@ -58,11 +52,8 @@ lTrait <- list(PlantHeight = plantheight,
                LA = la,
                LeafC = c,
                LeafN = n)
-# Switzerland ----
-bdm_traits <- readRDS("RDS_files/01_Traits_Swiss.rds")
 
-# Gapdata ---- 
-dfTRAIT <- readRDS("RDS_files/01_TRY_raw_traits.rds")
+dfTRY <- readRDS("RDS_files/01_Clean_TRY_data.rds")
 
 # Pollination and PFT ----
 dfPFT <- readRDS("RDS_files/Polmode_pft_vegetation.rds")
@@ -70,10 +61,6 @@ dfPFT <- readRDS("RDS_files/Polmode_pft_vegetation.rds")
 # Species data ----
 dfABUN_a <- readRDS("RDS_files/01_Species_abundance_a.rds") 
[…]

### 2022-09-08 03_CWM_estimates_pollen.R

@@ -26,12 +26,12 @@ if(!require(mcmc)) install.packages("mcmc")
 if(!require(coda)) install.packages("coda")
 if(!require(Hmisc)  )install.packages("Hmisc") 
 if(!require(furrr))  install.packages("furrr")
-if(!require(LCVP)) install.packages("LCVP")
-if(!require(lcvplants)) install.packages("lcvplants")
+
+set.seed(1)
 
 ## Load and prepare data -----
 # Trait data ----
-dfTRAIT <- readRDS("RDS_files/02_Gapfilled_traits.rds")
+dfTRAIT <- readRDS("RDS_files/01_TRY_raw_traits.rds")
 
 # Pollen data ----
 dfPOL_Scot <- readRDS("RDS_files/01_pollen_data_Scot.rds")
@@ -40,124 +40,23 @@ dfPOL <- bind_rows("Scotland" = dfPOL_Scot, "Switzerland" =
                      dfPOL_Swiss, .id = "country") %>% 
   ungroup()
 
-selectedcountry <- "Scotland"
-selectedtrait <- "LA"
-selectedabun <- str_subset(colnames(dfPOL), "percent")[57]
-# Calculate CWM ----
-cwmlogfunc <- function(selectedtrait, selectedcountry, selectedabun){
-  pol <- dfPOL %>% 
-    # select country
-    filter(country == selectedcountry) %>%
-    dplyr::select(sitename, pollentaxon, pollen_abun = all_of(selectedabun)) %>% 
-    # filter out unidentified
-    filter(!pollentaxon == "unindentified",
-           !is.na(pollen_abun)) %>% 
-    pivot_wider(names_from = pollentaxon, values_from = pollen_abun) %>% 
-    # filter empty columns (taxon not in that country)
-    discard(~all(is.na(.))) 
-  
-  # save sitenames
-  sitenames <- pol %>% 
-    pull(sitename)
-  # save pollen taxa
-  pollentaxa <- colnames(pol[, 2:ncol(pol)])
-  
-  Trait <- dfTRAIT %>% 
-    filter(country == selectedcountry) %>% 
-    filter(pollentaxon %in% pollentaxa) %>%
-    arrange(match(pollentaxon, pollentaxa)) %>% # order in the same way as pollen data
-    pull(selectedtrait) 
-  
-  Tax <- dfTRAIT %>%
-    filter(country == selectedcountry) %>% 
-    filter(pollentaxon %in% pollentaxa) %>%
[…]

### 2022-09-08 03_Pollen_diagrams.R
tidy script

@@ -20,17 +20,9 @@
 
 ## Load packages
 if(!require(tidyverse)) install.packages("tidyverse")
-
-## Load packages
-if(!require(tidyverse)) install.packages("tidyverse")
 if(!require(rioja)) install.packages("rioja") # plot pollen diagrams
 if(!require(ggrepel)) install.packages("ggrepel") # repel point labels
 if(!require(readxl)) install.packages("readxl") # repel point labels
-
-# - compare representation of pollen in vegetation
-# - classify by pollination mode
-# - classify by woody/non-woody
-# - compare with and without correction factors
 
 
 ## Prepare data ----
@@ -64,9 +56,9 @@ polmode <- polmode %>%
 ## Plot pollen data ----
 pol <- pollen %>% 
   ungroup %>% 
-  select(-adjustedpercent) %>% 
+  select(sitename, pollentaxon, percent) %>% 
   pivot_wider(names_from = pollentaxon, values_from = percent) %>% 
-  select(-sitename, -Liliaceae, -`Pedicularis palustris`) %>% 
+  mutate(across(where(is.numeric), ~.*100)) %>% 
   select(where(~ any(. > 1))) %>% # filter low values
   as.data.frame()
 sitenames <- pollen %>% pull(sitename)  %>% unique
@@ -98,8 +90,6 @@ pol_plot <-
     y.tks = 1:16,
     y.tks.labels = sitenames
   )
-rm(list=setdiff(ls(), c("pol_plot")))
-
 
 ## Prepare data - Switzerland ----
 pollen <- readRDS("RDS_files/01_Pollen_data_Swiss.rds")
@@ -137,10 +127,10 @@ polmode <- polmode %>%
 ## Plot pollen data ----
 pol <- pollen %>% 
   ungroup %>% 
-  select(-adjustedpercent) %>% 
+  select(sitename, pollentaxon, percent) %>% 
   pivot_wider(names_from = pollentaxon, values_from = percent) %>% 
-  select(-sitename) %>% 
-  select(where(~ any(. > 1))) %>% # filter low values
+  mutate(across(where(is.numeric), ~.*100)) %>% 
+  select(where(~ any(. > 5))) %>% # filter low values
   as.data.frame()
 sitenames <- pollen %>% pull(sitename) %>% unique
 bar_colour <- polmode %>% 
@@ -158,7 +148,7 @@ par(fig=c(0.01, 1, 0.07, 0.8),
[…]

### 2022-09-08 Manuscript.Rmd
adjusted introduction after comments

@@
 
+## Abstract
+
 ## Introduction
 
-In the past two decades, plant functional traits have become an important tool in ecology and biogeography for the study of consequences of changing plant composition on ecosystem functioning and the effect of environment factors on plant composition [@Hevia_2017].
-Recently, considerable interest in the trait-based approach has arisen in palaeoecology as well [@Brussel_2018; @Blaus_2020; @Carvalho_2019; @Reitalu_2015; @Sande_2019].
-Applying the trait-based approach could offer a new way of interpreting pollen data and increase our understanding of the drivers of vegetation change on timescales of centuries to millennial.
-For instance, @Sande_2019 combined functional trait and pollen data to study the effect of climate and human disturbances on vegetation composition in the Amazon on a 7000-year time scale.
+Functional traits provide a better insight in to the processes that drive biodiversity and how biodiversity affects ecosystem functioning [@Hevia_2017].
+While the functioning of modern ecosystems is regularly contingent on past human and environmental factors[@Levis_2018; @Jogiste_2017; @Blonder_2018], the observational duration of trait-based studies is generally short[@Estes_218].
+So to extend the time scales of ecosystem functioning research, considerable interest in the trait-based approach has arisen in palaeoecology [@Brussel_2018; @Blaus_2020; @Carvalho_2019; @Reitalu_2015; @Sande_2019].
+Applying the trait-based approach in palaeoecological studies could offer a new way of interpreting pollen data and increase our understanding of the drivers of vegetation change on timescales of centuries to millennia.
+For instance, @Sande_2019 found that on a 7000-year time scale human disturbance superseeds climate effects on plant trait composition in the Amazon.
 
 Although palaeoecological studies that have applied the trait-based approach show promising results, it is a relatively new approach of which the intricacies are underexplored.
-Little is known about how representative the pollen record is of plant trait composition.
-[-include what others have found-.]{.underline} To improve our understanding of the relationship between the trait composition of the vegetation and trait compositions reconstructed from pollen records, we here use surface pollen samples.
-Surface pollen assemblages approximate vegetation composition in recent years and are commonly used in palaeoecology to verify pollen-vegetation relationships [@Birks_2016].
-Here, we use surface samples of 16 Scottish and 26 Swiss sites to test the relationship between reconstructed trait composition from pollen data and vegetation trait composition.
+Little is known about how representative the pollen record is of past plant trait composition.
+Two main methodological issues arise when attempting to reconstruct functional trait composition from pollen data.
+First, pollen quantity does not exactly represent past plant abundance.
+Second, the low taxonomic resolution of the pollen record makes it not straightforward to assign traits to pollen taxa, as trait data is generally collected at the species level.
 
-Pollen quantity does not directly represent past plant abundance, taxa can become over- or underrepresented in the pollen record because of differences in pollen production and dispersal.
-[-something about pollination syndrome and plant function type]{.underline} - To correct for these differences pollen productivity factors can be use, these correction factors down-weight the contribution of dominant taxa.
-Furthermore, comparing between functional groups based on growth form or pollination syndrome could give insights in to the representativeness of the pollen assemblage to the vegetation.
-[-something about why pollination syndrome, distance from sampling site]{.underline} [and plant function type would matter to representation -]{.underline}
+Taxonomic differences in pollen dispersal and production can cause an over- or underrepresentation of taxa in the pollen record.
+Wind-pollinated and woody taxa are generally better represented in the pollen record than herbaceous and insect-pollinated taxa as their pollen are produced in larger quantities and dispersed over greater distances.
+The pollen source area for wind pollinated taxa can therefore be a couple magnitudes greater than that of wind pollinated taxa [\@Meltsov_2011](Bunting 2022?) (extend argument?
+bunting 2003, bunting 2022).
+It has therefore been suggested by @Blaus_2020 to treat woody and non-woody taxa separately in the reconstruction of functional composition from pollen data.
+Another or additional way to correct for taxon specific differences use pollen productivity estimates, these correction factors down-weight the contribution of dominant taxa [@Githumbi_2022].
 
-Perhaps the most challenging in reliably reconstructing past trait is to attribute functional trait data to the taxa found in pollen records, which can rarely be resolved to the species level [@Birks_2020].
-Trait variation within taxa can be substantial [@Siefert_2015].
+Perhaps the most challenging in reliably reconstructing past trait is to attribute functional trait data to the taxa found in pollen records.
+It is rarely possible to resolve the pollen record to the species level [@Birks_2020].
+Trait variation within taxa can be substantial depending on the trait and the diversity of the taxon [@Albert_2010; @Prinzing_2008].
 It can be expected that this especially problematic in the case of large, abundant plant families that cannot easily be resolved to a lower taxonomic level, such as the Poaceae family.
-Also small families where the trait variation is high could be problematic, such as is the case with the Fabaceae family where the height variation can be ...-fold (for instance Ulex shrub height and clover).
-Vegetation types with a large proportion of taxa with large trait variation might therefore be less reliably reconstructed.
-One way of dealing with the uncertainty caused by trait variation in pollen taxa is by using Bayesian models.
-Using a Bayesian approach for modelling trait composition from pollen data allows for the propagation of uncertainty in the taxon level trait reconstruction to the community level.
+@Blaus_2020 and @Carvalho_2019 proposed that the high species number included in herbaceous pollen taxon could be one of the reasons why herbaceous vegetation was less well represented by pollen based functional diversity reconstructions.
+In both studies, however, traits values of pollen taxa were calculated by averaging over the available species level trait information.
+In this study, we use a Bayesian approach for modelling trait composition from pollen data, thereby allowing for the propagation of uncertainty in the taxon level trait reconstruction to the community level.
 
 With this study, we aim to test the validity of using of pollen records for plant trait composition reconstruction.
-More specifically, we will test the effect of vegetation composition, taxonomic resolution and trait choice on the representativeness of pollen assemblages for functional diversity reconstructions.
-We examine the effectivity of pollen production factors to improve the pollen - functional diversity relationship.
-Lastly, we test the representation of plant functional groups based on growth form and pollination syndrome in the functional composition reconstruction.
+We use surface pollen samples as they approximate vegetation composition in recent years [@Birks_2016].
+We examine the effectiveness of pollen production factors to improve the pollen - functional diversity relationship.
+Furthermore, we test the representation of different plant functional groups by the pollen-based functional trait reconstruction.
+Laslty, we investigate the effect of taxonomic resolution on the reconstructions.
 
 ## Method
+
+### Study locations
+
+Here, we use surface samples of 16 Scottish and 26 Swiss sites to test the relationship between reconstructed trait composition from pollen data and vegetation trait composition.
 
@@
 
-### Vegetation surveys
+### Vegetation data
 
@@
 
-### Trait collection
+### Trait data
 
@@
 Five leaf samples per species for specific leaf area (SLA), leaf area (LA), leaf dry matter content (LDMC), leaf carbon and leaf nitrogen were collected in sealable plastic bags.
-Fresh weight and surface area were measured within 72 hours of collection using a carat scale and flatbad scanner.
+Fresh weight and surface area were measured within 72 hours of collection using a carat scale and flatbed scanner.
 Samples were dried at 70 C for 72 hours before measuring dry weight.
@@
 
-### Assigning species to pollen taxa
-
-For reconstructing plant trait composition from pollen data, assumptions need to be made about the species belonging to the pollen taxa in the data set.
-We used the *rgbif* package to download the current distribution of species belonging to each pollen taxon in Scotland and Switzerland from the Global Biodiversity Information Facility database (REF).
-Species distribution data was queried using the pollen taxonomic name.
-
-### Preparation of pollen data
+### Pollen data
 
@@
 
-    let to, on average, exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish
-    pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%) and `round(mean(missing_ppe_swiss$percent_missing),2)*100`% of the Swiss pollen data (max `r round(max(missing_ppe_swiss$percent_missing),2)*100`%).
+    let to, on average, exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%) and `round(mean(missing_ppe_swiss$percent_missing),2)*100`% of the Swiss pollen data (max `r round(max(missing_ppe_swiss$percent_missing),2)*100`%).
 
-    ### 
-    Calculation of community weighted mean
+-   Pollen were divided in to plant functional groups according to [@Reitalu_2019] trees shrubs (including Ericaceae family??), and herb (including grasses)
+
+### Assigning species to pollen taxa
+
+For reconstructing plant trait composition from pollen data, assumptions need to be made about the species belonging to the pollen taxa in the data set.
+We used the *rgbif* package to download the current distribution of species belonging to each pollen taxon in Scotland and Switzerland from the Global Biodiversity Information Facility database (REF).
+Species distribution data was queried using the pollen taxonomic name
+
+### Calculation of community weighted mean
 
 To test the representativeness of pollen for plant trait composition reconstructions, the community weighted mean trait values was calculated for the pollen data and for the vegetation data in the three nested circles: zone A (0-3.4m) zone B (10-100 m) and zone C (100 - 100 m).
-The trait distribution on the species or the pollen taxonomic level was estimated assuming that the trait data follows a lognormal distribution for the
-    likelihood, so that$$trait_{i}^{observed} \sim { lognormal(\mu_{j},\sigma_j)}$$Where *i* represents each trait observation of a taxon (*j*) from the trait data.
+The trait distribution on the species or the pollen taxonomic level was estimated assuming that the trait data follows a lognormal distribution for the likelihood, so that$$trait_{i}^{observed} \sim { lognormal(\mu_{j},\sigma_j)}$$Where *i* represents each trait observation of a taxon (*j*) from the trait data.
 $\mu_{j}$ and $\sigma_{j}$ are the mean trait value and the standard deviation per taxon, respectively.
@@
 The *CWM~­~^mean^* and *CWM^sd^* were subsequently used in the estimation of the correlation between the trait composition of the vegetation and the pollen data.
-We used a multivariate normal distribution to estimate correlation coefficient *ρ* so that,
-    $$[CWM_{s}^{veg},CWM_{s}^{veg}] \sim MultivariateNormal([\mu_{veg},\mu_{veg}],\sum) $$
+We used a multivariate normal distribution to estimate correlation coefficient *ρ* so that, $$[CWM_{s}^{veg},CWM_{s}^{veg}] \sim MultivariateNormal([\mu_{veg},\mu_{veg}],\sum) $$
 
@@
 
-    Correlations were calculated for all zones and three traits (currently): SLA, LA and plant
-    height.
-Furthermore, four different \"treatments\" were assigned to test what factors could influence representation of trait composition by pollen data: the use of correction factors, the separation of vegetation and pollen data in to plant functional type and pollination mode.
+Correlations were calculated for all zones and three traits (currently): SLA, LA and plant height.
+Furthermore, four different "treatments" were assigned to test what factors could influence representation of trait composition by pollen data: the use of correction factors, the separation of vegetation and pollen data in to plant functional type and pollination mode.
 Lastly, the vegetation was aggregated to genus level and family level to test the effect of taxonomic resolution on the reconstructions.
@@
 
-## ![](Figures/veg-Family.png)Appendix
+## ![](Figures/veg-Family.png)
+
+## Discussion
+
+-   representativeness of trait databases for trait measurements in the field?
+
+    -   Sandel (2015) - trait databases are biased towards big leaved and big seeded plants
+    -   Prinzing (2008) - Less lineages -- more trait variation
+
+-   Comparison between Switzerland and Scotland methods of sampling
+
+-   Some pollen types consider many more species than others for example in this study XX .
+This is also indentified as a problem in plant richness studies (Reitalu 2019) and trait rep (Carvalho, Blaus).
+although bayesian xxx
+
+-   size of the relevant source area (Bunting 2003)
+
+## Appendix
 
@@
 
+```
+
+### Site characteristics
+
+insert pollen plot
+
+```{r}
+# source("03_pollen_diagrams.R")
+```
+
+```{r}
+source("01_PFT_table_vegetation.R")
+sum
 ```

### 2022-09-08 05_Scatterplots.R
initial commit

@@ -1,0 +1,43 @@
+## ---------------------------
+##
+## Script name: 04_Create_scatterplots
+##
+## Purpose of script: Create scatterplots to go in to the appendix
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-09-08
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   
+##
+## ---------------------------
+
+
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
+
+## Load data
+dfCWM_veg <- readRDS("RDS_files/04_CWM_estimates_vegetation.rds")
+dfCWM_pol <- readRDS("RDS_files/04_CWM_estimates_pollen.rds")
+
+# join data
+
+# filter draws 
+
+# Produce plots
+(p_scatter <-
+    ggplot(cwm, aes(x = pollen_mean, y = veg_mean, color = zone)) +
+    geom_point() +
+    scale_fill_manual() +
+    scale_x_continuous("Pollen") +
+    scale_y_continuous("Vegetation") +
+    facet_wrap(country~trait, scales = "free") +
+    ggtitle("Corrected pollen data") +
+    theme_bw() +
+    theme(legend.position =  "bottom"))
+

### 2022-09-08 04_Correlation_bayesian.R
merge graphs in treatments

@@ -22,9 +22,16 @@
 if(!require(tidyverse)) install.packages("tidyverse")
 if(!require(rjags)) install.packages("rjags")
 if(!require(runjags)) install.packages("runjags")
-if(!require(mvtnorm)) install.packages("mvtnorm") #to generate correlated data with rmvnorm.
-if(!require(car)) install.packages("car") #To plot the estimated bivariate normal distribution.
-library(gridExtra)
+if(!require(gridExtra)) install.packages("gridExtra")
+
+# load PFT summary table
+# filter sites with to little vegetation cover of particular type
+sum <- readRDS("RDS_files/01_Percentage_cover_pft.rds")
+dropsites <- sum %>% 
+  transmute(drop_herb = case_when(`Non-Woody`<=5 ~ paste(zone, sitename)),
+         drop_trsh = case_when(Woody <= 5 ~ paste(zone,sitename)),
+         drop_wind = case_when(wind <= 5 ~ paste(zone,sitename)),
+         drop_nowind = case_when(`not wind` <= 5 ~ paste(zone,sitename)))
 
 # collate data
 traits <- c("LA","SLA","PlantHeight")
@@ -57,30 +64,19 @@ dfCWM_pol <- pollen_files %>%
                 as.data.frame() %>% 
                 rownames_to_column("parameter") %>% 
                 filter(., str_detect(parameter, "cwm\\[")) %>%
-               
-                mutate(label = .y,
-                       id = seq(1:nrow(.))
-                       )
-              ) %>% 
+                mutate(label = .y)) %>% 
   bind_rows() %>% 
   as_tibble() %>% 
-  dplyr::select(sitename,label, pollen_mean = Mean, pollen_sd = SD) %>% 
-  mutate(trait = str_extract(label, pattern = paste(traits, collapse = "|")) %>% 
-           str_remove_all("_"),
-         sitename = str_remove(sitename, pattern = "X"),
-         growthform = case_when(str_detect(label, pattern = "herb") ~ "herb",
-                                str_detect(label, pattern = "trsh") ~ "trsh",
-                                TRUE ~ "allpft"),
-         pollination = case_when(str_detect(label, pattern = "not wind") ~ "not wind",
[…]

### 2022-09-08 01_Swiss_data_compilation.R
fix unit of tree data

@@ -39,6 +39,10 @@ plot_id <- bdm_pollen %>%
 # check sites 
 bdm_meta <- bdm_meta %>% 
   filter(aID_STAO %in% plot_id)
+ 
+# bdm_meta <- bdm_meta %>% 
+#   filter(aID_STAO %in% bdm_sp$plot)
+# write.csv(bdm_meta, file = "Data/Swiss_LC/all_BDM_plots_coord.csv")
 
 # Prepare sp code for merge with trait and abundance data
 bdm_sp_codes <- bdm_sp_codes %>% 
@@ -54,7 +58,10 @@ bdm_sp_codes <- bdm_sp_codes %>%
                         `Helix hedera` = "Hedera helix",
                         `Leontondon helveticus` = "Leontodon helveticus",
                         `Leotondon Autumnalis` = "Leontodon autumnalis",
-                        `Cirisium spinosissimum` = "Cirsium spinosissimum"),
+                        `Cirisium spinosissimum` = "Cirsium spinosissimum",
+                        `vaccinium myrtillus` = "Vaccinium myrtillus",
+                        `vaccinium vitis idaea` = "Vaccinium vitis-idaea",
+                        `Anthoxanthum odoratum aggr.` = "Anthoxanthum odoratum"),
          Species_code = recode(Species_code, 
                                alcxan = "alcvul", # Alchemilla vulgaris current code alcxan is not in species or trait data, alcvul is
                                soldal = "solalp" # Soldanella alpinia idem
@@ -70,8 +77,7 @@ bdm_sp <- bdm_sp %>%
   filter(!abun == 0) %>% 
   # join species codes
   left_join(bdm_sp_codes, by = c("species_code" = "Species_code")) %>% 
-  # filter for relevant plots
-  filter(plot %in% plot_id) 
+  dplyr::select(plot, subplot, abun, binom, Species, species_code)
 
 # standardize species names
 spec <- bdm_sp %>% 
@@ -82,10 +88,11 @@ spec <- bdm_sp %>%
 lcvp <- lcvp_search(spec) # find synonyms
 
 synonyms <- lcvp %>% 
-  dplyr::select(binom = Search, stand_spec = Output.Taxon) %>% 
-  mutate(stand_spec = word(stand_spec, 1, 2)) # drop authorship
+  dplyr::select(binom = Search, stand.spec = Output.Taxon, family = Family) %>% 
+  mutate(stand.spec = word(stand.spec, 1, 2),# drop authorship
[…]

### 2022-09-20 01_PFT_table_vegetation.R
missing species

@@ -158,7 +158,11 @@ pft <- traits %>%
   # Ericaceae specified as herb in pollen data
   mutate(growthform = case_when(AccSpeciesName %in% c("Calluna vulgaris", 
                                                       "Erica cinerea",
-                                                      "Erica tetralix") ~ "herb",
+                                                      "Erica tetralix",
+                                                      "Vaccinium myrtillus",
+                                                      "Empetrum nigrum",
+                                                      "Vaccinium vitis-idaea",
+                                                      "Vaccinium nitens") ~ "herb",
                                 TRUE ~ growthform))
 
 

### 2022-09-20 01_Prepare_TRY_data.R
increase error risk

@@ -59,7 +59,7 @@ trait <- trait_raw %>%
           ErrorRisk
           ) %>%
   mutate(ErrorRisk2 = ifelse(is.na(ErrorRisk), 0, ErrorRisk)) %>%
-  filter(ErrorRisk2 < 8) 
+  filter(ErrorRisk2 < 4) # error risks based on Z-scores, see TRY 2019 paper
 
 # Get meta data
 meta <- trait %>% 
@@ -79,7 +79,6 @@ units <- trait %>%
 # Subsetting out traits and naming them
 # if origlname or unit is blank but there are no duplicates we are keeping it. 
 trait2 <- trait %>%
-  # remove problem data where there are replicates for ObservationID
   mutate(remove = case_when(TraitID == 14 & OriglName == "N amount%" ~ 1, 
                             TraitID == 14 & OriglName == "N_senesced_leaf" ~ 1,
                             TraitID == 15 & OriglName == "P_senesced_leaf" ~ 1, 
@@ -221,8 +220,10 @@ trait4 <- trait3 %>%
               names_from = CleanTraitName, values_from = StdValue) %>% 
   # remove duplicated values
   filter(!(duplicated(.[4:10]) & !is.na(OrigObsDataID))) %>% 
-  # conver plantheight from m to cm
-  mutate(PlantHeight = PlantHeight * 100)
+  # convert plantheight from m to cm
+  mutate(PlantHeight = PlantHeight * 100) %>% 
+  # remove zeros from the measurements
+  mutate(across(LeafN:LA, ~na_if(.,0)))
 
 # add pollen to species translation table ----
 spec <- readRDS("RDS_files/02_PollenType_species.rds") %>% 
@@ -231,13 +232,13 @@ spec <- readRDS("RDS_files/02_PollenType_species.rds") %>%
 # check for missing species in the trait data
 trait_sp <- trait4$AccSpeciesName %>% unique
 pol_sp <- spec$stand.spec %>% unique()
-
 misspec <-
   pol_sp[!pol_sp %in% trait_sp] %>%
   sort %>% unique
 
 # improve matching by standardizing species names
-# lcvp <- lcvp_search(str_subset(trait_sp, pattern = " ")) 
+# lcvp <- lcvp_search(str_subset(trait_sp, pattern = " "),
+#                     progress_bar = TRUE) 
[…]

### 2022-09-20 01_Prepare_pollen_data.R
tidy script

@@ -46,15 +46,20 @@ dfPOL <- pollen_raw %>%
   mutate(sitename = recode(sitename, C014.1 = "C014"),
          # convert count to numeric       
          count = as.numeric(count)) %>% 
+  # remove zeros
+  filter(!count == 0) %>% 
+  mutate(pollentaxon = recode(pollentaxon, Tsuga = "unindentified", 
+                              Liliaceae = "unindentified", Drosera = "unindentified")) %>%  # probably misidentified
   #correct pollen taxon names
   mutate(pollentaxon = 
            # remove "-type" to facilitate GBIF search
            str_remove(pollentaxon, "-type") %>% 
            # Typos
            recode(Caryophyllaccea = "Caryophyllaceae",
-                  Lilaceae = "Liliaceae",
                   "Juniperus communis" = "Juniperus")
          ) %>% 
+  # filter out unidentified
+  filter(!pollentaxon %in% c("unindentified")) %>% 
   filter(!is.na(count))
 
 # percentage data with missing ppe's 
@@ -77,7 +82,7 @@ dfPOL_nc <- dfPOL %>%
 # make function to continue draw from rnorm till RPP is positive (otherwise negative percentages)
 norm_positive <- function(RPP, RPP_sd){
   success <- FALSE
-  while(!success){
+  while (!success) {
     x <- rnorm(1, RPP, RPP_sd)
     success <- x > 0
     }
@@ -107,7 +112,7 @@ dfPOL <- dfPOL_nc %>%
   left_join(dfPOL_cor, by = c("sitename", "pollentaxon")) %>% 
   # add PFT and pollination mode
   left_join(polmode, by = "pollentaxon") %>% 
-  ungroup()
+  ungroup() 
 
 saveRDS(dfPOL, "RDS_files/01_Pollen_data_Scot.rds")
 
@@ -121,14 +126,18 @@ dfPOL <- read.csv("Data/Pollen_count_Switserland-checked.csv", skip = 1, sep = "
                values_to = "count") %>%
   mutate(count = as.numeric(count)) %>% # change variable types
   dplyr::select(sitename, pollentaxon = Sample.name, count) %>%  # select relevant rows
-  filter(!pollentaxon == "Tsuga") %>%  # probably misidentified
+  mutate(pollentaxon = recode(pollentaxon, Tsuga = "unindentified", 
+         Liliaceae = "unindentified", Drosera = "unindentified")) %>%  # probably misidentified
+  # remove zeros
[…]

### 2022-09-20 02_CWM_estimates_vegetation.R
fix convergence issues, tidy script

@@ -18,13 +18,12 @@
 ## ---------------------------
 
 ## Load packages
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(rjags))  install.packages("rjags")
-if(!require(runjags)) install.packages("runjags")
-if(!require(mcmc)) install.packages("mcmc")
-if(!require(coda)) install.packages("coda")
-if(!require(Hmisc)  )install.packages("Hmisc") 
-if(!require(furrr))  install.packages("furrr")
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(rjags))  install.packages("rjags")
+if (!require(runjags)) install.packages("runjags")
+if (!require(mcmc)) install.packages("mcmc")
+if (!require(coda)) install.packages("coda")
+if (!require(furrr))  install.packages("furrr")
 
 set.seed(1)
 
@@ -85,11 +84,6 @@ zoneC <- dfABUN_bc %>%
   dplyr::select(sitename, stand.spec, genus, family = fam, abun = spec_abun_c, growthform, polmode)
 lABUN <- list("zoneA" = zoneA, "zoneB" = zoneB, "zoneC" = zoneC)
 
-selectedtrait <- "LA"
-selectedabun <- "zoneA"
-selectedpolmode <- unique(dfPFT$polmode)
-selectedpft <- unique(dfPFT$growthform)
-selectedtaxres <- "stand.spec"
 
 trsh <- c("tree", "shrub")
 herb <- c("herb", "grass", "fern")
@@ -119,46 +113,30 @@ dfTRY <-
                         )
              
 # Scotland----
-if(1){ 
+if (0) { 
 cwm_func <- function(selectedabun, selectedtrait, selectedtaxres, selectedpolmode, 
                      selectedpft){
-  taxa <- lABUN %>% 
+  abun <- lABUN %>% 
     pluck(selectedabun) %>% 
-    arrange(sitename, selectedtaxres) %>% 
     # optional filter for polmode and pft
     filter(polmode %in% selectedpolmode) %>% 
     filter(growthform %in% selectedpft) %>% 
-    drop_na() %>% 
-    pull(selectedtaxres) %>% 
-    unique()
-  
-  Ab <- lABUN %>% 
-    pluck(selectedabun) %>% 
-    arrange(sitename, selectedtaxres) %>% 
-    # optional filter for polmode and pft
-    filter(polmode %in% selectedpolmode) %>% 
-    filter(growthform %in% selectedpft) %>% 
+    # rename column to selected taxonomic resolution
[…]

### 2022-09-20 03_CWM_estimates_pollen.R

@@ -19,13 +19,12 @@
 
 
 ## Load packages
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(rjags))  install.packages("rjags")
-if(!require(runjags)) install.packages("runjags")
-if(!require(mcmc)) install.packages("mcmc")
-if(!require(coda)) install.packages("coda")
-if(!require(Hmisc)  )install.packages("Hmisc") 
-if(!require(furrr))  install.packages("furrr")
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(rjags))  install.packages("rjags")
+if (!require(runjags)) install.packages("runjags")
+if (!require(mcmc)) install.packages("mcmc")
+if (!require(coda)) install.packages("coda")
+if (!require(furrr))  install.packages("furrr")
 
 set.seed(1)
 
@@ -40,85 +39,214 @@ dfPOL <- bind_rows("Scotland" = dfPOL_Scot, "Switzerland" =
                      dfPOL_Swiss, .id = "country") %>% 
   ungroup()
 
-# selectedcountry <- "Switzerland"
-# selectedtrait <- "SLA"
-# selectedabun <- str_subset(colnames(dfPOL), "percent")[52]
+selectedcountry <- "Scotland"
+selectedtrait <- "LA"
+selectedabun <- str_subset(colnames(dfPOL), "percent")[1]
 
-normaltraits <- c("PlantHeight","SLA","LA", "LeafN", "LDMC")
+normaltraits <- c("PlantHeight","SLA","LA")
+
+# No subset ----
+if (0) {
+  cwmlogfunc <- function(selectedtrait, selectedcountry, selectedabun){
+    trait <- dfTRAIT %>% 
+      filter(country == selectedcountry) %>% 
+      # drop NA's and 0's
+      dplyr::select(pollentaxon, all_of(selectedtrait)) %>% 
+      mutate(across(where(is.double), ~na_if(.,0))) %>% 
+      drop_na() %>% 
+      # Sort alphabetically
+      arrange(pollentaxon) %>% 
+      # check number of observations
+      group_by(pollentaxon) %>% 
+      mutate(ntrait = n()) %>% 
+      filter(ntrait > 3)
+    
+    pol <- dfPOL %>% 
+      # select country
+      filter(country == selectedcountry) %>%
+      dplyr::select(sitename, pollentaxon, pollen_abun = all_of(selectedabun)) %>% 
+      drop_na()
+    
+    # save pollen taxa
[…]

### 2022-09-20 03_Vegetation_pollen_comparison.R
add swiss data

@@ -18,21 +18,24 @@
 ## ---------------------------
 
 ## Load packages
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(rioja)) install.packages("rioja") # plot pollen diagrams
-if(!require(ggrepel)) install.packages("ggrepel") # repel point labels
-if(!require(readxl)) install.packages("readxl") # repel point labels
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(rioja)) install.packages("rioja") # plot pollen diagrams
+if (!require(ggrepel)) install.packages("ggrepel") # repel point labels
+if (!require(readxl)) install.packages("readxl") # repel point labels
 
 ## Prepare data ----
 dfPOL_Scot <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
 dfPOL_Swiss <- readRDS("RDS_files/01_Pollen_data_Swiss.rds")
-dfPOL <- bind_rows(dfPOL_Scot, dfPOL_Swiss)
+dfPOL <- bind_rows("Scotland" = dfPOL_Scot, "Switzerland" = dfPOL_Swiss, .id = "country") %>% 
+  dplyr::select(country,sitename, pollentaxon, percent, adjustedpercent_mean) %>% 
+  mutate(sitename = str_remove(sitename, pattern = "X"))
 polmode <- readRDS("RDS_files/01_Pollination_mode.rds")
 polspec <- readRDS("RDS_files/02_PollenType_species.rds") %>% 
   dplyr::select(pollentaxon, stand.spec) %>% distinct() %>% as_tibble()
 dfABUN_a <- readRDS("RDS_files/01_Species_abundance_a.rds")
 dfABUN_bc <- readRDS("RDS_files/01_Species_abundance_bc.rds") %>% ungroup
 bdm_abun <- readRDS("RDS_files/01_Species_abundance_Swiss.rds") 
+bdm_abun_bc <- readRDS("RDS_files/01_Species_abundance_zoneBC_swiss.rds")
 
 # Summarize vegetation data by pollen type
 zoneA <- dfABUN_a %>%
@@ -43,9 +46,8 @@ zoneA <- dfABUN_a %>%
   filter(distance %in% c("0 meter", "1.5-3 meter")) %>% 
   # calculate species abundance on the plot level
   group_by(sitename, pollentaxon) %>% 
-  summarise(abun = sum(abun)) %>% 
-  mutate(abun = abun/sum(abun)) %>% 
[…]

### 2022-09-20 04_Correlation_bayesian.R
implement Multi t distribution, improve plots

@@ -15,23 +15,28 @@
 ## Notes:
 ##   
 ##
+## References:
+## 1 https://www.sumsar.net/blog/2013/08/bayesian-estimation-of-correlation/
+## 2 https://www.sumsar.net/blog/2013/08/robust-bayesian-estimation-of-correlation/
+## 3 http://doingbayesiandataanalysis.blogspot.com/2015/12/prior-on-df-normality-parameter-in-t.html
 ## ---------------------------
 
+set.seed(1)
 
 ## Load packages
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(rjags)) install.packages("rjags")
-if(!require(runjags)) install.packages("runjags")
-if(!require(gridExtra)) install.packages("gridExtra")
-
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(rjags)) install.packages("rjags")
+if (!require(runjags)) install.packages("runjags")
+if (!require(furrr)) install.packages("furrr")
+  
 # load PFT summary table
 # filter sites with to little vegetation cover of particular type
 sum <- readRDS("RDS_files/01_Percentage_cover_pft.rds")
 dropsites <- sum %>% 
-  transmute(drop_herb = case_when(`Non-Woody`<=5 ~ paste(zone, sitename)),
-         drop_trsh = case_when(Woody <= 5 ~ paste(zone,sitename)),
-         drop_wind = case_when(wind <= 5 ~ paste(zone,sitename)),
-         drop_nowind = case_when(`not wind` <= 5 ~ paste(zone,sitename)))
+  transmute(drop_herb = case_when(`Non-Woody`<= 5 ~ paste(zone, sitename)),
+         drop_trsh = case_when(Woody <= 5 ~ paste(zone, sitename)),
+         drop_wind = case_when(wind <= 5 ~ paste(zone, sitename)),
+         drop_nowind = case_when(`not wind` <= 5 ~ paste(zone, sitename)))
 
 # collate data
 traits <- c("LA","SLA","PlantHeight")
@@ -67,16 +72,18 @@ dfCWM_pol <- pollen_files %>%
                 mutate(label = .y)) %>% 
   bind_rows() %>% 
   as_tibble() %>% 
-  mutate(correction = case_when(str_detect(label, pattern = "adjustedpercent_draw") ~ "correction draw",
+  mutate(correction = case_when(str_detect(label, pattern = "adjustedpercent_draw") ~ str_extract(label, pattern = "draw[0-9]*$"),
[…]

### 2022-09-20 05_Scatterplots.R
improve plots

@@ -1,6 +1,6 @@
 ## ---------------------------
 ##
-## Script name: 04_Create_scatterplots
+## Script name: 05_Scatterplots
 ##
 ## Purpose of script: Create scatterplots to go in to the appendix
 ##
@@ -19,25 +19,109 @@
 
 
 ## Load packages
-if(!require(tidyverse)) install.packages("tidyverse")
+if (!require(tidyverse)) install.packages("tidyverse")
 
 ## Load data
 dfCWM_veg <- readRDS("RDS_files/04_CWM_estimates_vegetation.rds")
-dfCWM_pol <- readRDS("RDS_files/04_CWM_estimates_pollen.rds")
+dfCWM_pol <- readRDS("RDS_files/04_CWM_estimates_pollen.rds") %>% 
+  filter(!str_detect(correction,pattern = "draw"))
 
 # join data
-
-# filter draws 
+dfCWM <- dfCWM_veg %>% 
+  left_join(dfCWM_pol, by = c("country","sitename", "trait", "growthform","pollination")) %>% 
+  # create treatment column
+  mutate(treatment = case_when((pollination == "allpol" & growthform == "allpft" &
+                                  taxres == "stand.spec") ~ "Correction factor",
+                               taxres %in% c("family","genus") ~ "Taxonomic resolution",
+                               (pollination != "allpol" & growthform == "allpft") ~ "Pollination mode",
+                               (growthform != "allpft" & pollination == "allpol") ~ "Growth form"),
+         zone = str_remove(zone, pattern = "Scotland |Switzerland "))
 
 # Produce plots
-(p_scatter <-
-    ggplot(cwm, aes(x = pollen_mean, y = veg_mean, color = zone)) +
+labs_trait <- c(PlantHeight = "Height (cm)",
+          LA = "Leaf area (mm2)",
+          SLA = "SLA (mg/mm2)")
+labs_zone <- c(zoneA = "",
+               zoneB = "",
+               zoneC = "")
+
+windows()
+(p_correction <-
+    dfCWM %>% 
+    filter(treatment == "Correction factor") %>% 
+    ggplot(aes(x = pollen_mean, y = veg_mean, color = correction, shape = country)) +
     geom_point() +
-    scale_fill_manual() +
     scale_x_continuous("Pollen") +
     scale_y_continuous("Vegetation") +
-    facet_wrap(country~trait, scales = "free") +
-    ggtitle("Corrected pollen data") +
[…]

### 2022-09-20 Manuscript.Rmd
introduction and methods adjustments

@@
   word_document:
-    reference_docx: Manuscript-Markdown-template.docx
+    reference_docx: "Manuscript-Markdown-template.docx"
     fig_caption: yes
@@
 
-source("Write_bib_file.R") # Compile library for reference list
 source("01_Prepare_pollen_data.R")
@@
 library(knitr)
+
+# Datasets
+bdm_meta <- readRDS("RDS_files/01_Meta_data_BDM.rds")
 ```
@@
 Wind-pollinated and woody taxa are generally better represented in the pollen record than herbaceous and insect-pollinated taxa as their pollen are produced in larger quantities and dispersed over greater distances.
-The pollen source area for wind pollinated taxa can therefore be a couple magnitudes greater than that of wind pollinated taxa [\@Meltsov_2011](Bunting 2022?) (extend argument?
+The pollen source area for wind pollinated taxa can therefore be a couple magnitudes greater than that of wind pollinated taxa [@Meltsov_2011] (extend argument?
 bunting 2003, bunting 2022).
@@
 With this study, we aim to test the validity of using of pollen records for plant trait composition reconstruction.
-We use surface pollen samples as they approximate vegetation composition in recent years [@Birks_2016].
 We examine the effectiveness of pollen production factors to improve the pollen - functional diversity relationship.
@@
 
-### Study locations
+### Study sites
 
-Here, we use surface samples of 16 Scottish and 26 Swiss sites to test the relationship between reconstructed trait composition from pollen data and vegetation trait composition.
+We collected pollen, trait and species composition data from 16 sites in the Cairngorms National Park, Scotland, United Kingdom.
+Sites were selected to represent four vegetation types in the area: grassland, heath land, deciduous forest and coniferous forest.
+Sampling locations were at least 1 km from each other to avoid pollen source areas from overlapping with each other.
+Pollen, trait and species data was also available from `r length(bdm_meta)` sites across Switzerland were monitored in the context of the Biodiversity Monitoring Switzerland (BDM).
+This is a systematic sampling grid for the long-term monitoring of plant and animal species in Switzerland [@BDM_2022].
+Of these sites `r table(bdm_meta$habitat01)[1]` were forest and `table(bdm_meta$habitat01)[2]` grassland, spanning an elevation between `r min(bdm_meta$elevation)` and `r max(bdm_meta$elevation)`.
+Trait and species composition data from `r nrow(read.csv("Data/Swiss_LC/all_BDM_plots_coord.csv")) - length(bdm_meta)` were also used in this study.
 
@@
 
-Pollen data, plant species composition data and trait data were collected in 16 sampling locations in June 2021 in the Cairngorms National Park, Scotland, United Kingdom.
-Study design mostly followed the protocol of @Bunting_2013 (Appendix).
-Sampling locations were at least 1 km from each other to avoid pollen source areas from overlapping with each other.
-Sites were selected to represent four vegetation types in the area: grassland, heath land, deciduous forest and coniferous forest.
-Every sampling location consisted of three nested concentric rings of 3.4m, 100m and 1000m in radius.
-The inner two rings were surveyed in the field and for the outer ring vegetation cover was estimated using QGIS and the Scottish Habitat and Land cover map of 2020.
-To determine the composition of the modern pollen assemblage at every site, surface samples of moss were collected at the centre of each sampling location.
-Moss was collected from multiple location within a 1m radius from the centre.
-Only living tissue was sampled.
-
-An additional 26 pollen samples were collected from sites in the Swiss Biodiversity Monitoring network in summer 2020, that were selected based on an elevation gradient (?).
-Only the inner ring (0-3.4m) was surveyed in these sites.
-[(Check with Tom the exact method in Switzerland)]{.underline}
+We compare trait composition of vegetation with trait composition reconstructed from surface pollen samples.
+Surface pollen samples are collected from moss polsters and approximate vegetation composition in recent years [@Birks_2016] [Other ref?].
+Study design mostly followed the protocol of @Bunting_2013 (Appendix), which allows for the calculation of vegetation composition in concentric circles around the moss sampling point.
+We calculated the vegetation composition in three nested concentric rings of 3.4m, 100m and 1000m in radius.
+The inner ring (3.4m) was surveyed in the field, the middle ring (100m) was surveyed in Scotland, but not in Switzerland.
+The vegetation cover of the outer ring (1000m) and the middle ring for Switzerland was estimated using QGIS and land cover maps [@SwissFC; @SwissLC;\@scotisch_government;\@QGIS_software].
 
@@
 
-In the Scottish sites, the inner ring (0-3.4m) 17 plots of 0.5 m were surveyed: one in the center, 3 in every wind direction between 3, 6 and 10 meters, and one in the NE, NW, SE and SW direction (See design in Appendix).
-In the middle ring (10-100m) between 4 and 6 plots of 10 by 10 meter were selected on a stratified random design based on vegetation types that were identified in the fields.
-Vegetation abundance was estimated using percentage cover.
+In the inner ring (3.4m) 4 plots of 0.25 m^2^ (0.16m^2^ in Switzerland) were surveyed, one in every wind direction.
+At the Scottisch sites the middle ring was surveyed as well, between 4 and 6 plots of 10 by 10 meter were selected on a stratified random design based on vegetation types that were identified in the fields.
+Species abundance was estimated using percentage cover in Scotland.
+In Switzerland, each subplot was a quadrat with nine points (a 3 x 3 grid), and percentage coverage of the species present at the nine points was estimated.
 All species names were standardized according to the Leipzig Plant Catalogue using the *lcvplants* and *lcvp* R packages [@freiberg2020lcvp].
-(Check with Tom the exact method in Switzerland)
 
@@
 
-Trait data was collected in the middle ring.
+Three traits were selected for the analysis, specific leaf area, leaf area and and plant height.
+These traits reflect different components in the spectrum of plant strategies.
+XXX.
+In Scotland, trait data was collected in the middle ring (100m).
 For determining plant height 10 flowering of individuals were selected at every site if present, non-flowering individuals were selected if flowering individuals were not available.
-Five leaf samples per species for specific leaf area (SLA), leaf area (LA), leaf dry matter content (LDMC), leaf carbon and leaf nitrogen were collected in sealable plastic bags.
+Five leaf samples per species for specific leaf area (SLA), leaf area (LA), leaf dry matter content (LDMC), leaf carbon and leaf nitrogen were collected in seal-able plastic bags.
 Fresh weight and surface area were measured within 72 hours of collection using a carat scale and flatbed scanner.
-Samples were dried at 70 C for 72 hours before measuring dry weight.
-For measuring leaf carbon and nitrogen samples using ICP-MS were crushed and X.
+Samples were dried at 70 C for 72 hours to constant dry weight before measuring dry weight.
+In Switzerland, the trait data was collected for species present at the nine points in the 3 by 3 grid.
 
-For reconstructing trait composition from pollen data, previously gapfilled data from the TRY database *(will replace this by the raw TRY data as discussed)* was used [@kattge2020try].
-
-(Check with Tom the exact method in Switzerland)
+For reconstructing trait composition from pollen data and for classifying vegetation data in to plant growth forms and pollination mode, trait data from the TRY database was used [@kattge2020try] (Appendix).
 
@@
 
--   Pollen were extracted from the moss using standardized pollen preparation protocol (REF)
-
--   Pollen were identified using light microscope and the key of (REF)
-
-    -   pollen correction factors- new data of Githumbi 2022 [@Githumbi_2022],
-
-    let to, on average, exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%) and `round(mean(missing_ppe_swiss$percent_missing),2)*100`% of the Swiss pollen data (max `r round(max(missing_ppe_swiss$percent_missing),2)*100`%).
-
--   Pollen were divided in to plant functional groups according to [@Reitalu_2019] trees shrubs (including Ericaceae family??), and herb (including grasses)
-
-### Assigning species to pollen taxa
+Moss was collected from multiple location within a 1m radius in the centre of the inner ring.
+Only living tissue was sampled.
+Pollen were extracted from the moss using standardized pollen preparation protocol (REF).
+Pollen were identified using light microscope and the key of (REF).
+To test the ability of pollen correction factors to improve pollen - vegetation relationships, pollen productivity estimates from @Githumbi_2022 were used.
+Pollen productivity estimates were not available for all taxa, taxa without pollen productivity factors were excluded.
+On average, this led to the exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%) and `round(mean(missing_ppe_swiss$percent_missing),2)*100`% of the Swiss pollen data (max `r round(max(missing_ppe_swiss$percent_missing),2)*100`%).
+Pollen were divided in to two plant functional groups according to @Reitalu_2019 consisting of woody (trees and shrubs) ornon-woody species (grasses, herbs and dwarf shrubs of the Ericaceae family).
+@Reitalu_2019 was also used to classify pollen in to pollination mode, either insect-pollinated or not insect-pollinated.
 
@@
 We used the *rgbif* package to download the current distribution of species belonging to each pollen taxon in Scotland and Switzerland from the Global Biodiversity Information Facility database (REF).
-Species distribution data was queried using the pollen taxonomic name
+Species distribution data was queried using the pollen taxonomic name, i.e.
+the lowest taxonomic name the pollen could be identified to .
 
@@
 
-To test the representativeness of pollen for plant trait composition reconstructions, the community weighted mean trait values was calculated for the pollen data and for the vegetation data in the three nested circles: zone A (0-3.4m) zone B (10-100 m) and zone C (100 - 100 m).
-The trait distribution on the species or the pollen taxonomic level was estimated assuming that the trait data follows a lognormal distribution for the likelihood, so that$$trait_{i}^{observed} \sim { lognormal(\mu_{j},\sigma_j)}$$Where *i* represents each trait observation of a taxon (*j*) from the trait data.
+We use the community weighted mean (CWM) to characterize functional composition, which here is specified as the trait distribution of a taxon, i.e.
+the species in the vegetation data and the pollen taxon in the pollen data, weighted by its abundance.
+First, the trait distribution on the species or the pollen taxonomic level was estimated.
+We assume that the trait data follows a lognormal distribution for the likelihood, so that$$trait_{i}^{observed} \sim { lognormal(\mu_{j},\sigma_j)}$$Where *i* represents each trait observation of a taxon (*j*) from the trait data.
 $\mu_{j}$ and $\sigma_{j}$ are the mean trait value and the standard deviation per taxon, respectively.
 Vague priors were used for the taxon mean and standard deviation estimation.
-We modeled the community weighted mean (CWM) as a mixture of normal distributions that were weighted by corrected pollen percentages, so that
+We then modeled the community weighted mean (CWM) as a mixture of normal distributions that were weighted by the species abundance or pollen percentage, so that
 
@@
 
-Where *n* is the total number of taxa, and $\pi_{j}$ the abundance of each pollen taxon/species at a given site.
-The *CWM~­~^mean^* and *CWM^sd^* were subsequently used in the estimation of the correlation between the trait composition of the vegetation and the pollen data.
-We used a multivariate normal distribution to estimate correlation coefficient *ρ* so that, $$[CWM_{s}^{veg},CWM_{s}^{veg}] \sim MultivariateNormal([\mu_{veg},\mu_{veg}],\sum) $$
+Where *n* is the total number of taxa, and $\pi_{j}$ the abundance of each pollen taxon or species at a given site *s*.
+The mean *CWM~­~^mean^* and the standard deviation *CWM^sd^* of this modeled distribution was subsequently used in the estimation of the correlation between the trait composition of the vegetation and the pollen data.
+
+### Comparison between pollen and vegetation community weighted mean
+
+To test the correlation between the pollen based reconstruction of CWM and the vegetation CWM, we estimated the correlation coefficient *ρ* we assumed the data follows a multivariate t likelihood distribution so that, $$[CWM_{s}^{pol},CWM_{s}^{veg}] \sim t_{df}([\mu_{pol},\mu_{veg}],\sum) $$
+
+Where we estimated the degrees of freedom ($df$) of the multivariate t-distribution with and exponential prior with a mean of 30 as the t-distribution becomes almost normal at values of 30 and above (Kruscke, dbda),
+
+$$
+df \sim Exp(\lambda)
+$$
+
+$$
+\lambda = 1/30
+$$
+
+We used an uninformative, flat prior for $\rho$ so that,
 
@@
 
-Correlations were calculated for all zones and three traits (currently): SLA, LA and plant height.
-Furthermore, four different "treatments" were assigned to test what factors could influence representation of trait composition by pollen data: the use of correction factors, the separation of vegetation and pollen data in to plant functional type and pollination mode.
+Correlations were calculated for all zones and three traits: SLA, LA and plant height.
+Furthermore, four different treatments were assigned to test what factors could influence representation of trait composition by pollen data: the use of correction factors, the separation of vegetation and pollen data in to plant functional type and pollination mode.
 Lastly, the vegetation was aggregated to genus level and family level to test the effect of taxonomic resolution on the reconstructions.
@@
 
-![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Pollen_uncorrected.png)
-
-![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Pollen_corrected.png)
-
 ### Pollination mode
-
-![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Wind_pollinated.png)
-
-![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/NotWind_pollinated.png)
 
@@
 
-![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/TRSH.png)
-
-![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/HERB.png)
-
 ### Effect of taxonomic resolution
-
-![](Figures/veg_genus.png)
-
-## ![](Figures/veg-Family.png)
 
@@
 
+    -   vegetation surveys - all species versus max nine
+
+    -   estimation of the zones
+
 -   Some pollen types consider many more species than others for example in this study XX .
-This is also indentified as a problem in plant richness studies (Reitalu 2019) and trait rep (Carvalho, Blaus).
+This is also identified as a problem in plant richness studies (Reitalu 2019) and trait rep (Carvalho, Blaus).
 although bayesian xxx
@@
 -   size of the relevant source area (Bunting 2003)
+
+    -   The spatial resolution of a pollen assemblage is difficult to assess and is dependent on several factors, such as differences in taxon-specific differences in pollen productivity and dispersal and landscape structure
+
+    -   here we surveyed sites with different openness.
+
+-   pollen based trait reconstructions have been shown to correctly reflect environmental gradients as elevation (Sande) and and climatic variations (Barboni)
+
+## References
+
+    <div id="refs"></div>
 
@@
 
-```{r fig.height=6, fig.width=10, fig.cap="Scatter plot of vegetation basal area % against pollen abundance %"}
+```{r echo=FALSE, fig.height=6, fig.width=10}
 source("03_Vegetation_pollen_comparison.R") # pollen plot
@@
 
-```{r fig.height=6, fig.width=10, fig.cap="Scatter plot of vegetation basal area % against adjusted pollen abundance %"}
-
+```{r echo=FALSE, fig.height=6, fig.width=10}
 source("03_Vegetation_pollen_comparison.R") # pollen plot
 pol_veg_plot_adj
-
 ```
@@
 
-```{r}
+```{r echo=FALSE}
 # source("03_pollen_diagrams.R")
-```
-
-```{r}
-source("01_PFT_table_vegetation.R")
-sum
 ```
@@
 
-```{r}
-source("02_Prepare_trait_data_gapfilled.R")
+```{r echo=FALSE}
+sum_trait <- readRDS("RDS_files/01_summary_table_TRY_data.rds")
 kable(sum_trait)
@@
 
-```{r}
-source("01_Prepare_trait_data.R", encoding = 'UTF-8', echo=TRUE)
+```{r echo=FALSE}
+source("01_Prepare_trait_data_vegetation.R",encoding ="UTF-8")
 kable(sum_ph)
@@
 ```
-
-## References

### 2022-09-20 03_Vegetation_pollen_comparison.R
fix labeling issue

@@ -107,30 +107,8 @@ zoneC_swiss <- bdm_abun_bc %>%
 veg <- bind_rows(zoneA, zoneA_swiss, zoneB, zoneB_swiss, zoneC, zoneC_swiss) %>% 
   left_join(dfPOL, by = c("country","sitename","pollentaxon")) 
 
-# Add pollination mode and plant functional type
-polmode <- polmode %>% 
-  mutate(pollentaxon = dplyr::recode(pollentaxon,  
-                              "Apiaceae undiff." = "Apiaceae",
-                              "Caryophyllaceae undiff." = "Caryophyllaceae",
-                              "Cyperaceae undiff." = "Cyperaceae",
-                              "Ericacea-type" = "Ericales (tetrad)",
-                              "Fraxinus excelsior" = "Fraxinus",
-                              "Geranium" = "Geraniaceae",
-                              "Juniperus" = "Juniperus communis",
-                              "Pedicularis" = "Pedicularis palustris",
-                              "Ranunculuceae undiff." = "Ranunculaceae",
-                              "Rosaceae undiff." = "Rosaceae",
-                              "Viola palustris-type" = "Viola")) %>% 
-  add_row(pollentaxon = "Asteraceae", fam = "Asteraceae", growthform = "herb", pollination = "not wind") %>% 
-  add_row(pollentaxon = "Lamiaceae", fam = "Lamiaceae", growthform = "herb", pollination = "not wind") %>% 
-  add_row(pollentaxon = "Malvaceae", fam = "Malvaceae", growthform = "tree", pollination = "not wind") %>% 
-  add_row(pollentaxon = "Plantago", fam = "Plantaginaceae", growthform = "herb", pollination = "wind") %>% 
-  add_row(pollentaxon = "Liliaceae")  %>% 
-  add_row(pollentaxon = "Pteridophyte") %>% 
-  add_row(pollentaxon = "Tsuga ")
 
-
-## Calculate pollen representation value (Julier 2018, https://doi.org/10.1080/01916122.2017.1356392)
+## Calculate pollen representation pot (Julier 2018, https://doi.org/10.1080/01916122.2017.1356392)
 ## Plot pollen representation per zone ----
 zone_lab <- c("Inner ring (3.4 m)", "Middle ring (100 m)", "Outer ring (1km)")
 names(zone_lab) <- c("zoneA", "zoneB", "zoneC")
[…]

### 2022-10-07 01_PFT_table_pollen.R
tidied

@@ -19,8 +19,8 @@
 
 
 ## Load packages
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(readxl)) install.packages("readxl")
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(readxl)) install.packages("readxl")
 
 ## Load data
 dfPOL_Scot <- readRDS("RDS_files/01_Pollen_data_Scot.rds")

### 2022-10-07 04_Correlation_bayesian.R

@@ -81,9 +81,9 @@ dfCWM_pol <- pollen_files %>%
          sitename = str_remove(sitename, "X")
          ) %>% 
   dplyr::select(country, sitename, pollen_mean = Mean, 
-                pollen_sd = SD, trait, correction, pollination, growthform) %>% 
+                pollen_sd = SD, trait, correction, pollination, growthform) 
   # convert to original scale
-  mutate(across(c("pollen_mean", "pollen_sd"), ~exp(.)))
+  # mutate(across(c("pollen_mean", "pollen_sd"), ~exp(.)))
 
 dfCWM_veg <- veg_files %>% 
   folderpath.fun(.) %>% 
@@ -107,7 +107,7 @@ dfCWM_veg <- veg_files %>%
          !(pollination == "wind" & zone == "zoneA" & sitename %in% c("665174", "C009","C015")),
          !(pollination == "not wind" & zone == "zoneA" & sitename %in% c("563166"))) %>% 
   # convert to original scale
-  mutate(across(c("veg_mean", "veg_sd"), ~exp(.))) %>% 
+  # mutate(across(c("veg_mean", "veg_sd"), ~exp(.))) %>% 
   # combine to facilitate mapping
   mutate(zone = paste(country, zone)) %>% 
   filter(!sitename == "563186")
@@ -159,8 +159,9 @@ cor_bay <- function(cwm, selectedzone, selectedtrait, selectedtreatment){
         x_rand[i,1:2] ~ dmt(mu[], prec[,,i], df)
         }
       #data# N, CWM_mean, CWM_sd
-      #monitor# rho
+      #monitor# rho, df
     }"
+  
   
   inits_list <-  list(mu = c(min(CWM_mean)*0.1, max(CWM_mean)*10),
                     rho = cor(CWM_mean[, 1], CWM_mean[, 2]))
@@ -197,7 +198,7 @@ combo <- cwm %>%
   distinct()
 
 if (Sys.info()[4] == "LUIN67311") {  
-  plan(multisession, workers = 1)
+  plan(multisession, workers = 2)
 } else {
   plan(multisession, workers = 6)
 }
@@ -336,9 +337,12 @@ cwm_trsh <- dfCWM_pol %>%
          correction == "no correction") %>% 
   left_join(dfCWM_veg, by = c("country", "sitename", "trait", "growthform", "pollination")) %>% 
   filter(taxres == "stand.spec")
+combo_trsh <- cwm_trsh %>% 
+  dplyr::select(trait, zone) %>% 
[…]

### 2022-10-07 01_PFT_table_vegetation.R
add missing species NWSS data

@@ -18,8 +18,8 @@
 ## ---------------------------
 
 # Libraries
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(data.table)) install.packages("data.table")
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(data.table)) install.packages("data.table")
 
 # Pollination and PFT for vegetation----
 traits <- fread(
@@ -140,7 +140,19 @@ pft <- traits %>%
                              "Juncus bufonius", "Larix decidua", "Lathyrus palustris",
                              "Galium odoratum","Galium verum", "Geranium robertianum",
                              "Clematis vitalba", "Helianthemum salicifolium",
-                             "Rhinanthus glacialis", "Potentilla brevifolia"),
+                             "Rhinanthus glacialis", "Potentilla brevifolia",
+                             "Blechnum spicant","Sambucus nigra" ,"Abies grandis",  "Abies procera", 
+                             "Acer campestre"  , "Aesculus hippocastanum", "Alnus glutinosa","Alnus incana",
+                             "Betula pubescens", "Betula", "Botrychium lunaria",  "Carpinus betulus",
+                             "Castanea sativa", "Chamaecyparis lawsoniana", "Crataegus monogyna", "Cryptomeria japonica",
+                             "Cupressus", "Ilex aquifolium", "Larix kaempferi","Malus sylvestris",
+                             "Nothofagus", "Picea omorika", "Picea sitchensis", "Pinus contorta",
+                             "Pinus nigra", "Pinus", "Populus", "Populus tremula",
+                             "Prunus avium"   , "Prunus padus"  , "Pseudotsuga menziesii",
+                             "Quercus robur","Quercus rubra", "Quercus", "Salix caprea",
+                             "Salix cinerea","Sequoia sempervirens", "Sorbus aria", "Sorbus torminalis",
+                             "Taxus baccata","Thuja plicata", "Tilia", "Tsuga heterophylla",
+                             "Ulmus glabra","Ulmus procera"), 
           growthform = c("grass","grass", "grass", "grass", "herb",
                          "grass","herb", "shrub", "herb", "herb",
                          "fern", "herb", "herb", "herb", "fern",
@@ -153,9 +165,22 @@ pft <- traits %>%
                          "grass", "shrub","grass","grass", "herb",
                          "fern", "shrub", "grass", "tree", "herb",
[…]

### 2022-10-07 01_Prepare_TRY_data.R
tidied

@@ -21,10 +21,10 @@
 ## ---------------------------
 
 ## Load packages
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(data.table)) install.packages("data.table")
-if(!require(LCVP)) install.packages("LCVP")
-if(!require(lcvplants)) install.packages("lcvplants")
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(data.table)) install.packages("data.table")
+if (!require(LCVP)) install.packages("LCVP")
+if (!require(lcvplants)) install.packages("lcvplants")
 
 # increase memory limit
 memory.limit(999999)
@@ -115,78 +115,78 @@ deadID <- meta %>%
 #subset plant that were not measured in natural conditions
 #get list of observations that were not in natural settings
 setting <- meta %>%
-  mutate(drop = ifelse(OrigValueStr == "Botanical garden"|
-                         OrigValueStr == "botanical garden (Bergius Botanical Garden, Stockholm, Sweden)"|
-                         OrigValueStr == "Botanical gardens, greenhouses and other atypical habitats"|
-                         OrigValueStr == "Chamber"|
-                         OrigValueStr == "climate chamber"|
-                         OrigValueStr == "Climate chamber"|
-                         OrigValueStr == "Climate Chamber"|
-                         OrigValueStr == "Climate chamber, non-limiting conditions, (cf. dataset reference)"|
-                         OrigValueStr == "climate chambers"|
-                         OrigValueStr == "Common Garden"|
-                         OrigValueStr == "Controlled climate chamber"|
-                         OrigValueStr == "controlled environment room"|
-                         OrigValueStr == "drought treatment"|
-                         OrigValueStr == "FACE"|
-                         OrigValueStr == "FE"|
-                         OrigValueStr == "C"|
-                         OrigValueStr == "Field Experiment"|
-                         OrigValueStr == "FW"|
-                         OrigValueStr == "G"|
-                         OrigValueStr == "GH"|
-                         OrigValueStr == "Glasshouse"|
-                         OrigValueStr == "Greehouse"|
-                         OrigValueStr == "Green house"|
-                         OrigValueStr == "greenhouse"|
-                         OrigValueStr == "Greenhouse"|
-                         OrigValueStr == "Greenhouse, grrowth container"|
-                         OrigValueStr == "groth chamber"|
-                         OrigValueStr == "growth-chamber"|
-                         OrigValueStr == "growth chamber"|
[…]

### 2022-10-07 01_Prepare_pollen_data.R
tidied

@@ -17,14 +17,12 @@
 ##
 ## ---------------------------
 
-options(scipen=999)
-
-## Load packages
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(readxl)) install.packages("readxl")
-
-# Seed 
+options(scipen = 999)
 set.seed(1)
+## Load packages
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(readxl)) install.packages("readxl")
+
 
 # Scotland ----
 pollen_raw <- read.csv("Data/Pollen_count_Scotland-checked.csv", skip = 1, sep = ";")

### 2022-10-07 01_Prepare_trait_data_vegetation.R
make summary table more readable

@@ -123,15 +123,17 @@ cn <- cn_raw %>%
   dplyr::select(-measurementID)
 saveRDS(cn, "RDS_files/01_CN.rds")
 
-## SUmmary table per pollentaxon
+## Summary table per pollentaxon
+bdm_traits <- readRDS("RDS_files/01_Traits_Swiss.rds")
 spec <- readRDS("RDS_files/02_PollenType_species.rds")
 
-sum_ph <- plantheight %>% 
-  left_join(spec, by = c("species" = "stand.spec")) %>% 
-  group_by(pollentaxon) %>% 
+sum_ph <- plantheight[,c("stand.spec", "PlantHeight")] %>% 
+  bind_rows(la[,c("stand.spec", "LA")]) %>% 
+  bind_rows(sla[,c("stand.spec", "SLA")]) %>% 
+  group_by(stand.spec) %>% 
   summarise(nspec = length(unique(stand.spec)),
-            nobs = n(),
-            PlantHeight = paste(round(mean(PlantHeight), 1), "+-", round(sd(PlantHeight), 1)))
+            nobs = across(c(SLA,LA,PlantHeight),~sum(!is.na(.))))
+
 sum_la <- la %>% 
   left_join(spec, by = c("species" = "stand.spec")) %>% 
   group_by(pollentaxon) %>% 
@@ -145,4 +147,3 @@ sum_sla <- sla %>%
             nobs = n(),
             PlantHeight = paste(round(mean(SLA), 1), "+-", round(sd(SLA), 1)))
 
-rm(list=setdiff(ls(), c("sum_ph", "sum_la", "sum_sla")))

### 2022-10-07 01_Species_abundance.Rmd
add NWSS data

@@
 Use Sample raster values to compute landcover types for vegetation
-plots. Export as Veg_type_zoneB_sites.csv and Veg_type_zoneA_sites.csv
+plots. Export as Veg_type_zoneB_sites.csv and Veg_type_zoneA_sites.csv,
 
-## Prepare data
+Canopy data for zoneB was extracted for the zone B sites from the
+Scottish Native Woodland survey:
+<https://forestry.gov.scot/forests-environment/biodiversity/native-woodlands/native-woodland-survey-of-scotland-nwss>
+
+Patterson, G., Nelson, D., Robertson, P. and Tullis, J., 2014.
+*Scotland's Native Woodlands: Results from the Native Woodland Survey of
+Scotland*. [online] Edinburgh: Forestry Commission Scotland. Available
+at:
+<https://forestry.gov.scot/publications/74-scotland-s-native-woodlands-results-from-the-native-woodland-survey-of-scotland>
+[Accessed 3 October 2022].
+
+-   Join NWSS woodland cover shape file with Species Structures table -
+    by SCPTDATA_I
+
+-   Union zone buffer (B or C separately) and NWSS woodland vector layer
+
+-   remove all that don't overlap with the zone buffer
+
+-   calculate area of each polygon with field calculator \$area
+
+-   export to csv
+
+## Zone A
+
+```{r}
+# load data
+spec_a_raw <- read_xlsx("Data/Species_zone_A.xlsx", skip = 1)
+canop_a_raw <- read_xlsx("Data/Canopy_zone_A.xlsx")
+
+canop_a <- canop_a_raw %>% 
+  pivot_longer(-sitename, names_to = "species", values_to = "abun") %>% 
+  mutate(direction = "Canopy", distance = "0-10 meter")
+# Make tidy
+spec_a <- spec_a_raw %>% 
+  pivot_longer(3:19, names_to = "plot", values_to = "abun") %>% 
+  rename(sitename = Site, species = Species) %>% 
+  # remove moss
+  filter(!species %in% c("Moss", "Sphagnum", "Polytrichum commune",
+                         "Sphagnum squarrosum")) %>% 
+  # create direction and distance column
+  mutate(direction = case_when(str_detect(plot, "N") ~ "North",
+                               str_detect(plot, "S") ~ "South",
+                               str_detect(plot, "E") ~ "East",
+                               str_detect(plot, "W") ~ "West",
+                               str_detect(plot, "NE") ~ "Northeast",
+                               str_detect(plot, "NW") ~ "Northwest",
+                               str_detect(plot, "SW") ~ "Southwest",
+                               str_detect(plot, "SE") ~ "Southeast",
+                               plot == "Center" ~ "Center"),
+         distance = case_when(plot %in% c("N...4", "E...5", "S...6","W...7") ~ 
+                                "1.5-3 meter",
+                              plot %in% c("N...8", "E...9", "S...10", "W...11")  ~ 
+                                "3-6 meter",
+                              plot %in% c("N...12","NE","SE","NW","SW", "E...14", "S...16", "W...18") ~ 
+                                "6-10 meter",
+                              plot == "Center" ~ "0 meter")) %>% 
+  bind_rows(canop_a) %>% 
+  # create unique id for plot in site
+  group_by(sitename, direction, distance) %>% 
+  mutate(plot = cur_group_id())
+
+```
+
+## Zone B & C
 
@@
 
-```{r load and prepare vegetation data}
+```{r load and prepare species data}
 # Load in vegetation type data for the plots
 zone_b_site <- read.csv("Data/20220106-Veg_type_zoneB_sites.csv")
+spec_b_raw <- read_xlsx("Data/species_zone_B.xlsx") # load in species data
 
-# Make tidy
+# Make vegetation type tidy
 zone_b_site <- zone_b_site %>% 
@@
                            "9" = "E5: Woodland fringes and clearings and tall forb stands",
-                           "10"= "F2: Arctic, alpine and subalpine scrub",
+                           "10" = "F2: Arctic, alpine and subalpine scrub",
                            "11"  = "F3: Temperate and mediterranean-montane scrub",
-                           "12"= "F4: Temperate shrub heathland",
+                           "12" = "F4: Temperate shrub heathland",
                            "13" = "F9: Riverine and fen scrubs",
-                           "14"= "G1: Broadleaved deciduous woodland",
+                           "14" = "G1: Broadleaved deciduous woodland",
                            "15" = "G3: Coniferous woodland",
@@
 
-# load in vegetation data
-spec_b_raw <- read_xlsx("Data/species_zone_B.xlsx")
-
-# make tidy
+# Tidy and join 
 spec_b <- spec_b_raw %>% 
@@
   # join with vegetation type data
-  left_join(zone_b_site, by = c("sitename", "plot"))
+  left_join(zone_b_site, by = c("sitename", "plot")) %>% 
+  # filter trees from species survey (will be covered by NWSS)
+  filter(!species %in% c("Betula pendula", "Pinus sylvestris", 
+                        "Juniperus communis", "Salix sp.",
+                        "Larix decidua", "Salix repens"))
 ```
 
-## Zone B & C
+tree abundance
+
+```{r}
+canop_b_raw <- 
+  read.csv("Data/NWSS_zoneB_ID_forest_cover.csv") %>% 
+  # clean column names
+  rename_with(~str_remove(., "NWSS_Species_Structures_5718542285393555851.â...NWSS_Species_Structures_"),
+              matches("NWSS_Species_Structures_5718542285393555851.â...NWSS_Species_Structures_"))
+canop_c_raw <- 
+  read.csv("Data/NWSS_zoneC_ID_forest_cover.csv") %>% 
+  # clean column names
+  rename_with(~str_remove(., "NWSS_Species_Structures_5718542285393555851.â...NWSS_Species_Structures_"),
+              matches("NWSS_Species_Structures_5718542285393555851.â...NWSS_Species_Structures_"))
+NWSS_sp <- 
+  read.csv("Data/NWSS_Species_Structures.csv") 
+NWSS <- 
+  read.csv("Data/NWSS_basis.csv") %>%  # inventory of NWSS 
+  # clean column names
+  rename_with(~str_remove(., "NWSS_Species_Structures_"),
+              matches("NWSS_Species_Structures_"))
+
+# calculate average species abundance of species per habitat type
+# species percentage of NWSS (not available for all polygon)
+habitats <- canop_c_raw$DOM_HABITA %>% unique
+sp_habitat <- NWSS %>% 
+  filter(DOM_HABITA %in% habitats) %>% 
+  group_by(DOM_HABITA, SPECIES) %>% 
+  summarise(mean_pct = mean(SP_STR_PCT))
+
+# - calculate woodland cover for every site
+# - weight woodland cover by mean abundance of species per type
+canop_b <- canop_b_raw %>% 
+  # select relevant columns
+  dplyr::select(sitename = name, SCPTDATA_I, TYPE, DOM_HABITA, woodland_area = Woodland_area) %>% 
+  # fill in area without woodland
+  mutate(TYPE = if_else(TYPE == "", "Open land habitat", TYPE),
+         # rename site names
+         sitename = str_remove(sitename, "-A")) %>% 
+  # calculate woodland type percentage
+  group_by(sitename) %>% 
+  # calculate percentages
+  mutate(total_area = sum(woodland_area)) %>% 
+  mutate(woodland_percentage = woodland_area/total_area) %>% 
+  # join with species abundance
+  left_join(sp_habitat, by = "DOM_HABITA") %>% 
+  # create column with scientific species name
+  mutate(species = recode(SPECIES, "Common alder*" = "Alnus glutinosa",
+                          "Rowan *" = "Sorbus aucuparia",
+                          "Sitka spruce" = "Picea sitchensis",
+                          "European larch" = "Larix decidua",
+                          "Lodgepole pine" = "Pinus contorta",
+                          "Downy birch *" = "Betula pubescens",
+                          "Scots pine *" = "Pinus sylvestris",
+                          "Hazel *" = "Corylus avellana",
+                          "Eared willow *+" = "Salix aurita", 
+                          "Grey willow *" = "Salix cinerea", 
+                          "Hybrid larch" = "Larix decidua", 
+                          "Shrubs - preferentially browsed +" = "Unidentified",
+                          "Dwarf juniper *+" = "Juniperus communis", 
+                          "Silver birch *" = "Betula pendula", 
+                          "Beech" = "Fagus sylvatica", 
+                          "Lawsons cypress" = "Chamaecyparis lawsoniana", 
+                          "Common alder*" = "Alnus glutinosa", 
+                          "Douglas fir" = "Pseudotsuga menziesii", 
+                          "Yew *" = "Taxus baccata", 
+                          "Upright juniper * +" = "Juniperus communis", 
+                          "Norway spruce" = "Picea abies", 
+                          "Bird cherry *" = "Prunus padus", 
+                          "Intermediate oak *" = "Quercus sp.", 
+                          "Sessile oak *" = "Quercus petraea", 
+                          "Ash *" = "Fraxinus excelsior", 
+                          "Holly *" = "Ilex aquifolium", 
+                          "Aspen *" = "Populus tremula", 
+                          "Goat willow*" = "Salix caprea", 
+                          "Elder *" = "Sambucus nigra", 
+                          "Broom *+" = "Cytisus scoparius", 
+                          "Other non-native trees/shrubs" = "Unidentified", 
+                          "Sycamore" = "Acer pseudoplatanus", 
+                          "Grand fir" = "Abies grandis", 
+                          "Gorse *+" = "Ulex europaeus", 
+                          "Gean wild cherry *" = "Prunus avium",
+                          "Austrian pine" = "Pinus nigra",
+                          "Blackthorn *+" = "Prunus spinosa",
+                          "Birch *" = "Betula sp.",
+                          "Coast redwood" = "Sequoia sempervirens",
+                          "Bramble *+" = "Rubus fruticosus",
+                          "Crab apple" = "Malus sylvestris",
+                          "Dog rose *+" = "Rosa canina",
+                          "Geulder rose *+" = "Viburnum opulus",
+                          "Grey alder" = "Alnus incana",
+                          "Hawthorn *" = "Crataegus monogyna",
+                          "Field maple" = "Acer campestris",
+                          "Hornbeam" = "Carpinus betulus",               
+                          "Horse chestnut" = "Aesculus hippocastanum",
+                          "Japanese knotweed +" = "Reynoutria japonica",
+                          "Japanese larch" = "Larix kaempferi",
+                          "Leyland cypress" = "Cupressus sp.",        
+                          "Lime all" = "Tilia sp.",
+                          "Noble fir" = "Abies procera",
+                          "Non-native willow" = "Salix sp.",
+                          "Norway maple" = "Acer platanoides",       
+                          "Nothofagus all" = "Nothofagus sp.",
+                          "Omorika spruce" = "Picea omorika",         
+                          "Other native willows *" = "Salix sp.",
+                          "Other pines" = "Pinus sp.",
+                          "Other whitebeam"  = "Sorbus aria",       
+                          "Pedunculate oak *" = "Quercus robur",     
+                          "Poplar" = "Populus sp.",
+                          "Red oak" = "Quercus rubra",               
+                          "Rhododendron ponticum +" = "Rhododendron ponticum",
+                          "Scots pine" = "Pinus sylvestris",
+                          "Silver fir" = "Abies alba",
+                          "Snowberry+" = "Symphoricarpos sp.",
+                          "Sweet chestnut" = "Castanea sativa",
+                          "Western hemlock" = "Tsuga heterophylla",
+                          "Western red cedar" = "Thuja plicata",
+                          "Wild service tree" = "Sorbus torminalis",      
+                          "Wych elm *" = "Ulmus glabra",
+                          "English elm" = "Ulmus procera",
+                          "Japanese cedar" = "Cryptomeria japonica",
+                          "Montane willow all *+" = "Salix sp."
+                          )) %>% 
+  # filter non trees
+  mutate(growthform = case_when(str_detect(SPECIES, "\\+") ~ "shrub",
+                                SPECIES %in% c("Japanese knotweed +") ~ "herb",
+                                TRUE ~ "tree")) %>% 
+  filter(growthform == "tree",
+         !species %in% c("Unidentified","")) %>% 
+  # weight species abundance by woodland percentage
+  group_by(sitename, species) %>% 
+  mutate(sp_abun = mean_pct * woodland_percentage) %>% 
+  summarise(spec_abun_b = sum(sp_abun))
+
+# PAWS =  Planted Woods on Ancient Woodland Sites
+canop_c <- canop_c_raw %>% 
+  # select relevant columns
+  dplyr::select(sitename = name, SCPTDATA_I, TYPE, DOM_HABITA, woodland_area = Woodland_area) %>% 
+  # fill in area without woodland
+  mutate(TYPE = if_else(TYPE == "", "Open land habitat", TYPE),
+         # rename site names
+         sitename = str_remove(sitename, "-A")) %>% 
+  # calculate woodland type percentage
+  group_by(sitename) %>% 
+  # calculate percentages
+  mutate(total_area = sum(woodland_area)) %>% 
+  mutate(woodland_percentage = woodland_area/total_area) %>% 
+  # join with species abundance
+  left_join(sp_habitat, by = "DOM_HABITA") %>% 
+  # create column with scientific species name
+  mutate(species = recode(SPECIES, "Common alder*" = "Alnus glutinosa",
+                          "Rowan *" = "Sorbus aucuparia",
+                          "Sitka spruce" = "Picea sitchensis",
+                          "European larch" = "Larix decidua",
+                          "Lodgepole pine" = "Pinus contorta",
+                          "Downy birch *" = "Betula pubescens",
+                          "Scots pine *" = "Pinus sylvestris",
+                          "Hazel *" = "Corylus avellana",
+                          "Eared willow *+" = "Salix aurita", 
+                          "Grey willow *" = "Salix cinerea", 
+                          "Hybrid larch" = "Larix decidua", 
+                          "Shrubs - preferentially browsed +" = "Unidentified",
+                          "Dwarf juniper *+" = "Juniperus communis", 
+                          "Silver birch *" = "Betula pendula", 
+                          "Beech" = "Fagus sylvatica", 
+                          "Lawsons cypress" = "Chamaecyparis lawsoniana", 
+                          "Common alder*" = "Alnus glutinosa", 
+                          "Douglas fir" = "Pseudotsuga menziesii", 
+                          "Yew *" = "Taxus baccata", 
+                          "Upright juniper * +" = "Juniperus communis", 
+                          "Norway spruce" = "Picea abies", 
+                          "Bird cherry *" = "Prunus padus", 
+                          "Intermediate oak *" = "Quercus sp.", 
+                          "Sessile oak *" = "Quercus petraea", 
+                          "Ash *" = "Fraxinus excelsior", 
+                          "Holly *" = "Ilex aquifolium", 
+                          "Aspen *" = "Populus tremula", 
+                          "Goat willow*" = "Salix caprea", 
+                          "Elder *" = "Sambucus nigra", 
+                          "Broom *+" = "Cytisus scoparius", 
+                          "Other non-native trees/shrubs" = "Unidentified", 
+                          "Sycamore" = "Acer pseudoplatanus", 
+                          "Grand fir" = "Abies grandis", 
+                          "Gorse *+" = "Ulex europaeus", 
+                          "Gean wild cherry *" = "Prunus avium",
+                          "Austrian pine" = "Pinus nigra",
+                          "Blackthorn *+" = "Prunus spinosa",
+                          "Birch *" = "Betula sp.",
+                          "Coast redwood" = "Sequoia sempervirens",
+                          "Bramble *+" = "Rubus fruticosus",
+                          "Crab apple" = "Malus sylvestris",
+                          "Dog rose *+" = "Rosa canina",
+                          "Geulder rose *+" = "Viburnum opulus",
+                          "Grey alder" = "Alnus incana",
+                          "Hawthorn *" = "Crataegus monogyna",
+                          "Field maple" = "Acer campestris",
+                          "Hornbeam" = "Carpinus betulsu",               
+                          "Horse chestnut" = "Aesculus hippocastanum",
+                          "Japanese knotweed +" = "Reynoutria japonica",
+                          "Japanese larch" = "Larix kaempferi",
+                          "Leyland cypress" = "Cupressus sp.",        
+                          "Lime all" = "Tilia sp.",
+                          "Noble fir" = "Abies procera",
+                          "Non-native willow" = "Salix sp.",
+                          "Norway maple" = "Acer platanoides",       
+                          "Nothofagus all" = "Nothofagus sp.",
+                          "Omorika spruce" = "Picea omorika",         
+                          "Other native willows *" = "Salix sp.",
+                          "Other pines" = "Pinus sp.",
+                          "Other whitebeam"  = "Sorbus aria",       
+                          "Pedunculate oak *" = "Quercus robur",     
+                          "Poplar" = "Populus sp.",
+                          "Red oak" = "Quercus rubra",               
+                          "Rhododendron ponticum +" = "Rhododendron ponticum",
+                          "Scots pine" = "Pinus sylvestris",
+                          "Silver fir" = "Abies alba",
+                          "Snowberry+" = "Symphoricarpos sp.",
+                          "Sweet chestnut" = "Castanea sativa",
+                          "Western hemlock" = "Tsuga heterophylla",
+                          "Western red cedar" = "Thuja plicata",
+                          "Wild service tree" = "Sorbus torminalis",      
+                          "Wych elm *" = "Ulmus glabra",
+                          "English elm" = "Ulmus procera",
+                          "Japanese cedar" = "Cryptomeria japonica",
+                          "Montane willow all *+" = "Salix sp.",
+                          "Arran whitebeam *" = "Sorbus arranensis"
+         )) %>% 
+  # filter non trees
+  mutate(growthform = case_when(str_detect(SPECIES, "\\+") ~ "shrub",
+                                SPECIES %in% c("Japanese knotweed +") ~ "herb",
+                                TRUE ~ "tree")) %>% 
+  filter(growthform == "tree",
+       !species %in% c("Unidentified","")) %>% 
+  # weight species abundance by woodland percentage
+  group_by(sitename, species) %>% 
+  mutate(sp_abun = mean_pct * woodland_percentage) %>% 
+  summarise(spec_abun_c = sum(sp_abun))
+
+# join df's from both zones
+canop <- canop_b %>% 
+  left_join(canop_c, by = c("sitename","species"))
+```
 
@@
                          "Sphagnum squarrosum")) %>% 
-  # calculate avarage species cover per vegetation type using the plots belonging to that vegetation type
+  # calculate average species cover per vegetation type using the plots belonging to that vegetation type
   group_by(sitename, veg_type, species) %>% 
@@
   # join with vegetation cover data
-  full_join(veg_type, by = c("sitename", "veg_type")) %>% 
-  # weight species abundance by veg cover
-  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c) %>% 
-  # drop unneeded columns
-  dplyr::select(sitename, veg_type, species, veg_cov_b, veg_cov_c,
-                spec_abun_b,spec_abun_c)
+  full_join(veg_type, by = c("sitename", "veg_type"))
 
@@
   summarise(cov_b = round(sum(veg_cov_b), 2), cov_c = round(sum(veg_cov_c), 2)) %>%
-  filter(!cov_b == 0|!cov_c == 0)
+  filter(!cov_b == 0 | !cov_c == 0)
 
@@
   summarise(abun = mean(abun))
-
 veg_cover_bc_na <- veg_cover_bc %>% 
   filter(is.na(species)) %>%
-  dplyr::select(-species) %>% 
-  left_join(missing, by = "veg_type") %>% 
-  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c) %>% 
-  dplyr::select(sitename, veg_type, species, veg_cov_b, veg_cov_c, spec_abun_b,
-         spec_abun_c)
+  dplyr::select(-species, -abun) %>% 
+  left_join(missing, by = "veg_type")
 
@@
   filter(!is.na(species)) %>% 
+  # bind with df with missing veg types
   bind_rows(., veg_cover_bc_na) %>% 
-  arrange(sitename, veg_type, species)
-```
-
-## Zone A
-
-```{r}
-# load data
-spec_a_raw <- read_xlsx("Data/Species_zone_A.xlsx", skip = 1)
-canop_a_raw <- read_xlsx("Data/Canopy_zone_A.xlsx")
-
-canop_a <- canop_a_raw %>% 
-  pivot_longer(-sitename, names_to = "species", values_to = "abun") %>% 
-  mutate(direction = "Canopy", distance = "0-10 meter")
-# Make tidy
-spec_a <- spec_a_raw %>% 
-  pivot_longer(3:19, names_to = "plot", values_to = "abun") %>% 
-  rename(sitename = Site, species = Species) %>% 
-  # remove moss
-  filter(!species %in% c("Moss", "Sphagnum", "Polytrichum commune",
-                         "Sphagnum squarrosum")) %>% 
-  # create direction and distance column
-  mutate(direction = case_when(str_detect(plot, "N") ~ "North",
-                               str_detect(plot, "S") ~ "South",
-                               str_detect(plot, "E") ~ "East",
-                               str_detect(plot, "W") ~ "West",
-                               str_detect(plot, "NE") ~ "Northeast",
-                               str_detect(plot, "NW") ~ "Northwest",
-                               str_detect(plot, "SW") ~ "Southwest",
-                               str_detect(plot, "SE") ~ "Southeast",
-                               plot == "Center" ~ "Center"),
-         distance = case_when(plot %in% c("N...4", "E...5", "S...6","W...7") ~ 
-                                "1.5-3 meter",
-                              plot %in% c("N...8", "E...9", "S...10", "W...11")  ~ 
-                                "3-6 meter",
-                              plot %in% c("N...12","NE","SE","NW","SW", "E...14", "S...16", "W...18") ~ 
-                                "6-10 meter",
-                              plot == "Center" ~ "0 meter")) %>% 
-  bind_rows(canop_a) %>% 
-  # create unique id for plot in site
-  group_by(sitename, direction, distance) %>% 
-  mutate(plot = cur_group_id())
-
+  # weight species abundance by veg cover
+  group_by(sitename, species) %>% 
+  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c) %>%
+  summarise(spec_abun_b = sum(spec_abun_b),
+            spec_abun_c = sum(spec_abun_b)) %>% 
+  # bind with tree species data
+  bind_rows(canop) %>% 
+  arrange(sitename, species)
 ```
@@
 ```{r standardize species names}
-# load in generalized species list, spelling checked and harmonized with
-# LCVP
+# load in generalized species list, spelling checked and harmonized with LCVP
 harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx") 
@@
   mutate(stand.spec = str_remove(stand.spec, pattern = " sp\\."))
+
 spec_a <- spec_a %>% 
@@
 veg_cover_bc <- veg_cover_bc %>% 
-  left_join(harm, by = "species")
+  left_join(harm, by = "species") %>% 
+  filter(!is.na(stand.spec))
 
@@
             cov_c = round(sum(veg_cov_c), 2)) %>%
-  filter(!cov_b == 0|!cov_c == 0)
+  filter(!cov_b == 0 |!cov_c == 0)
 

### 2022-10-07 04_Correlation_bayesian.R
add function for plotting

@@ -28,7 +28,8 @@ if (!require(tidyverse)) install.packages("tidyverse")
 if (!require(rjags)) install.packages("rjags")
 if (!require(runjags)) install.packages("runjags")
 if (!require(furrr)) install.packages("furrr")
-  
+if (!require(broom)) install.packages("broom")  
+
 # load PFT summary table
 # filter sites with to little vegetation cover of particular type
 sum <- readRDS("RDS_files/01_Percentage_cover_pft.rds")
@@ -44,8 +45,9 @@ traits <- c("LA","SLA","PlantHeight")
 pollen_files <- list.files("RDS_files/") %>% 
   str_subset(pattern = "03_CWM_estimates_") %>% 
   str_subset(paste(traits, collapse = "|")) %>% 
-  str_subset(pattern = "pollen") 
-  
+  str_subset(pattern = "pollen") %>% 
+  str_subset(pattern = "gapfilled", negate = TRUE)
+
 pollen_lab <- pollen_files %>% 
   str_subset(paste(traits, collapse = "|")) %>% 
   str_remove(., "03_CWM_estimates_") %>% 
@@ -54,7 +56,10 @@ pollen_lab <- pollen_files %>%
 veg_files <- list.files("RDS_files/") %>% 
   str_subset(pattern = "03_CWM_estimates_") %>% 
   str_subset(paste(traits, collapse = "|")) %>% 
-  str_subset(pattern = "pollen", negate = TRUE)
+  str_subset(pattern = "pollen", negate = TRUE) %>% 
+  str_subset(pattern = "gapfilled", negate = TRUE) %>% 
+  str_subset(pattern = "TRY", negate = TRUE)
+
 veg_lab <- veg_files %>% 
   str_subset(paste(traits, collapse = "|")) %>% 
   str_remove(., "03_CWM_estimates_") %>% 
@@ -81,9 +86,10 @@ dfCWM_pol <- pollen_files %>%
          sitename = str_remove(sitename, "X")
          ) %>% 
   dplyr::select(country, sitename, pollen_mean = Mean, 
-                pollen_sd = SD, trait, correction, pollination, growthform) 
+                pollen_sd = SD, trait, correction, pollination, growthform) %>% 
   # convert to original scale
-  # mutate(across(c("pollen_mean", "pollen_sd"), ~exp(.)))
+  mutate(across(c("pollen_mean", "pollen_sd"), ~exp(.))) %>% 
+  filter(country == "Scotland") 
 
 dfCWM_veg <- veg_files %>% 
[…]

### 2022-10-07 05_Scatterplots.R
remove swiss data, fix range on scales

@@ -13,7 +13,7 @@
 ## ---------------------------
 ##
 ## Notes:
-##   
+##
 ##
 ## ---------------------------
 
@@ -22,35 +22,164 @@
 if (!require(tidyverse)) install.packages("tidyverse")
 
 ## Load data
-dfCWM_veg <- readRDS("RDS_files/04_CWM_estimates_vegetation.rds")
-dfCWM_pol <- readRDS("RDS_files/04_CWM_estimates_pollen.rds") %>% 
-  filter(!str_detect(correction,pattern = "draw"))
+dfCWM_veg <- readRDS("RDS_files/04_CWM_estimates_vegetation_Scotland.rds") %>%
+  mutate(across(c("veg_mean", "veg_sd"),~exp(.)))
+dfCWM_pol <- readRDS("RDS_files/04_CWM_estimates_pollen_Scotland.rds") %>%
+  filter(!str_detect(correction,pattern = "draw")) %>%
+  mutate(across(c("pollen_mean", "pollen_sd"),~exp(.)))
 
 # join data
-dfCWM <- dfCWM_veg %>% 
-  left_join(dfCWM_pol, by = c("country","sitename", "trait", "growthform","pollination")) %>% 
+dfCWM <- dfCWM_veg %>%
+  left_join(dfCWM_pol, by = c("country","sitename", "trait", "growthform","pollination")) %>%
   # create treatment column
   mutate(treatment = case_when((pollination == "allpol" & growthform == "allpft" &
                                   taxres == "stand.spec") ~ "Correction factor",
                                taxres %in% c("family","genus") ~ "Taxonomic resolution",
                                (pollination != "allpol" & growthform == "allpft") ~ "Pollination mode",
                                (growthform != "allpft" & pollination == "allpol") ~ "Growth form"),
-         zone = str_remove(zone, pattern = "Scotland |Switzerland "))
+         zone = str_remove(zone, pattern = "Scotland |Switzerland ")) %>%
+  filter(country == "Scotland")
 
-# Produce plots
-labs_trait <- c(PlantHeight = "Height (cm)",
-          LA = "Leaf area (mm2)",
-          SLA = "SLA (mg/mm2)")
+# Plot labels
+labs_trait <- c("Height (cm)",
+          "Leaf area (mm2)",
+          "SLA (mm2/mg)")
+names(labs_trait) <- c("PlantHeight", "LA", "SLA")
 labs_zone <- c(zoneA = "",
[…]

### 2022-11-10 01_PFT_table_vegetation.R
reduced repetition in code

@@ -261,92 +261,54 @@ dfPFT <- full_join(pft, polmode, by = "AccSpeciesName") %>%
 saveRDS(dfPFT, "RDS_files/Polmode_pft_vegetation.rds")
 
 # Make summary table per site -----
+dfPFT <- readRDS("RDS_files/Polmode_pft_vegetation.rds")
 zoneA <- dfABUN_a %>%
-  arrange(sitename, stand.spec) %>% 
-  # join with pft and polmode data
-  left_join(dfPFT, 
-            by = c("stand.spec" = "AccSpeciesName")) %>% 
   rename(family = fam) %>% 
   ungroup %>% 
   mutate(zone = "zoneA",
          country = "Scotland")
 zoneB <- dfABUN_bc %>% 
-  # join with pft and polmode data
-  left_join(dfPFT, 
-            by = c("stand.spec" = "AccSpeciesName")) %>% 
   ungroup() %>% 
   dplyr::select(sitename, stand.spec, genus, family = fam,
-                abun = spec_abun_b, growthform, polmode) %>% 
+                abun = spec_abun_b) %>% 
   mutate(zone = "zoneB",
          country = "Scotland")
 zoneC <- dfABUN_bc %>% 
-  # join with pft and polmode data
-  left_join(dfPFT, 
-            by = c("stand.spec" = "AccSpeciesName")) %>% 
   ungroup %>% 
   dplyr::select(sitename, stand.spec, genus, family = fam, 
-                abun = spec_abun_c, growthform, polmode) %>% 
+                abun = spec_abun_c) %>% 
   mutate(zone = "zoneC",
          country = "Scotland")
 bdm_zoneA <- bdm_abun %>%
-  # join with pft and polmode data
-  left_join(dfPFT, 
-            by = c("stand.spec" = "AccSpeciesName")) %>% 
-  dplyr::select(sitename, stand.spec, genus, family, abun,
-                growthform, polmode) %>%
+  dplyr::select(sitename, stand.spec, genus, family, abun) %>%
   mutate(sitename = as.character(sitename)) %>%
   mutate(zone = "zoneA",
          country = "Switzerland")
-bdm_abunbc <- readRDS("RDS_files/01_Species_abundance_zoneBC_swiss.rds")
-bdm_zoneB <- bdm_abunbc %>%
-  # join with pft and polmode data
-  left_join(dfPFT, 
[…]

### 2022-11-10 01_Prepare_pollen_data.R
correct misnamed site

@@ -136,7 +136,8 @@ dfPOL <- read.csv("Data/Pollen_count_Switserland-checked.csv", skip = 1, sep = "
            str_remove(pollentaxon, "-type") %>% 
            # Typos
            recode(Caryophyllaccea = "Caryophyllaceae")
-  )
+  ) %>% 
+  mutate(sitename = recode(sitename, "X563186" = "X563166" )) # miss named site
 
 # percentage data with missing ppe's 
 missing_ppe_swiss <- dfPOL %>%
@@ -180,5 +181,3 @@ dfPOL <- dfPOL_nc %>%
   ungroup()
 
 saveRDS(dfPOL, "RDS_files/01_Pollen_data_Swiss.rds")
-
-rm(list=setdiff(ls(), c("missing_ppe_scot", "missing_ppe_swiss")))

### 2022-11-10 01_Prepare_trait_data_vegetation.R
change unit of LA

@@ -94,7 +94,9 @@ la <- sla_ldmc_raw %>%
   arrange(sitename, species, measurementID) %>% 
   # merge with species harmonisation list
   left_join(harm, by = "species") %>% 
-  filter(!is.na(LA)) # check notes from raw data file for reason why missing
+  filter(!is.na(LA)) %>%  # check notes from raw data file for reason why missing
+  # convert to cm2
+  mutate(LA = LA/100)
 saveRDS(la, "RDS_files/01_LA.rds")
     
 ## C ----

### 2022-11-10 01_Species_abundance.Rmd
fix joining issue

@@
 canop <- canop_b %>% 
-  left_join(canop_c, by = c("sitename","species"))
+  full_join(canop_c, by = c("sitename", "species")) %>% 
+  filter(!(is.na(spec_abun_b) & is.na(spec_abun_c))) %>% 
+  replace_na(list(spec_abun_b = 0,
+                  spec_abun_c = 0))
 ```

### 2022-11-14 01_Swiss_data_compilation.R
LA to cm2, species codes checked

@@ -19,9 +19,9 @@
 ## ---------------------------
 
 # Load packages
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(LCVP)) install.packages("LCVP")
-if(!require(lcvplants)) install.packages("lcvplants")
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(LCVP)) install.packages("LCVP")
+if (!require(lcvplants)) install.packages("lcvplants")
 
 # Load data
 bdm_meta <- read_csv("Data/bdm_metadata.csv") # contains plot codes, X/Y coordinates (on the Swiss grid), elevation, some basic climate data (MAT, TAP) and a habitat classification
@@ -38,9 +38,10 @@ plot_id <- bdm_pollen %>%
 
 # check sites 
 bdm_meta <- bdm_meta %>% 
-  filter(aID_STAO %in% plot_id)
- 
-# bdm_meta <- bdm_meta %>% 
+  filter(aID_STAO %in% plot_id) 
+
+# saveRDS(bdm_meta, "RDS_files/01_Meta_data_BDM.rds")
+# bdm_meta <- bdm_meta %>%
 #   filter(aID_STAO %in% bdm_sp$plot)
 # write.csv(bdm_meta, file = "Data/Swiss_LC/all_BDM_plots_coord.csv")
 
@@ -63,10 +64,15 @@ bdm_sp_codes <- bdm_sp_codes %>%
                         `vaccinium vitis idaea` = "Vaccinium vitis-idaea",
                         `Anthoxanthum odoratum aggr.` = "Anthoxanthum odoratum"),
          Species_code = recode(Species_code, 
-                               alcxan = "alcvul", # Alchemilla vulgaris current code alcxan is not in species or trait data, alcvul is
-                               soldal = "solalp" # Soldanella alpinia idem
-                               ) # soraur,lonneg, rubfru, giumon missing from codes 
-         ) 
+                               soldal = "solalp", # Soldanella alpinia idem
+                               soraur = "sorauc", # Sorbus aucuparia
+                               lonneg = "lonnig", # Lonicera nigra
+                               giumon = "geumon" # Geum montanum
+                               ) 
+         ) %>% 
+  # missing codes
+  add_row(Species_code = c("Rubfru", "Alcxan"),
+          Species = c("Rubus fruticosus", "Alchemilla xanthochlora"))
 
 ## Abundance ----
 # Join species codes with abundance data
[…]

### 2022-11-14 02_CWM_estimates_vegetation.R
reduce duplication in script

@@ -17,6 +17,9 @@
 ##
 ## ---------------------------
 
+set.seed(1)
+
+
 ## Load packages
 if (!require(tidyverse)) install.packages("tidyverse")
 if (!require(rjags))  install.packages("rjags")
@@ -25,7 +28,8 @@ if (!require(mcmc)) install.packages("mcmc")
 if (!require(coda)) install.packages("coda")
 if (!require(furrr))  install.packages("furrr")
 
-set.seed(1)
+## Load functions
+source("cwm_functions.R")
 
 ## Load and prepare data -----
 # Trait data ----
@@ -38,52 +42,55 @@ sla <- readRDS("RDS_files/01_SLA.rds") %>%
   ungroup()
 la <- readRDS("RDS_files/01_LA.rds") %>% 
   ungroup()
-c <- readRDS("RDS_files/01_CN.rds") %>% 
-  ungroup() %>% 
-  rename(LeafC = C)
-n <- readRDS("RDS_files/01_CN.rds") %>% 
-  ungroup() %>% 
-  rename(LeafN = N)
+
 
 lTrait <- list(PlantHeight = plantheight,
                LDMC = ldmc,
                SLA = sla,
-               LA = la,
-               LeafC = c,
-               LeafN = n)
+               LA = la)
 
-dfTRY <- readRDS("RDS_files/01_Clean_TRY_data.rds")
+dfTRY_raw <- readRDS("RDS_files/02_Gapfilled_traits.rds")
 
 # Pollination and PFT ----
 dfPFT <- readRDS("RDS_files/Polmode_pft_vegetation.rds")
 
-# Species data ----
+# Species to pollen table
+polspec <- readRDS("RDS_files/02_PollenType_species.rds") %>% 
+  dplyr::select(pollentaxon, stand.spec) %>% distinct() %>% as_tibble()
+
+# Species abundance data ----
 dfABUN_a <- readRDS("RDS_files/01_Species_abundance_a.rds") 
 dfABUN_bc <- readRDS("RDS_files/01_Species_abundance_bc.rds")
 
 zoneA <- dfABUN_a %>%
   arrange(sitename, stand.spec) %>% 
-  # select plots that are comparable with Switzerland
-  filter(distance %in% c("0 meter", "1.5-3 meter")) %>%
   # join with pft and polmode data
   left_join(dfPFT, 
             by = c("stand.spec" = "AccSpeciesName")) %>% 
   rename(family = fam) %>% 
[…]

### 2022-11-14 03_CWM_estimates_pollen.R

@@ -17,6 +17,7 @@
 ##
 ## ---------------------------
 
+set.seed(1)
 
 ## Load packages
 if (!require(tidyverse)) install.packages("tidyverse")
@@ -24,13 +25,13 @@ if (!require(rjags))  install.packages("rjags")
 if (!require(runjags)) install.packages("runjags")
 if (!require(mcmc)) install.packages("mcmc")
 if (!require(coda)) install.packages("coda")
-if (!require(furrr))  install.packages("furrr")
+if (!require(furrr)) install.packages("furrr")
 
-set.seed(1)
+source("cwm_functions.R")
 
 ## Load and prepare data -----
-# Trait data ----
-dfTRAIT <- readRDS("RDS_files/01_TRY_raw_traits.rds")
+# Trait data ---- 
+dfTRAIT <- readRDS("RDS_files/02_Gapfilled_traits_pollen.rds")
 
 # Pollen data ----
 dfPOL_Scot <- readRDS("RDS_files/01_pollen_data_Scot.rds")
@@ -40,515 +41,129 @@ dfPOL <- bind_rows("Scotland" = dfPOL_Scot, "Switzerland" =
   ungroup()
 
 selectedcountry <- "Scotland"
-selectedtrait <- "LA"
+selectedtrait <- "PlantHeight"
 selectedabun <- str_subset(colnames(dfPOL), "percent")[1]
 
 normaltraits <- c("PlantHeight","SLA","LA")
 
-# No subset ----
-if (0) {
-  cwmlogfunc <- function(selectedtrait, selectedcountry, selectedabun){
-    trait <- dfTRAIT %>% 
-      filter(country == selectedcountry) %>% 
-      # drop NA's and 0's
-      dplyr::select(pollentaxon, all_of(selectedtrait)) %>% 
-      mutate(across(where(is.double), ~na_if(.,0))) %>% 
-      drop_na() %>% 
-      # Sort alphabetically
-      arrange(pollentaxon) %>% 
-      # check number of observations
-      group_by(pollentaxon) %>% 
-      mutate(ntrait = n()) %>% 
-      filter(ntrait > 3)
-    
-    pol <- dfPOL %>% 
-      # select country
-      filter(country == selectedcountry) %>%
-      dplyr::select(sitename, pollentaxon, pollen_abun = all_of(selectedabun)) %>% 
-      drop_na()
-    
-    # save pollen taxa
[…]

### 2022-11-14 03_Vegetation_pollen_comparison.R
update study design

@@ -27,7 +27,7 @@ if (!require(readxl)) install.packages("readxl") # repel point labels
 dfPOL_Scot <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
 dfPOL_Swiss <- readRDS("RDS_files/01_Pollen_data_Swiss.rds")
 dfPOL <- bind_rows("Scotland" = dfPOL_Scot, "Switzerland" = dfPOL_Swiss, .id = "country") %>% 
-  dplyr::select(country,sitename, pollentaxon, percent, adjustedpercent_mean) %>% 
+  dplyr::select(country, sitename, pollentaxon, percent, adjustedpercent_mean) %>% 
   mutate(sitename = str_remove(sitename, pattern = "X"))
 polmode <- readRDS("RDS_files/01_Pollination_mode.rds")
 polspec <- readRDS("RDS_files/02_PollenType_species.rds") %>% 
@@ -35,15 +35,12 @@ polspec <- readRDS("RDS_files/02_PollenType_species.rds") %>%
 dfABUN_a <- readRDS("RDS_files/01_Species_abundance_a.rds")
 dfABUN_bc <- readRDS("RDS_files/01_Species_abundance_bc.rds") %>% ungroup
 bdm_abun <- readRDS("RDS_files/01_Species_abundance_Swiss.rds") 
-bdm_abun_bc <- readRDS("RDS_files/01_Species_abundance_zoneBC_swiss.rds")
 
 # Summarize vegetation data by pollen type
 zoneA <- dfABUN_a %>%
   # join with pollen to species table
-  left_join(polspec, by = "stand.spec") %>% 
-  mutate(pollentaxon = if_else(is.na(pollentaxon), "Not in pollen data", pollentaxon)) %>% 
-  # select plots that are comparable with Switzerland
-  filter(distance %in% c("0 meter", "1.5-3 meter")) %>% 
+  full_join(polspec, by = "stand.spec") %>% 
+  # mutate(pollentaxon = if_else(is.na(pollentaxon), "Not in pollen data", pollentaxon)) %>% 
   # calculate species abundance on the plot level
   group_by(sitename, pollentaxon) %>% 
   summarise(abun = sum(abun, na.rm = TRUE)) %>% 
@@ -52,20 +49,20 @@ zoneA <- dfABUN_a %>%
          country = "Scotland")
 zoneA_swiss <- bdm_abun %>% 
   # join with pollen to species table
[…]

### 2022-11-14 04_Correlations.R
script replaced by 04_correlation_bayesian

@@ -1,138 +1,0 @@
-## ---------------------------
-##
-## Script name: Correlations 
-##
-## Purpose of script: Calculating correlation between CWM species based and pollen based
-##
-## Author: Annegreet Veeken
-##
-## Date created: 08-02-2022
-## ---------------------------
-##
-## Notes:
-##   - re-run when leafC and leafN are available
-##
-## ---------------------------
-
-## Load libraries
-library(tidyverse)
-library(rjags) 
-library(runjags)
-library(mcmc)
-library(coda)
-library(gridExtra)
-
-## Load data and prepare
-dfCWM <- readRDS("RDS_files/03_Posterior_CWM.rds") 
-
-# Make df per trait, make tidy
-lCWM <- dfCWM %>%
-  dplyr::select(-id) %>% 
-  group_by(trait) %>% 
-  group_split(.keep = TRUE) %>% 
-  purrr::map(., ~pivot_wider(., id_cols = c(trait, sitename, country), names_from = zone ,
-                             values_from = Mean))
-
-
-
-# dfCWM <- readRDS("RDS_files/03_Posterior_CWM.rds") %>%
-#   as_tibble() %>%
-#   pivot_wider(names_from = zone, values_from = c(Mean,SD))
-# 
-# dfCWM %>% 
-#   filter(trait == "SLA") %>% 
-#   ggplot(aes(x = exp(Mean_pollencorrected), y = Mean_zoneA)) +
-#   geom_point(col = "darkorange") +
-#   geom_smooth(method = "lm", col = "darkorange") 
-# 
-# 
-# ggplot(dfCWM, aes(x = Mean_pollencorrected, y = Mean_zoneB)) +
-#   geom_point(col = "darkorchid") +
-#   geom_smooth(method = "lm", col = "darkorchid")+
-#   facet_wrap(~trait)
-# 
-# ggplot(dfCWM, aes(x = Mean_pollencorrected, y = Mean_speciesC)) +
-#   geom_point(col = "cyan4") +
-#   geom_smooth(method = "lm", col = "cyan4")+
-#   facet_wrap(~trait)
-
-## Plots ----
-plot_func <- function(df,traitname){
-p <- ggplot(df, aes(x = pollencorrected, y =zoneA)) +
-  geom_point(col = "darkorange") +
-  geom_smooth(method = "lm", col = "darkorange")+
[…]

### 2022-11-14 05_CWM_elevation_switzerland.R
initial commit

@@ -1,0 +1,249 @@
+## ---------------------------
+##
+## Script name:  05_CWM_elevation_switzerland
+##
+## Purpose of script: Calculating CWM change along elevation gradient from pollen and vegetation
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-09-22
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   
+## References:
+##
+## ---------------------------
+
+set.seed(1)
+
+## Load packages
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(rjags)) install.packages("rjags")
+if (!require(runjags)) install.packages("runjags")
+
+## Load data
+bdm_meta <- read_csv("Data/bdm_metadata.csv") %>% 
+  mutate(sitename = as.character(aID_STAO))
+# collate data
+traits <- c("LA","SLA","PlantHeight")
+
+pollen_files <- list.files("RDS_files/") %>% 
+  str_subset(pattern = "03_CWM_estimates_") %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
+  str_subset(pattern = "pollen") %>% 
+  str_subset(pattern = "gapfilled", negate = TRUE) %>% 
+  str_subset(pattern = "Switzerland") %>% 
+  str_subset(pattern = "draw", negate = TRUE)
+
+pollen_lab <- pollen_files %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
+  str_remove(., "03_CWM_estimates_") %>% 
+  str_remove(., ".rds")
+
+veg_files <- list.files("RDS_files/") %>% 
+  str_subset(pattern = "03_CWM_estimates_") %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
+  str_subset(pattern = "Switzerland") 
+
+veg_lab <- veg_files %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
+  str_remove(., "03_CWM_estimates_") %>% 
+  str_remove(., ".rds")
+
+folderpath.fun <- function(x)
+{paste("RDS_files/", x, sep = "/")}
+
+dfCWM_pol <- pollen_files %>% 
+  folderpath.fun(.) %>% 
+  purrr::map2(pollen_lab, ~readRDS(.) %>% 
+                as.data.frame() %>% 
+                rownames_to_column("parameter") %>% 
+                filter(., str_detect(parameter, "cwm\\[")) %>%
[…]

### 2022-11-14 05_Scatterplots.R
log transformed values

@@ -22,12 +22,9 @@
 if (!require(tidyverse)) install.packages("tidyverse")
 
 ## Load data
-dfCWM_veg <- readRDS("RDS_files/04_CWM_estimates_vegetation_Scotland.rds") %>%
-  mutate(across(c("veg_mean", "veg_sd"),~exp(.)))
+dfCWM_veg <- readRDS("RDS_files/04_CWM_estimates_vegetation_Scotland.rds") 
 dfCWM_pol <- readRDS("RDS_files/04_CWM_estimates_pollen_Scotland.rds") %>%
-  filter(!str_detect(correction,pattern = "draw")) %>%
-  mutate(across(c("pollen_mean", "pollen_sd"),~exp(.)))
-
+  filter(!str_detect(correction, pattern = "draw")) 
 # join data
 dfCWM <- dfCWM_veg %>%
   left_join(dfCWM_pol, by = c("country","sitename", "trait", "growthform","pollination")) %>%
@@ -41,21 +38,21 @@ dfCWM <- dfCWM_veg %>%
   filter(country == "Scotland")
 
 # Plot labels
-labs_trait <- c("Height (cm)",
-          "Leaf area (mm2)",
-          "SLA (mm2/mg)")
+labs_trait <- c("Height (cm)(log)",
+          "Leaf area (cm2)(log)",
+          "SLA (mm2/mg)(log)")
 names(labs_trait) <- c("PlantHeight", "LA", "SLA")
 labs_zone <- c(zoneA = "",
                zoneB = "",
                zoneC = "")
 
-# TRY data pollen, field data vegetation ----
+# gapfilled data pollen, field data + gapfilled vegetation ----
 windows()
 cwm_range <- dfCWM %>%
   group_by(trait) %>%
-  summarise(across(c("veg_mean","pollen_mean"), ~range(.))) %>%
-  transmute(max_lim = max(veg_mean, pollen_mean),
-         min_lim = min(veg_mean, pollen_mean)) %>%
+  summarise(across(c("veg_mean","pollen_mean"), ~range(., na.rm = TRUE))) %>%
+  transmute(max_lim = max(veg_mean, pollen_mean, na.rm = TRUE),
+            min_lim = min(veg_mean, pollen_mean, na.rm = TRUE)) %>%
   pivot_longer(-trait, names_to = "limits") %>%
   mutate(pollen_mean = value, veg_mean = value,
          correction = NA, taxres = NA, pollination = NA, growthform = NA)
[…]

### 2022-11-14 02_Prepare_trait_data_gapfilled.R
initial commit

@@ -1,0 +1,142 @@
+## ---------------------------
+##
+## Script name: 02_prepare_trait_data_gapfilled.R
+##
+## Purpose of script: Prepare gapfilled data for CWM calculation
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-07-19
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   
+##
+## ---------------------------
+
+
+## Load packages
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(LCVP)) install.packages("LCVP")
+if (!require(lcvplants)) install.packages("lcvplants")
+
+## Add pollen type to gapdata
+# Species per pollen type list
+spec_pollen <- readRDS("RDS_files/02_PollenType_species.rds")
+spec <- readRDS("RDS_files/01_Species_abundance_bc.rds") %>% 
+  pull(stand.spec) %>% 
+  c(., spec_pollen) %>% 
+  unlist() %>% 
+  unique
+
+# Gap-filled trait data
+gapdata <-  
+  read.csv("C:/Users/lgxgv/OneDrive - The University of Nottingham/Data analysis/Data/GapfilledTraitDataFS.csv") 
+gapdata <- gapdata %>% 
+    # Clean species names
+    mutate(Species = 
+           str_replace(Spp, "\\.", " ") %>%  # replace first . occurrence by space
+           str_replace("\\.", "-") %>% # replace second . occurrence by - for species names with hyphens
+             str_remove(" sp")) # remover sp epiphet 
+# check for missing species in the trait data
+misspec <-
+  spec[!spec %in% gapdata$Species] %>%
+  sort %>% unique
+
+# standardize species names gapfilled data
+# gapdat_sp <- gapdata$Species %>% 
+#   str_subset(pattern = "sp\\.", negate = TRUE) %>% 
+#   unique() %>% 
+#   lcvp_search(., progress_bar = TRUE)
+# saveRDS(gapdat_sp, "RDS_files/02_LCVP_search_gapdata.rds")
+gapdat_sp <- readRDS("RDS_files/02_LCVP_search_gapdata.rds")
+synonyms <- gapdat_sp %>%
+  mutate(
+    Input.Taxon = word(Input.Taxon, 1, 2) %>% # keep binomial without authority
+      str_remove("_x"),# x added for unknown reason
[…]

### 2022-11-14 Manuscript.Rmd
update after comments

@@
 
-# Reconstructing plant trait composition from palaeoecological data: testing the validity of the use of pollen data for functional trait reconstruction
+# Reconstructing plant trait composition from palaeoecological data: testing the validity of the use of pollen data for studying long-term changes in plant trait composition
 
@@
 Functional traits provide a better insight in to the processes that drive biodiversity and how biodiversity affects ecosystem functioning [@Hevia_2017].
-While the functioning of modern ecosystems is regularly contingent on past human and environmental factors[@Levis_2018; @Jogiste_2017; @Blonder_2018], the observational duration of trait-based studies is generally short[@Estes_218].
+While the functioning of modern ecosystems is regularly contingent on past human and environmental factors [@Levis_2018; @Jogiste_2017; @Blonder_2018; ], the observational duration of trait-based studies is generally at most a couple of decades [@Estes_218].
 So to extend the time scales of ecosystem functioning research, considerable interest in the trait-based approach has arisen in palaeoecology [@Brussel_2018; @Blaus_2020; @Carvalho_2019; @Reitalu_2015; @Sande_2019].
-Applying the trait-based approach in palaeoecological studies could offer a new way of interpreting pollen data and increase our understanding of the drivers of vegetation change on timescales of centuries to millennia.
-For instance, @Sande_2019 found that on a 7000-year time scale human disturbance superseeds climate effects on plant trait composition in the Amazon.
+Applying the trait-based approach in palaeoecological studies could offer a new way of interpreting pollen data and increase our understanding of the drivers of vegetation change on time-scales of centuries to millennia.
+For instance, @Sande_2019 found that on a 7000-year time scale human disturbance supersedes climate effects on plant trait composition in the Amazon.
 
-Although palaeoecological studies that have applied the trait-based approach show promising results, it is a relatively new approach of which the intricacies are underexplored.
+Although palaeoecological studies that have applied the trait-based approach show promising results, it is a relatively new approach of which the intricacies are under-explored.
 Little is known about how representative the pollen record is of past plant trait composition.
@@
 
-Taxonomic differences in pollen dispersal and production can cause an over- or underrepresentation of taxa in the pollen record.
+Taxonomic differences in pollen dispersal and production can cause an over- or under-representation of taxa in the pollen record.
 Wind-pollinated and woody taxa are generally better represented in the pollen record than herbaceous and insect-pollinated taxa as their pollen are produced in larger quantities and dispersed over greater distances.
-The pollen source area for wind pollinated taxa can therefore be a couple magnitudes greater than that of wind pollinated taxa [@Meltsov_2011] (extend argument?
+The source area of a pollen of a wind-pollinated taxon can therefore be a couple magnitudes greater than that of a insect-pollinated taxon [@Meltsov_2011] (extend argument?
 bunting 2003, bunting 2022).
-It has therefore been suggested by @Blaus_2020 to treat woody and non-woody taxa separately in the reconstruction of functional composition from pollen data.
-Another or additional way to correct for taxon specific differences use pollen productivity estimates, these correction factors down-weight the contribution of dominant taxa [@Githumbi_2022].
+Over- or under-representation can be corrected for with the use of taxon-specific pollen productivity estimates, these correction factors down-weight the contribution of dominant taxa [@Githumbi_2022].
+Furthermore, it has been suggested by @Blaus_2020 to treat woody and non-woody taxa separately in the reconstruction of functional composition from pollen data as their representative pollen source area differs greatly.
 
-Perhaps the most challenging in reliably reconstructing past trait is to attribute functional trait data to the taxa found in pollen records.
-It is rarely possible to resolve the pollen record to the species level [@Birks_2020].
+Perhaps the most challenging in reliably reconstructing past trait composition is to attribute functional trait data to the taxa found in pollen records, as it is rarely possible to resolve the pollen record to the species level [@Birks_2020].
 Trait variation within taxa can be substantial depending on the trait and the diversity of the taxon [@Albert_2010; @Prinzing_2008].
-It can be expected that this especially problematic in the case of large, abundant plant families that cannot easily be resolved to a lower taxonomic level, such as the Poaceae family.
+It can be expected that this is especially problematic in the case of large, abundant plant families that cannot easily be resolved to a lower taxonomic level, such as the Poaceae family.
 @Blaus_2020 and @Carvalho_2019 proposed that the high species number included in herbaceous pollen taxon could be one of the reasons why herbaceous vegetation was less well represented by pollen based functional diversity reconstructions.
@@
 
-With this study, we aim to test the validity of using of pollen records for plant trait composition reconstruction.
-We examine the effectiveness of pollen production factors to improve the pollen - functional diversity relationship.
-Furthermore, we test the representation of different plant functional groups by the pollen-based functional trait reconstruction.
-Laslty, we investigate the effect of taxonomic resolution on the reconstructions.
+With this study, we aim to test the validity of using pollen records for past plant trait composition reconstruction.
+We address two questions.
+Firstly, what factors influence the pollen - vegetation trait relationship?
+Using a data set from the Cairngorms, Scotland, we investigate the effect of spatial scale, plant functional group, correction factors and taxonomic resolution on the representativeness of the pollen assemblages for functional trait reconstruction.
+Secondly, can pollen based trait reconstruction predict changes in trait composition along environmental gradients?
+Using pollen, vegetation and trait data collected along an elevation gradient in Switzerland we test if pollen assemblages can predict changes in trait composition with elevation.
 
@@
 Sampling locations were at least 1 km from each other to avoid pollen source areas from overlapping with each other.
-Pollen, trait and species data was also available from `r length(bdm_meta)` sites across Switzerland were monitored in the context of the Biodiversity Monitoring Switzerland (BDM).
+Pollen, trait and species data was also available from `r nrow(bdm_meta)` sites across Switzerland were monitored in the context of the Biodiversity Monitoring Switzerland (BDM).
 This is a systematic sampling grid for the long-term monitoring of plant and animal species in Switzerland [@BDM_2022].
-Of these sites `r table(bdm_meta$habitat01)[1]` were forest and `table(bdm_meta$habitat01)[2]` grassland, spanning an elevation between `r min(bdm_meta$elevation)` and `r max(bdm_meta$elevation)`.
-Trait and species composition data from `r nrow(read.csv("Data/Swiss_LC/all_BDM_plots_coord.csv")) - length(bdm_meta)` were also used in this study.
+Trait and species composition data from an additional `r nrow(read.csv("Data/Swiss_LC/all_BDM_plots_coord.csv")) - nrow(bdm_meta)` sites that were surveyed in this network were also used in this study.
 
@@
 
-We compare trait composition of vegetation with trait composition reconstructed from surface pollen samples.
-Surface pollen samples are collected from moss polsters and approximate vegetation composition in recent years [@Birks_2016] [Other ref?].
-Study design mostly followed the protocol of @Bunting_2013 (Appendix), which allows for the calculation of vegetation composition in concentric circles around the moss sampling point.
-We calculated the vegetation composition in three nested concentric rings of 3.4m, 100m and 1000m in radius.
-The inner ring (3.4m) was surveyed in the field, the middle ring (100m) was surveyed in Scotland, but not in Switzerland.
-The vegetation cover of the outer ring (1000m) and the middle ring for Switzerland was estimated using QGIS and land cover maps [@SwissFC; @SwissLC;\@scotisch_government;\@QGIS_software].
+The data that was collected in Scotland was used to test if pollen-based trait reconstructions represents vegetation trait composition.
+We compare the community weighted mean (CWM) trait value of the vegetation with the CWM trait values reconstructed from surface pollen samples.
+Surface pollen samples are collected from moss polsters and approximate vegetation composition in recent years [@Birks_2016] [+Other ref?].
+Study design mostly followed the protocol of @Bunting_2013, which allows for the calculation of vegetation composition in concentric circles around the moss sampling point to test the spatial scale that is represented by the moss sample (Appendix).
+We calculated the vegetation composition in three nested concentric rings of 10 m, 100 m and 1000 m in radius.
+The inner ring (10 m) and the middle ring was surveyed in the field.
+The vegetation cover of the outer ring (1000 m) was estimated using QGIS and land cover maps [\@scotisch_government; \@QGIS_software].
+To test the effect of spatial scale on the representativeness of pollen assemblages for trait composition reconstruction, we test compare vegetation CWM and pollen CWM in the rings of 10 m, 100 m and 1000 m.
+Furthermore, four different treatments were assigned to test what factors could influence representation of trait composition by pollen data.
+Firstly, the use of correction factors, than the separation of vegetation and pollen data in to plant functional type and pollination mode.
+Lastly, the vegetation was aggregated to genus level and family level to test the effect of taxonomic resolution on the reconstructions.
 
-### Vegetation data
+To test if pollen-trait reconstructions can predict relative changes in vegetation trait composition along environmental gradients, we calculate changes in CWM trait values along an elevation gradient in Switzerland.
+The sites span an elevation between `r min(bdm_meta$elevation)` m and `r max(bdm_meta$elevation)`m.
+Of these sites `r table(bdm_meta$habitat01)[1]` were in forests and `r table(bdm_meta$habitat01)[2]` in grassland.
 
-In the inner ring (3.4m) 4 plots of 0.25 m^2^ (0.16m^2^ in Switzerland) were surveyed, one in every wind direction.
-At the Scottisch sites the middle ring was surveyed as well, between 4 and 6 plots of 10 by 10 meter were selected on a stratified random design based on vegetation types that were identified in the fields.
-Species abundance was estimated using percentage cover in Scotland.
-In Switzerland, each subplot was a quadrat with nine points (a 3 x 3 grid), and percentage coverage of the species present at the nine points was estimated.
+### Vegetation survey
+
+At the Scottish sites, in the inner ring (10 m) 17 sub-plots of 0.25 m^2^ were surveyed: one in the centre, 3 in every wind direction at 3, 6 and 10 meters, and one in the NE, NW, SE and SW direction (See design in Appendix).
+In the middle ring (10-100 m) between 4 and 6 plots of 5 by 5 meter were selected on a stratified random design based on vegetation types that were identified in the fields.
+Vegetation abundance was estimated using percentage cover.
+In Switzerland, the plots were 1.8 m in radius and contained 4 sub-plots.
+Each sub-plot was a quadrant of 0.16 m^2^ with nine points (a 3 x 3 grid), and percentage coverage of the species present at the nine points was estimated.
 All species names were standardized according to the Leipzig Plant Catalogue using the *lcvplants* and *lcvp* R packages [@freiberg2020lcvp].
 
-### Trait data
+### Trait data collection
 
@@
 XXX.
-In Scotland, trait data was collected in the middle ring (100m).
+In Scotland, trait data was collected in the middle ring (100 m).
 For determining plant height 10 flowering of individuals were selected at every site if present, non-flowering individuals were selected if flowering individuals were not available.
-Five leaf samples per species for specific leaf area (SLA), leaf area (LA), leaf dry matter content (LDMC), leaf carbon and leaf nitrogen were collected in seal-able plastic bags.
+Five leaf samples per species for specific leaf area (SLA) and leaf area (LA) collected in seal-able plastic bags.
 Fresh weight and surface area were measured within 72 hours of collection using a carat scale and flatbed scanner.
-Samples were dried at 70 C for 72 hours to constant dry weight before measuring dry weight.
+Samples were dried at 70 ***°***C for 72 hours to constant dry weight before measuring dry weight.
 In Switzerland, the trait data was collected for species present at the nine points in the 3 by 3 grid.
+XXXX
 
-For reconstructing trait composition from pollen data and for classifying vegetation data in to plant growth forms and pollination mode, trait data from the TRY database was used [@kattge2020try] (Appendix).
+For reconstructing trait composition from pollen data and for classifying vegetation data in to plant growth forms and pollination mode, previously gapfilled trait data from the TRY database was used [@kattge2020try] [\@Schrodt_2015].
+XXX
 
@@
 
-Moss was collected from multiple location within a 1m radius in the centre of the inner ring.
+Moss was collected from multiple location within a 1 m radius in the centre of each plot.
 Only living tissue was sampled.
@@
 Pollen productivity estimates were not available for all taxa, taxa without pollen productivity factors were excluded.
-On average, this led to the exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%) and `round(mean(missing_ppe_swiss$percent_missing),2)*100`% of the Swiss pollen data (max `r round(max(missing_ppe_swiss$percent_missing),2)*100`%).
-Pollen were divided in to two plant functional groups according to @Reitalu_2019 consisting of woody (trees and shrubs) ornon-woody species (grasses, herbs and dwarf shrubs of the Ericaceae family).
+On average, this led to the exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%).
+Pollen were divided in to two plant functional groups according to @Reitalu_2019 consisting of woody (trees and shrubs) or non-woody species (grasses, herbs and dwarf shrubs of the Ericaceae family).
 @Reitalu_2019 was also used to classify pollen in to pollination mode, either insect-pollinated or not insect-pollinated.
 
-For reconstructing plant trait composition from pollen data, assumptions need to be made about the species belonging to the pollen taxa in the data set.
+For reconstructing plant trait composition from pollen data, assumptions needed to be made about the species belonging to the pollen taxa in the data set.
+Therefore we created a pollen to species translation table.
 We used the *rgbif* package to download the current distribution of species belonging to each pollen taxon in Scotland and Switzerland from the Global Biodiversity Information Facility database (REF).
 Species distribution data was queried using the pollen taxonomic name, i.e.
-the lowest taxonomic name the pollen could be identified to .
+the lowest taxonomic name the pollen could be identified to.
+All trait data that was available for the species in the translation table was then used in the calculation trait values on the pollen taxonomic level .
 
@@
 First, the trait distribution on the species or the pollen taxonomic level was estimated.
-We assume that the trait data follows a lognormal distribution for the likelihood, so that$$trait_{i}^{observed} \sim { lognormal(\mu_{j},\sigma_j)}$$Where *i* represents each trait observation of a taxon (*j*) from the trait data.
+We assume that the trait data of SLA, LA and plant height follow a lognormal distribution for the likelihood, so that$$trait_{i}^{observed} \sim { lognormal(\mu_{j},\sigma_j)}$$Where *i* represents each trait observation of a taxon (*j*) from the trait data.
 $\mu_{j}$ and $\sigma_{j}$ are the mean trait value and the standard deviation per taxon, respectively.
 Vague priors were used for the taxon mean and standard deviation estimation.
-We then modeled the community weighted mean (CWM) as a mixture of normal distributions that were weighted by the species abundance or pollen percentage, so that
+We then modelled the community weighted mean (CWM) as a mixture of normal distributions that were weighted by the species abundance or pollen percentage, so that
 
@@
 Where *n* is the total number of taxa, and $\pi_{j}$ the abundance of each pollen taxon or species at a given site *s*.
-The mean *CWM~­~^mean^* and the standard deviation *CWM^sd^* of this modeled distribution was subsequently used in the estimation of the correlation between the trait composition of the vegetation and the pollen data.
+The mean *CWM~­~^mean^* and the standard deviation *CWM^sd^* of this modelled distribution was subsequently used in the estimation of the correlation between the trait composition of the vegetation and the pollen data.
 
-### Comparison between pollen and vegetation community weighted mean
+### Correlation between pollen and vegetation community weighted mean
 
-To test the correlation between the pollen based reconstruction of CWM and the vegetation CWM, we estimated the correlation coefficient *ρ* we assumed the data follows a multivariate t likelihood distribution so that, $$[CWM_{s}^{pol},CWM_{s}^{veg}] \sim t_{df}([\mu_{pol},\mu_{veg}],\sum) $$
+To test the correlation between the pollen based reconstruction of CWM and the vegetation CWM, we used a multivariate normal distribution to estimate correlation coefficient $\rho$ so that,
 
-Where we estimated the degrees of freedom ($df$) of the multivariate t-distribution with and exponential prior with a mean of 30 as the t-distribution becomes almost normal at values of 30 and above (Kruscke, dbda),
-
-$$
-df \sim Exp(\lambda)
-$$
-
-$$
-\lambda = 1/30
-$$
-
-We used an uninformative, flat prior for $\rho$ so that,
+$$  [CWM_{s}^{veg},CWM_{s}^{veg}] \sim MultivariateNormal([\mu_{veg},\mu_{veg}],\sum) $$
 
@@
 
-Correlations were calculated for all zones and three traits: SLA, LA and plant height.
-Furthermore, four different treatments were assigned to test what factors could influence representation of trait composition by pollen data: the use of correction factors, the separation of vegetation and pollen data in to plant functional type and pollination mode.
-Lastly, the vegetation was aggregated to genus level and family level to test the effect of taxonomic resolution on the reconstructions.
+Correlations were calculated for the different treatments, all zones and three traits: SLA, LA and plant height.
+
+### Community weighted mean changes with elevation
+
+To test if relative trait compositional changes are reflected by the pollen data along an elevation gradient, we fitted a generalized linear model with elevation as the predictor, so that
+
+$$CWM^{mean} \sim Normal(\alpha + \beta Y_{elev}, CWM^{sd})$$
+
+Where $\alpha$ and $\beta$ are the intercept and the slope, respectively, which are supplied by uninformative priors. $Y_{elev}$ is the elevation as the linear predictor.
 
@@
 
+![Estimated correlation between the pollen data and vegetation data.
+Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Pollen_correction_factors.png)
+
+\* add uncertainty around pollen correction factors
+
 ### Pollination mode
+
+![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Pollination_mode.png)
 
@@
 
+![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/PFT.png)
+
 ### Effect of taxonomic resolution
+
+![Estimated correlation between the pollen data and vegetation data.
+Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Taxonomic_resolution.png)
+
+### CWM change with elevation
+
+![CWM trait change with elevation for Leaf area, plant height and specific leaf area.
+Left column shows the CWM values for the vegetation data, right column shows the CWM values for the pollen data.
+Green lines show the relationships from the linear model, grey area shows the 95% credibility interval.
+Leaf area and plant height were log transformed before modelling.](Figures/20221016-Swiss_elevation_combined.png)
 
@@
     -   estimation of the zones
+
+-   Interspecific trait variation in LA
 
@@
 
-```{r echo=FALSE, fig.height=6, fig.width=10}
-source("03_Vegetation_pollen_comparison.R") # pollen plot
-pol_veg_plot
-```
+![Scatter plot of vegetation basal area % against pollen abundance %](Figures/Pollen_rep_percent.png)
 
-```{r echo=FALSE, fig.height=6, fig.width=10}
-source("03_Vegetation_pollen_comparison.R") # pollen plot
-pol_veg_plot_adj
-```
-
-### Site characteristics
-
-insert pollen plot
-
-```{r echo=FALSE}
-# source("03_pollen_diagrams.R")
-```
+![Scatter plot of vegetation basal area % against pollen abundance % adjusted with pollen productivity estimates.](Figures/Pollen_rep_adjustedpercent.png)
 
@@
 
-![Design of Swiss trait sampling and vegetation surveys](Other/BDM_study_design.png)
+![Study design in Scotland.
+The moss sample was collected in the middle of the inner ring.
+Vegetation surveys were performed in the inner and middle ring.
+Trait data was collected in the middle ring.
+The outer ring was surveyed using land cover maps and QGIS](images/paste-941650FE.png){alt="Study design in Scotland.
+The moss sample was collected in the middle of the inner ring.
+Vegetation surveys were performed in the inner and middle ring.
+Trait data was collected in the middle ring.
+The outer ring was surveyed using land cover maps and QGIS" width="573"}
 
-![Design of Scottish vegetation survey in zone A](Other/Scotland_field_design.png)
+![Design of Scottish vegetation survey in zone A](Other/Scotland_field_design.png){alt="Design of Scottish vegetation survey in zone A"}
 
-### Trait values
+![Design of Swiss trait sampling and vegetation surveys. The moss sample was collected in the centre of the plot.](Other/BDM_study_design.png)
 
-```{r echo=FALSE}
-sum_trait <- readRDS("RDS_files/01_summary_table_TRY_data.rds")
-kable(sum_trait)
-```
+### Trait data
 
-```{r echo=FALSE}
-source("01_Prepare_trait_data_vegetation.R",encoding ="UTF-8")
-kable(sum_ph)
-kable(sum_la)
-kable(sum_sla)
-```
+![](Figures/LA_trait_values.png)
+
+![](Figures/PlantHeight_trait_values.png)
+
+![](Figures/SLA_trait_values.png)

### 2022-11-14 Harmonization_table.R
tidy script

@@ -1,13 +1,34 @@
-# install.packages(c("tidyverse", "xlsx"))
-library(tidyverse)
-library(xlsx)
+## ---------------------------
+##
+## Script name: 
+##
+## Purpose of script: Create harmonization table for pollen types
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-07-13
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   
+##
+## ---------------------------
+
+
+## Load packages
+if(!require(tidyverse)) install.packages("tidyverse")
 
 # Accepted pollen types
-p_vars <- read.csv("Data/EPD data tables/P_VARS.csv", sep = ";")
+p_vars <- read_csv("Data/EPD data tables/P_VARS.csv", sep = ";")
 # Ecological group names
-p_group <- read.csv("Data/EPD data tables/P_GROUP.csv", sep = ";")
+p_group <- read_csv("Data/EPD data tables/P_GROUP.csv", sep = ";")
 # Group codes
-group_code <- read.csv("Data/EPD data tables/GROUP_CODES.csv", sep = ";")
+group_code <- read_csv("Data/EPD data tables/GROUP_CODES.csv", sep = ";")
+
+
 
 # Filter for accepted pollen types names
 acc_varname <- p_vars %>% 

### 2022-11-28 04_Correlation_bayesian.R
changed plot annotation

@@ -333,7 +333,7 @@ nobs_pft <- cwm_herb %>%
   dplyr::select(label,  zone, trait, Mean, treatment = "growthform") %>% 
   ungroup() %>% 
   arrange(trait) %>% 
-  mutate(x = rep(c(3,1.75,2,0.75,1), 3))
+  mutate(x = rep(c(2.84,1.75,2,0.75,1), 3))
 
 p_rho <- corrplot(rhos_pft, nobs_pft, c("trsh" = "Woody", "herb" = "Non-Woody"))
 ggsave("Figures/PFT_regression.png", p_rho, width = 7, height = 3)
@@ -349,8 +349,7 @@ cwm_gen <- dfCWM_pol %>%
          pollination == "allpol",
          correction == "no correction") %>% 
   left_join(dfCWM_veg, by = c("country", "sitename", "trait", "growthform", "pollination")) %>% 
-  filter(taxres == "genus") %>% 
-  mutate(x = rep(c(0.75,1,1.75,2,2.75,3), 3))
+  filter(taxres == "genus") 
 
 rhos_gen <- purrr::map2_dfr(combo$zone, combo$trait, 
                                    ~cor_bay(cwm_gen, selectedzone = .x, selectedtrait = .y,
@@ -378,8 +377,10 @@ nobs_taxres <- cwm_fam %>%
   left_join(rhos_taxres[rhos_taxres$parameter == "slope", ], by = c("trait",  "zone", "taxres" = "treatment")) %>%
   dplyr::select(label,  zone, trait, Mean, treatment = "taxres") %>%
   ungroup() %>%
-  arrange(trait) %>%
-  mutate(x = 0)
+  arrange(trait) %>% 
+  mutate(x = rep(c(2.66,2.84,3,
+                   1.66,1.84,2,
+                   0.66,0.84,1), 3))
 
 p_rho <- corrplot(rhos_taxres, nobs_taxres, c("family" = "Family",
                                               "genus" = "Genus",

### 2022-11-28 Manuscript.Rmd
manscript writing, prettifying figures

@@
 ## Abstract
+
+-   needs more testing
 
@@
 Secondly, can pollen based trait reconstruction predict changes in trait composition along environmental gradients?
-Using pollen, vegetation and trait data collected along an elevation gradient in Switzerland we test if pollen assemblages can predict changes in trait composition with elevation.
+Using pollen, vegetation and trait data collected along an elevation gradient in Switzerland we test if pollen assemblages can predict relative changes in trait composition with elevation.
 
@@
 Where *n* is the total number of taxa, and $\pi_{j}$ the abundance of each pollen taxon or species at a given site *s*.
-The mean *CWM~­~^mean^* and the standard deviation *CWM^sd^* of this modelled distribution was subsequently used in the estimation of the correlation between the trait composition of the vegetation and the pollen data.
+The mean ($CWM_{mean}$) and the standard deviation ($CWM_{sd}$)of this modelled distribution was subsequently used in the estimation of the correlation between the trait composition of the vegetation and the pollen data.
 
-### Correlation between pollen and vegetation community weighted mean
+### Relationship between pollen and vegetation community weighted mean
 
-To test the correlation between the pollen based reconstruction of CWM and the vegetation CWM, we used a multivariate normal distribution to estimate correlation coefficient $\rho$ so that,
+To test the relationship between the pollen based reconstruction of CWM and the vegetation CWM, we standardized the CWM values and fitted a students-t distributed likelihood for pollen CWM ($CWM_{mean}^{pol}$) with vegetation CWM ($CWM_{mean}^{veg}$) as the single predictor so that,
 
-$$  [CWM_{s}^{veg},CWM_{s}^{veg}] \sim MultivariateNormal([\mu_{veg},\mu_{veg}],\sum) $$
+$$  CWM_{mean}^{pol} \sim t(\beta_{t} CWM_{mean}^{veg},CWM_{sd}^{pol},df) $$
+
+The slope $\beta_{t}$ is a correlation-like metric defined with a flat, uninformative prior. $$
+\beta \sim Uniform(-1,1)
+$$
+
+We estimated the degrees of freedom ($df$) of the t-distribution with and exponential prior with a mean of 30, the t-distribution becomes almost normal at values of 30 and above (Kruscke, dbda),
 
 $$
-\sum \sim \begin{pmatrix} \sigma_{veg}^{2} & \rho\sigma_{veg}\sigma_{pol} \\ \rho\sigma_{veg}\sigma_{pol} & \sigma_{pol}^{2} \end{pmatrix}
+df \sim Exp(\lambda)
 $$
@@
 $$
-\rho \sim Uniform(-1,1)
+\lambda = 1/30
 $$
 
-Correlations were calculated for the different treatments, all zones and three traits: SLA, LA and plant height.
+We ran this model for every trait and zone in Scotland, and for the following different treatments: the use of correction factors, sub-setting the data in plant functional type, sub-setting the data in pollination mode and modelling on different taxonomic resolutions
 
@@
 
-To test if relative trait compositional changes are reflected by the pollen data along an elevation gradient, we fitted a generalized linear model with elevation as the predictor, so that
+To test if trait compositional changes of the vegetation along an elevation gradient are reflected by pollen based trait reconstructions, we fitted a generalized linear model with elevation as the predictor, so that the likelihood of the CWM value is defined as follows,
 
-$$CWM^{mean} \sim Normal(\alpha + \beta Y_{elev}, CWM^{sd})$$
+$$CWM^{mean} \sim t(\alpha + \beta Y_{elev}, CWM^{sd},df)$$
 
-Where $\alpha$ and $\beta$ are the intercept and the slope, respectively, which are supplied by uninformative priors. $Y_{elev}$ is the elevation as the linear predictor.
+Where $\alpha$ and $\beta$ are the intercept and the slope, respectively, which are supplied by uninformative priors.
+$Y_{elev}$ is the elevation as the linear predictor.
+We modelled this relationship for the grassland and forest plots separately.
 
@@
 
-### Effect of correction factors
+Over all treatments and spatial scales, we only found evidence of a significant positive relationship between pollen-based reconstructed CWM trait values and vegetation CWM values for plant height.
+We found little support for a positive association between pollen and vegetation trait value for leaf area and specific leaf area.
+The spatial scale matters for the relationship between pollen plant height and vegetation plant height, only at scales between 10-100 m the relationship is significant.
 
-![Estimated correlation between the pollen data and vegetation data.
-Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Pollen_correction_factors.png)
+Although pollen productivity estimates do seem to improve the representation of the taxa in the vegetation by the pollen data (Supplementary information), it does not help to improve the pollen-vegetation trait relationships.
+Taking in account uncertainty around pollen productivity estimates, does XXX by between --- and ---
+
+![Estimaed correlation between the pollen data and vegetation data.
+Whisker plots show the mean estimated correlation coefficient and the 95% credibility interval.
+Sample size is noted above the mean correlation coefficient.](Figures/Pollen_correction_factors_regression.png)
 
@@
 
-### Pollination mode
-
-![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Pollination_mode.png)
-
-### Plant functional group
-
-![Estimated correlation between the pollen data and vegetation data. Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/PFT.png)
-
-### Effect of taxonomic resolution
+We expected that the trait composition of wind-pollinated and woody vegetation would be better represented by the pollen data, especially at greater spatial scale.
+The plant height of not wind-pollinated taxa is well represented by the pollen data at the smallest spatial scale (0-10 m), but not at greater spatial scales or for other traits.
+The subset of wind-pollinated taxa does not show significant relsationships.
+Sub-setting pollen and vegetation data in to woody and non-woody vegetation did not improve pollen - vegetation trait relationships.
 
 ![Estimated correlation between the pollen data and vegetation data.
-Whisker plots show the median estimated correlation coefficient and the 95% credibility interval.](Figures/Taxonomic_resolution.png)
+Whisker plots show the mean estimated correlation coefficient and the 95% credibility interval.
+Sample size is noted above the mean correlation coefficient.](Figures/Pollination_mode_regression.png)
+
+![Estimated correlation between the pollen data and vegetation data.
+Whisker plots show the mean estimated correlation coefficient and the 95% credibility interval.
+Sample size is noted above the mean correlation coefficient.](Figures/PFT_regression.png)
+
+We expected pollen to better represent trait composition at higher taxonomic resolution, family or genus level, because of the general low taxonomic resolution of pollen data.
+We did not find that increasing the taxonomic resolution on which the trait distribution are modelled, improves the relationship between pollen and vegetation trait values.
+
+![Estimated correlation between the pollen data and vegetation data.
+Whisker plots show the mean estimated correlation coefficient and the 95% credibility interval.
+Sample size is noted above the mean correlation coefficient.](Figures/Taxonomic_resolution_regression.png)
 
@@
 
-![CWM trait change with elevation for Leaf area, plant height and specific leaf area.
-Left column shows the CWM values for the vegetation data, right column shows the CWM values for the pollen data.
-Green lines show the relationships from the linear model, grey area shows the 95% credibility interval.
-Leaf area and plant height were log transformed before modelling.](Figures/20221016-Swiss_elevation_combined.png)
+We expected to pollen to reproduce similar trends in trait change along an elevation gradient.
+Pollen do not represent the changes in vegetation CWM values well.
+Pollen underestimate LA and SLA values and the decrease in LA and SLA in grasslands that was observed in the vegetation is not predicted by the pollen data.
+Also the steep decrease in plant height along the elevation gradient is not reproduced by the pollen data, and the plant height in grassland is overestimated.
+
+![Vegetation CWM trait change with elevation for leaf area, plant height and specific leaf area and two vegetation types: forest and grassland.
+Darker shaded area shows the 95% credibility interval and lighter shade area shows the 95% prediction interval.](Figures/CWM_elevation_vegetaion.png)
+
+![Pollen CWM trait change with elevation for leaf area, plant height and specific leaf area and two vegetation types: forest and grassland.
+Darker shaded area shows the 95% credibility interval and lighter shade area shows the 95% prediction interval.](Figures/CWM_elevation_pollen.png)
 
 ## Discussion
+
+Pollen are the main proxy for reconstructing past plant composition, but the relationship between pollen assemblages and the vegetation composition is complicated.
+Although X-Y have showed positive relationships between pollen and vegetation composition, these are reasons why to be cautious
 
@@
     -   Prinzing (2008) - Less lineages -- more trait variation
+
+-   Look for elevation studies in pollen ecology
+
+-   Being careful about apply trait based palaeoecology
+
+    -   bigger sample size
 
@@
 
+-   positive relationships found by others
+
 -   Some pollen types consider many more species than others for example in this study XX .
@@
 although bayesian xxx
+
+-   Relationship is not significant, though still could be positive, as enocmpossed by the 95% confidence interval. More data, less uncertainty in the CWM estimates .
 
@@
 
-    -   here we surveyed sites with different openness.
+    -   here we surveyed sites with different openness .
+
+    -   The relationship between spatial scale and pollen records is complex (\@Abraham_2022) and dependant on the species richness and landscape openness
+
+        -   forest data not significantly related to florisitc data while open landscapes are.
+
+        -   highest correlation between pollen and species within 1.5-70 m in open landscapes
+
+        -   high correlation with pollen richness and landscape openess
+
+        -   Changes in the relationship with changing spatial scale are largely driven by the numbers of species newly appearing as the radius of surveyed area increases, especially as new habitats are added (DIRECT COPY)
+
+-   High productivity of tree pollen bias pollen-plant diversity relationships @Odgaard_1999
+
+-   Trees intercept pollen in their canopies @Jackson_1999
 
@@
 
-![Scatter plot of vegetation basal area % against pollen abundance %](Figures/Pollen_rep_percent.png)
+![Scatter plot of vegetation basal area % against pollen abundance %. Coloured dots represent either wind or non-wind pollinated taxa.](Figures/Pollen_rep_percent.png)
 
-![Scatter plot of vegetation basal area % against pollen abundance % adjusted with pollen productivity estimates.](Figures/Pollen_rep_adjustedpercent.png)
+![Scatter plot of vegetation basal area % against pollen abundance % adjusted with pollen productivity estimates.
+Grey dots show the effect of uncertainty around pollen productivity estimates.
+Coloured dots represent either wind or non-wind pollinated taxa.](Figures/Pollen_rep_adjustedpercent.png)
 
@@
 
+```{r}
+bdm_meta <- readRDS("RDS_files/04_zCWM_estimates_vegetation_Switzerland.rds")
+
+ggplot(data = bdm_meta) + 
+  geom_point(aes(x = elevation, y = TAP, colour = habitat01)) +
+  scale_color_manual("", 
+                     values = c(forests = "darkorange", grasslands = "purple"),
+                     label = c(forests = "Forests", grasslands = "Grassland")) + 
+  scale_x_continuous("Elevation (m.s.l.)") +
+  scale_y_continuous("Total annual precipitation (mm)") +
+  theme_bw()
+
+ggplot(data = bdm_meta) + 
+  geom_point(aes(x = elevation, y = MAT, colour = habitat01)) +
+  scale_color_manual("", 
+                     values = c(forests = "darkorange", grasslands = "purple"),
+                     label = c(forests = "Forests", grasslands = "Grassland")) + 
+  scale_x_continuous("Elevation (m.s.l.)") +
+  scale_y_continuous("Mean annual temperature (°C)") +
+  theme_bw()
+
+```
+
 ### Trait data
 
-![](Figures/LA_trait_values.png)
+![Comparison between the trait data collected in the field with trait data obtained from TRY](Figures/LA_trait_values.png)
 
-![](Figures/PlantHeight_trait_values.png)
+![Comparison between the trait data collected in the field with trait data obtained from TRY](Figures/PlantHeight_trait_values.png)
 
-![](Figures/SLA_trait_values.png)
+![Comparison between the trait data collected in the field with trait data obtained from TRY](Figures/SLA_trait_values.png)
+
+### Taxon level trait values
+
+![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_LA.png) ![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_SLA.png) ![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_PH.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_LA.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_SLA.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_PH.png) \### Scatterplots pollen and vegetation trait values

### 2022-11-28 04_Correlation_bayesian.R

@@ -186,9 +186,9 @@ corrplot <- function(rhos, nobs, labels){
                   position = position_dodge(0.5)) +
     geom_point(shape = 21, size = 2, position = position_dodge(0.5)) +
     # labels and scales
-    scale_x_discrete("", labels = c("zoneA" = "Zone A", 
-                                    "zoneB" = "Zone B", 
-                                    "zoneC"  = "Zone C")) +
+    scale_x_discrete("", labels = c("zoneA" = "Inner ring (0-10 m)", 
+                                    "zoneB" = "Middle ring (10-100 m)", 
+                                    "zoneC"  = "Outer ring (100-1000 m)")) +
     scale_y_continuous("", limits = c(-1,1)) +
     scale_fill_manual("", values = c("darkorange", "purple", "cyan4"),
                       labels = labels) +

### 2022-11-28 05_CWM_elevation_switzerland.R

@@ -270,7 +270,8 @@ purrr::map(traits, ~lm_bay(cwm = cwm_pol, selectedtrait = .x,
 purrr::map(traits, ~lm_bay(cwm = cwm_pol, selectedtrait = .x, 
                                       selectedhabitat = "grasslands", selecteddata = "Pollen"))
 
-
+# Plotting ----
+# load in lm results
 lm_files <- list.files("RDS_files/") %>% 
   str_subset(pattern = "05_lm")
 
@@ -279,12 +280,15 @@ lm_values <- lm_files %>%
   purrr::map_df(~readRDS(.)) 
 
 # plot model results
-cwm_range <- cwm_veg %>%
-  left_join(cwm_pol, by = c("sitename", "trait")) %>% 
+cwm_range <- lm_values %>% 
+  mutate(trait = factor(trait, level = c("LA", "PlantHeight", "SLA"),
+                        labels = c(LA = "Leaf~area~(mm^{2})(log)",
+                                   PlantHeight = "Height~(cm)(log)",
+                                   SLA = "SLA~(mm^{2}/mg)(log)"))) %>% 
   group_by(trait) %>%
-  summarise(across(c("cwm_mean.x","cwm_mean.y"), ~range(., na.rm = T))) %>%
-  transmute(max_lim = max(cwm_mean.x, cwm_mean.y, na.rm = T),
-            min_lim = min(cwm_mean.x, cwm_mean.y, na.rm = T)) %>%
+  summarise(across(c("uncertainty_lower","uncertainty_upper"), ~range(., na.rm = T))) %>%
+  transmute(max_lim = max(uncertainty_upper, na.rm = T),
+            min_lim = min(uncertainty_lower, na.rm = T)) %>%
   pivot_longer(-trait, names_to = "limits", values_to = "cwm_mean") %>%
   mutate(elevation = rep(c(400,2600,400,2600)), habitat01 = NA) 
  

### 2022-11-29 Manuscript.Rmd
writng manuscript, small figure adjustmetns

@@
 -   needs more testing
+
+-   we advocate for more testing with greater sample size
 
@@
 Pollen are the main proxy for reconstructing past plant composition, but the relationship between pollen assemblages and the vegetation composition is complicated.
-Although X-Y have showed positive relationships between pollen and vegetation composition, these are reasons why to be cautious
+Here the relationship is further complicated by throwing traits in the mix.
+Here we find little evidence for pollen being a reliable proxy for plant trait composition.
+Exploring the possible causes for the lack of a relationship between pollen CWM values and vegetation CWM values, we suggest that this either has to do with the representation of the vegetation by the pollen data (abundance, species richness) or the large trait variation within pollen types.
+
+Previous studies have showed that pollen assemblages are a reliable proxy for plant richness (@Abraham_2022, @Reitalu_2019), but these are reasons why to be cautious.
+Taxa can become over- or underrepresented in the pollen record because of taxon-specific differences in pollen production.
+Here we applied pollen productivity estimates to correct for the overproduction of certain taxa.
+While apply these correction factors improved the representation of the vegetation by the pollen data (supplementary), Applying this correction did not improve the relationship between pollen and vegetation trait values, however.
+Wind-pollinated taxa and woody taxa are generally best represented by pollen assemblages, but separating the data in to these two categories had no effect.
+
+The spatial scale that is represented by the pollen data is an important factor influencing the representativeness of pollen for vegetation reconstruction.
+All our treatments were assessed at three spatial scales.
+It seems that the intermediate scale, up to 100 m, is the most representative for the pollen data.
+But the relevant spatial scale is also dependent on the openness of the landscape (Bunting, ABraham).
+
+Despite here showing little evidence for pollen being a reliable proxy for past trait composition, we do not disencourage the use of it, we advocate for caution of the use., rather encaurage further testing of the validity of this proxy.
+And finding ways to improve it.
+X-Y studies have found a relationship, and looking at the confidence interval of the data we do not disproof the existence of a possitive correlation.
+We take into account the uncertainty around pollen trait values, but we had a very low sample size and a diverse landscape.
+Also moss samples vs.
+lake samples.
+
+the traits of species that do not show up in the dat.a
+
+-
+
+-   traits of species that do not show up in the pollen record.
+
+These studies have showed
 
@@
     -   Prinzing (2008) - Less lineages -- more trait variation
+    -   Sterwart (2022, preprint) - gapfilling performs well (?)
 
@@
     -   bigger sample size
-
--   Comparison between Switzerland and Scotland methods of sampling
-
-    -   vegetation surveys - all species versus max nine
-
-    -   estimation of the zones
 
@@
 -   pollen based trait reconstructions have been shown to correctly reflect environmental gradients as elevation (Sande) and and climatic variations (Barboni)
+
+-   exploring the reasons why we did not find correlations between pollen and vegetation values it could be:
+
+    -   low sample size
+
+    -   low pollen count (not picking up rare pollen in the vegetation) (run CWM without the species in the pollen record)
+
+    -   large trait variation in pollen
+
+    -   Is this a pollen problem or a trait problem?
 
@@
 
-![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_LA.png) ![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_SLA.png) ![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_PH.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_LA.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_SLA.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_PH.png) \### Scatterplots pollen and vegetation trait values
+![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_LA.png) ![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_SLA.png) ![Estimated trait values of pollen taxa and the data](Figures/Trait_values_pollen_PH.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_LA.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_SLA.png) ![Estimated trait values of vegetation taxa and the data](Figures/Trait_values_vegetation_PH.png)
+
+### Scatter plots pollen and vegetation trait values
+
+![](Figures/Scatter_correction.png)
+
+![](Figures/Scatter_growthform.png)
+
+![](Figures/Scatter_pollination.png)
+
+### ![](Figures/Scatter_taxonomic_resolution.png)Effect of removal of taxa on CWM trait values
+
+![](Figures/CWM_pollen_removed_taxa.png)

### 2022-11-29 05_Scatterplots.R

@@ -35,12 +35,16 @@ dfCWM <- dfCWM_veg %>%
                                (pollination != "allpol" & growthform == "allpft") ~ "Pollination mode",
                                (growthform != "allpft" & pollination == "allpol") ~ "Growth form"),
          zone = str_remove(zone, pattern = "Scotland |Switzerland ")) %>%
-  filter(country == "Scotland")
+  filter(country == "Scotland") %>% 
+  mutate(trait = factor(trait, level = c("LA", "PlantHeight", "SLA"),
+                        labels = c(LA = "Leaf~area~(mm^{2})(log)",
+                                   PlantHeight = "Height~(cm)(log)",
+                                   SLA = "SLA~(mm^{2}/mg)(log)")))
 
 # Plot labels
-labs_trait <- c("Height (cm)(log)",
-          "Leaf area (cm2)(log)",
-          "SLA (mm2/mg)(log)")
+labs_trait <- c("Height (cm)",
+          "Leaf area (mm2)",
+          "SLA (mm2/mg)")
 names(labs_trait) <- c("PlantHeight", "LA", "SLA")
 labs_zone <- c(zoneA = "",
                zoneB = "",
@@ -50,9 +54,9 @@ labs_zone <- c(zoneA = "",
 windows()
 cwm_range <- dfCWM %>%
   group_by(trait) %>%
-  summarise(across(c("veg_mean","pollen_mean"), ~range(., na.rm = TRUE))) %>%
-  transmute(max_lim = max(veg_mean, pollen_mean, na.rm = TRUE),
-            min_lim = min(veg_mean, pollen_mean, na.rm = TRUE)) %>%
+  summarise(across(c("veg_mean","pollen_mean"), ~range(.))) %>%
+  transmute(max_lim = max(veg_mean, pollen_mean),
+         min_lim = min(veg_mean, pollen_mean)) %>%
   pivot_longer(-trait, names_to = "limits") %>%
   mutate(pollen_mean = value, veg_mean = value,
          correction = NA, taxres = NA, pollination = NA, growthform = NA)
@@ -71,8 +75,7 @@ cwm_range <- dfCWM %>%
     facet_wrap(zone~trait,
                scales = "free",
                labeller = labeller(trait = labs_trait,
-                                    zone = labs_zone)
-               ) +
+                                    zone = labs_zone)) +
     theme_bw() +
     theme(legend.position =  "bottom",
           strip.background = element_blank(),
[…]

### 2022-11-29 05_Scatterplots.R
small figure adjustment

@@ -35,16 +35,12 @@ dfCWM <- dfCWM_veg %>%
                                (pollination != "allpol" & growthform == "allpft") ~ "Pollination mode",
                                (growthform != "allpft" & pollination == "allpol") ~ "Growth form"),
          zone = str_remove(zone, pattern = "Scotland |Switzerland ")) %>%
-  filter(country == "Scotland") %>% 
-  mutate(trait = factor(trait, level = c("LA", "PlantHeight", "SLA"),
-                        labels = c(LA = "Leaf~area~(mm^{2})(log)",
-                                   PlantHeight = "Height~(cm)(log)",
-                                   SLA = "SLA~(mm^{2}/mg)(log)")))
+  filter(country == "Scotland") 
 
 # Plot labels
-labs_trait <- c("Height (cm)",
-          "Leaf area (mm2)",
-          "SLA (mm2/mg)")
+labs_trait <- c("Height (cm)(log)",
+                "Leaf area (cm2)(log)",
+                "SLA (mm2/mg)(log)")
 names(labs_trait) <- c("PlantHeight", "LA", "SLA")
 labs_zone <- c(zoneA = "",
                zoneB = "",
@@ -54,9 +50,9 @@ labs_zone <- c(zoneA = "",
 windows()
 cwm_range <- dfCWM %>%
   group_by(trait) %>%
-  summarise(across(c("veg_mean","pollen_mean"), ~range(.))) %>%
-  transmute(max_lim = max(veg_mean, pollen_mean),
-         min_lim = min(veg_mean, pollen_mean)) %>%
+  summarise(across(c("veg_mean","pollen_mean"), ~range(., na.rm = TRUE))) %>%
+  transmute(max_lim = max(veg_mean, pollen_mean, na.rm = TRUE),
+            min_lim = min(veg_mean, pollen_mean, na.rm = TRUE)) %>%
   pivot_longer(-trait, names_to = "limits") %>%
   mutate(pollen_mean = value, veg_mean = value,
          correction = NA, taxres = NA, pollination = NA, growthform = NA)
@@ -75,7 +71,8 @@ cwm_range <- dfCWM %>%
     facet_wrap(zone~trait,
                scales = "free",
                labeller = labeller(trait = labs_trait,
-                                    zone = labs_zone)) +
+                                   zone = labs_zone)
+    ) +
     theme_bw() +
     theme(legend.position =  "bottom",
[…]

### 2022-11-30 02_PollenType_Species_GBIF.R
extend species list

@@ -23,10 +23,10 @@
 
 
 ## Load packages
-if(!require(tidyverse)) install.packages("tidyverse")
-if(!require(rgbif)) install.packages("rgbif")
-if(!require(LCVP)) devtools::install_github("idiv-biodiversity/LCVP")
-if(!require(lcvplants)) devtools::install_github("idiv-biodiversity/lcvplants")
+if (!require(tidyverse)) install.packages("tidyverse")
+if (!require(rgbif)) install.packages("rgbif")
+if (!require(LCVP)) devtools::install_github("idiv-biodiversity/LCVP")
+if (!require(lcvplants)) devtools::install_github("idiv-biodiversity/lcvplants")
 
 # Load in pollen data
 dfPOL_scot <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
@@ -112,7 +112,7 @@ countrykey <- countrykeys[1]
 taxonkey <- taxonkeys[1:2]
 
 ## create function  that downloads, cleans, saves and creates conversion table for pollen type to species
-if(1){
+if (1) {
 gbifsaveclean <- function(taxonkey, countrykey){
   # perform gbif download request
   occsearch <- occ_download(
@@ -263,78 +263,108 @@ specpol <- specpol  %>%
   add_row(lcvp_scot, country = "Scotland") %>%
   add_row(lcvp_swiss, country = "Switzerland") %>% 
   # missing species
-  add_row(family = "Cyperaceae", genus = "Carex", stand.spec= "Carex", 
+  add_row(family = "Cyperaceae", genus = "Carex", stand.spec = "Carex", 
              country = "Switzerland", pollentaxon = "Cyperaceae") %>% 
-  add_row(family = "Juncaceae", genus = "Juncus", stand.spec= "Juncus", 
+  add_row(family = "Juncaceae", genus = "Juncus", stand.spec = "Juncus", 
           country = "Switzerland", pollentaxon = "Cyperaceae") %>%
-  add_row(family = "Caryophyllaceae", genus = "Cerastium", stand.spec= "Cerastium", 
+  add_row(family = "Caryophyllaceae", genus = "Cerastium", stand.spec = "Cerastium", 
           country = "Switzerland", pollentaxon = "Caryophyllaceae") %>% 
-  add_row(family = "Ranunculaceae", genus = "Ranunculus", stand.spec= "Ranunculus", 
+  add_row(family = "Ranunculaceae", genus = "Ranunculus", stand.spec = "Ranunculus", 
[…]

### 2022-11-30 04_Correlation_bayesian.R
change plot labels

@@ -186,9 +186,9 @@ corrplot <- function(rhos, nobs, labels){
                   position = position_dodge(0.5)) +
     geom_point(shape = 21, size = 2, position = position_dodge(0.5)) +
     # labels and scales
-    scale_x_discrete("", labels = c("zoneA" = "Inner ring (0-10 m)", 
-                                    "zoneB" = "Middle ring (10-100 m)", 
-                                    "zoneC"  = "Outer ring (100-1000 m)")) +
+    scale_x_discrete("", labels = c("zoneA" = "Inner ring (10 m)", 
+                                    "zoneB" = "Middle ring (100 m)", 
+                                    "zoneC"  = "Outer ring (1000 m)")) +
     scale_y_continuous("", limits = c(-1,1)) +
     scale_fill_manual("", values = c("darkorange", "purple", "cyan4"),
                       labels = labels) +
@@ -333,7 +333,7 @@ nobs_pft <- cwm_herb %>%
   dplyr::select(label,  zone, trait, Mean, treatment = "growthform") %>% 
   ungroup() %>% 
   arrange(trait) %>% 
-  mutate(x = rep(c(2.84,1.75,2,0.75,1), 3))
+  mutate(x = rep(c(2.86,1.75,2,0.75,1), 3))
 
 p_rho <- corrplot(rhos_pft, nobs_pft, c("trsh" = "Woody", "herb" = "Non-Woody"))
 ggsave("Figures/PFT_regression.png", p_rho, width = 7, height = 3)
@@ -386,7 +386,7 @@ p_rho <- corrplot(rhos_taxres, nobs_taxres, c("family" = "Family",
                                               "genus" = "Genus",
                                               "stand.spec" = "Species"))
 
-ggsave("Figures/Taxonomic_resolution_regression.png", p_rho, width = 7, height = 3)
+ggsave("Figures/Taxonomic_resolution_regression.png", p_rho, width = 7, height = 4)
 
 ## Pollen correction + uncertainty ----
 cor_bay_draw <- function(cwm, selectedzone, selectedtrait, 

### 2022-11-30 SI_Remove_taxa.R
make plot

@@ -65,19 +65,19 @@ labs_trait <- c("Height (cm)(log)",
 names(labs_trait) <- c("PlantHeight", "LA", "SLA")
 
 # Run model
-plan(multisession(workers = 4))
+plan(multisession(workers = 2))
 furrr::future_map(unique(dfPOL_Scot$pollentaxon),
                     ~cwm_pol_remove(selectedtrait = "LA",
                                 selectedcountry = "Scotland",
                                 selectedabun = "percent",
                                 removedspec = .x),
                   .options = furrr_options(seed = TRUE))
-furrr::future_map(unique(dfPOL_Scot$pollentaxon),
-                  ~cwm_pol_remove(selectedtrait = "SLA",
-                              selectedcountry = "Scotland",
-                              selectedabun = "percent",
-                              removedspec = .x),
-                  .options = furrr_options(seed = TRUE))   
+# furrr::future_map(unique(dfPOL_Scot$pollentaxon),
+#                   ~cwm_pol_remove(selectedtrait = "SLA",
+#                               selectedcountry = "Scotland",
+#                               selectedabun = "percent",
+#                               removedspec = .x),
+#                   .options = furrr_options(seed = TRUE))   
 furrr::future_map(unique(dfPOL_Scot$pollentaxon),
                   ~cwm_pol_remove(selectedtrait = "PlantHeight",
                               selectedcountry = "Scotland",
@@ -93,18 +93,30 @@ pollen_files <-
   str_subset(pattern = paste(unique(dfPOL_Scot$pollentaxon), collapse = "|")) %>% 
   str_subset(pattern = "reduced_sp", negate = T)
 
-dfCWM_pol <- pollen_files %>% 
-  purrr::map2(., str_remove(pollen_files, "RDS_files/"), ~readRDS(.x) %>% 
-                as.data.frame())  %>% 
-  bind_rows()  
+pollen_files_all <-  list.files("RDS_files/",full.names = TRUE) %>% 
+  str_subset(pattern = "03_CWM_estimates_pollen_Scotland_LA_percent.rds|03_CWM_estimates_pollen_Scotland_SLA_percent.rds|03_CWM_estimates_pollen_Scotland_PlantHeight_percent.rds")
 
-ggplot(dfCWM_pol, aes(x = Mean, y = removed_spec)) +
[…]

### 2022-11-30 SI_Vegetation_pollen_comparison.R
add representation and trait values plot

@@ -11,9 +11,10 @@
 ## Email: veeken.g.a@gmail.com
 ##
 ## ---------------------------
-##
+## References: 
+## - pollen representation plot (Julier 2018, https://doi.org/10.1080/01916122.2017.1356392)
 ## Notes:
-##   - 
+##  
 ##     
 ## ---------------------------
 
@@ -75,18 +76,15 @@ zoneC <- dfABUN_bc %>%
   mutate(pollentaxon = if_else(is.na(pollentaxon), "Not in pollen data", pollentaxon)) %>% 
   # calculate species abundance on the plot level
   group_by(sitename, pollentaxon) %>% 
-  summarise(abun = sum(spec_abun_c)) %>% 
-  mutate(abun = abun/sum(abun)) %>% 
+  summarise(abun = sum(spec_abun_c, na.rm = T)) %>% 
+  mutate(abun = abun/sum(abun, na.rm = T)) %>% 
   filter(!is.nan(abun)) %>% 
   mutate(zone = "zoneC",
          country = "Scotland")
 
 veg <- bind_rows(zoneA, zoneA_swiss, zoneB, zoneC) %>% 
-  left_join(dfPOL, by = c("country","sitename","pollentaxon")) %>% 
-  filter(!is.na(abun))
+  left_join(dfPOL, by = c("country","sitename","pollentaxon")) 
 
-
-## Calculate pollen representation pot (Julier 2018, https://doi.org/10.1080/01916122.2017.1356392)
 ## Plot pollen representation per zone ----
 zone_lab <- c("Switzerland (1.8 m)", "Inner ring (10 m)", "Middle ring (100 m)", "Outer ring (1km)")
 names(zone_lab) <- c("Switzerland","zoneA", "zoneB", "zoneC")
@@ -99,18 +97,20 @@ nopol <- veg %>%
   ungroup() 
 pol_veg <- veg %>% 
   group_by(country, zone, pollentaxon) %>% 
-  summarise(veg = mean(abun,na.rm = T)*100, 
-            pol = mean(percent,na.rm = T)*100) %>% 
+  summarise(veg = mean(abun, na.rm = T)*100, 
+            pol = mean(percent, na.rm = T)*100) %>% 
+  filter(!(is.nan(veg) & is.nan(pol))) %>% 
   mutate(r_rel = pol/veg,
          r_rel_abs =  case_when(r_rel == Inf ~ "not in vegetation data",
                             is.nan(r_rel) ~ "not in pollen data")
             ) %>% 
[…]

### 2022-11-30 SI_Trait_data_field_TRY.R
fix plot labels

@@ -21,17 +21,12 @@
 ## Load packages
 if (!require(tidyverse)) install.packages("tidyverse")
 
-# Trait data from TRY for pollen based reconstruction
-TRY <- readRDS("RDS_files/01_TRY_raw_traits.rds") %>% 
-  dplyr::select(stand.spec, pollentaxon, PlantHeight, SLA, LA) %>% 
-  filter(!is.na(pollentaxon)) %>% 
-  distinct() %>% 
-  mutate(source = "TRY") 
 gapfilled <- readRDS("RDS_files/02_Gapfilled_traits_pollen.rds") %>% 
+  filter(country == "Switzerland") %>% 
   dplyr::select(stand.spec, pollentaxon, PlantHeight, SLA, LA) %>% 
   filter(!is.na(pollentaxon)) %>% 
   distinct() %>% 
-  mutate(source = "Gapfilled") 
+  mutate(source = "Gapfilled TRY data") 
 spec <- readRDS("RDS_files/02_PollenType_species.rds") %>% 
   dplyr::select(-country) %>% 
   distinct()
@@ -41,98 +36,97 @@ plantheight <- readRDS("RDS_files/01_PlantHeight.rds") %>%
   ungroup() %>% 
   left_join(spec, by = "stand.spec") %>% 
   select(stand.spec,pollentaxon, PlantHeight) %>% 
-  mutate(source = "Field")
+  mutate(source = "Field Scotland")
 sla <- readRDS("RDS_files/01_SLA.rds") %>% 
   ungroup() %>% 
   left_join(spec, by = "stand.spec") %>% 
   select(stand.spec,pollentaxon, SLA) %>% 
-  mutate(source = "Field")
+  mutate(source = "Field Scotland")
 la <- readRDS("RDS_files/01_LA.rds") %>% 
   ungroup() %>% 
   left_join(spec, by = "stand.spec") %>% 
   select(stand.spec,pollentaxon, LA) %>% 
-  mutate(source = "Field")
+  mutate(source = "Field Scotland")
 
 lTrait <- list(PlantHeight = plantheight,
-               LDMC = ldmc,
                SLA = sla,
-               LA = la,
-               LeafC = c,
-               LeafN = n)
+               LA = la)
 # Load swiss data
 bdm_ph <- readRDS("RDS_files/01_Traits_Swiss.rds") %>% 
+  # filter trees, only measured up to 200 cm
+  filter(!PFT == "TR") %>% 
   left_join(spec, by = "stand.spec") %>% 
[…]

### 2022-12-28 Manuscript.Rmd
update from macbook

@@
     fig_caption: yes
-bibliography: Validation_study_references.bib
 ---
@@
 
+-   need to take in to account trait variation pollen types
+
 ## Introduction
@@
 Functional traits provide a better insight in to the processes that drive biodiversity and how biodiversity affects ecosystem functioning [@Hevia_2017].
-While the functioning of modern ecosystems is regularly contingent on past human and environmental factors [@Levis_2018; @Jogiste_2017; @Blonder_2018; ], the observational duration of trait-based studies is generally at most a couple of decades [@Estes_218].
+While the functioning of modern ecosystems is regularly contingent on past human and environmental factors [@Levis_2018; @Jogiste_2017; @Blonder_2018], the observational duration of trait-based studies is generally at most a couple of decades [@Estes_218].
 So to extend the time scales of ecosystem functioning research, considerable interest in the trait-based approach has arisen in palaeoecology [@Brussel_2018; @Blaus_2020; @Carvalho_2019; @Reitalu_2015; @Sande_2019].
@@
 bunting 2003, bunting 2022).
-Over- or under-representation can be corrected for with the use of taxon-specific pollen productivity estimates, these correction factors down-weight the contribution of dominant taxa [@Githumbi_2022].
+Over- or underrepresentation can be corrected for with the use of taxon-specific pollen productivity estimates, these correction factors down-weight the contribution of dominant taxa [@Githumbi_2022].
 Furthermore, it has been suggested by @Blaus_2020 to treat woody and non-woody taxa separately in the reconstruction of functional composition from pollen data as their representative pollen source area differs greatly.
 
-Perhaps the most challenging in reliably reconstructing past trait composition is to attribute functional trait data to the taxa found in pollen records, as it is rarely possible to resolve the pollen record to the species level [@Birks_2020].
+The second challenge in reliably reconstructing past trait composition is to attribute functional trait data to the taxa found in pollen records, as it is rarely possible to resolve the pollen record to the species level [@Birks_2020].
 Trait variation within taxa can be substantial depending on the trait and the diversity of the taxon [@Albert_2010; @Prinzing_2008].
@@
 @Blaus_2020 and @Carvalho_2019 proposed that the high species number included in herbaceous pollen taxon could be one of the reasons why herbaceous vegetation was less well represented by pollen based functional diversity reconstructions.
-In both studies, however, traits values of pollen taxa were calculated by averaging over the available species level trait information.
 In this study, we use a Bayesian approach for modelling trait composition from pollen data, thereby allowing for the propagation of uncertainty in the taxon level trait reconstruction to the community level.
+
+Alternative to forming a direct representation of past plant trait composition, pollen based trait reconstruction might be better suited to study relative changes in trait composition.
+For instance, @Sande202 showed that pollen assemblages reflect relative trait changes in trait composition of the vegetation.
 
@@
 This is a systematic sampling grid for the long-term monitoring of plant and animal species in Switzerland [@BDM_2022].
-Trait and species composition data from an additional `r nrow(read.csv("Data/Swiss_LC/all_BDM_plots_coord.csv")) - nrow(bdm_meta)` sites that were surveyed in this network were also used in this study.
+Trait and species composition data, but not pollen data, from an additional `r nrow(read.csv("Data/Swiss_LC/all_BDM_plots_coord.csv")) - nrow(bdm_meta)` sites that were surveyed in this network were also used in this study.
 
@@
 We compare the community weighted mean (CWM) trait value of the vegetation with the CWM trait values reconstructed from surface pollen samples.
-Surface pollen samples are collected from moss polsters and approximate vegetation composition in recent years [@Birks_2016] [+Other ref?].
+Surface pollen samples are collected from moss polsters and approximate vegetation composition in recent years[@Pardoe2010].
 Study design mostly followed the protocol of @Bunting_2013, which allows for the calculation of vegetation composition in concentric circles around the moss sampling point to test the spatial scale that is represented by the moss sample (Appendix).
@@
 
-Three traits were selected for the analysis, specific leaf area, leaf area and and plant height.
-These traits reflect different components in the spectrum of plant strategies.
-XXX.
+We used three traits for the analysis, specific leaf area, leaf area and and plant height.
+These traits reflect different components in the spectrum of plant strategies [@Diaz2016].
 In Scotland, trait data was collected in the middle ring (100 m).
-For determining plant height 10 flowering of individuals were selected at every site if present, non-flowering individuals were selected if flowering individuals were not available.
-Five leaf samples per species for specific leaf area (SLA) and leaf area (LA) collected in seal-able plastic bags.
+For determining plant height, we randomly selected 10 flowering individuals of each species at every site.
+Non-flowering individuals were selected if flowering individuals were not available.
+For determining specific leaf area (SLA) and leaf area (LA), we collected five leaf samples per species in sealable plastic bags.
 Fresh weight and surface area were measured within 72 hours of collection using a carat scale and flatbed scanner.
 Samples were dried at 70 ***°***C for 72 hours to constant dry weight before measuring dry weight.
-In Switzerland, the trait data was collected for species present at the nine points in the 3 by 3 grid.
-XXXX
+In Switzerland, trait data was collected for the species present at the nine points in the 3 by 3 grid, otherwise trait measurement was the same.
 
-For reconstructing trait composition from pollen data and for classifying vegetation data in to plant growth forms and pollination mode, previously gapfilled trait data from the TRY database was used [@kattge2020try] [\@Schrodt_2015].
-XXX
+For reconstructing trait composition from pollen data and previously gapfilled trait data from the TRY database was used [@kattge2020try] [\@Schrodt_2015].
+for classifying vegetation data in to plant growth forms and pollination mode,
 
@@
 
-Pollen are the main proxy for reconstructing past plant composition, but the relationship between pollen assemblages and the vegetation composition is complicated.
-Here the relationship is further complicated by throwing traits in the mix.
-Here we find little evidence for pollen being a reliable proxy for plant trait composition.
-Exploring the possible causes for the lack of a relationship between pollen CWM values and vegetation CWM values, we suggest that this either has to do with the representation of the vegetation by the pollen data (abundance, species richness) or the large trait variation within pollen types.
+Pollen assemblages are the main proxy for studying past vegetation features, and are a reliable proxy for plant species richness and XX.
+However the relationship with vegetation trait composition is complex.
+Here we find little evidence for pollen being a reliable proxy for plant trait composition.but the relationship between pollen assemblages and the vegetation composition is complex Exploring the possible causes for the lack of a relationship between pollen CWM values and vegetation CWM values, we suggest that this either has to do with the representation of the vegetation by the pollen data (abundance, species richness) or the large trait variation within pollen types.
+
+vegetation structure (Switzerland) and scale (Scotland)
 
@@
 
--
+## Conclusion
+Characterizing the factors that influence the relationship between vegetation trait composition and , and validating the relationship between trait composition of the vegetation and is an important step towards the application of traits in palaeoecology.
+
+-   swiss - grass land receiving input from forests
 
@@
 
-### Pollen representation
+## Pollen percentages
+
+![Pollen percentage of the most abundant (\> 5%) taxa in Scotland.
+The effect of the application of pollen productivity estimates to the representation of pollen taxa is shown by the coloured bars.
+![Pollen percentage of the most abundant (\> 5%) taxa in Switzerland.
+Sites are sorted in order of increasing elevation (top to bottom).
+The effect of the application of pollen productivity estimates to the representation of pollen taxa is shown by the coloured bars.](Figures/Pollen_percentage_swiss.png)](Figures/Pollen_percentage_scotland.png)
+
+## Pollen representation
 
@@
 
-### ![](Figures/Scatter_taxonomic_resolution.png)Effect of removal of taxa on CWM trait values
+### ![](Figures/Scatter_taxonomic_resolution.png)
+
+### Effect of removal of taxa on CWM trait values
 

### 2022-12-28 01_PFT_table_vegetation.R

@@ -289,6 +289,7 @@ dfABUN  <- bind_rows(zoneA, zoneB, zoneC,
                      bdm_zoneA)
 
 sum_polmode <- dfABUN %>% 
+  filter(!stand.spec == "Unindentified") %>% 
   left_join(dfPFT, by = c("stand.spec" = "AccSpeciesName")) %>% 
   group_by(zone, sitename, polmode) %>% 
   # percentage of pollination type
@@ -297,6 +298,7 @@ sum_polmode <- dfABUN %>%
   pivot_wider(names_from = polmode, values_from = abun)
   
 sum <- dfABUN %>% 
+  filter(!stand.spec == "Unindentified") %>% 
   left_join(dfPFT, by = c("stand.spec" = "AccSpeciesName")) %>% 
   mutate(growthform = case_when(growthform %in% c("tree", "shrub") ~ "Woody",
                                 growthform %in% c("herb", "grass", "fern") ~ "Non-Woody")

### 2022-12-28 02_CWM_estimates_vegetation.R

@@ -19,7 +19,6 @@
 
 set.seed(1)
 
-
 ## Load packages
 if (!require(tidyverse)) install.packages("tidyverse")
 if (!require(rjags))  install.packages("rjags")
@@ -137,158 +136,158 @@ if (Sys.info()[4] == "LUIN67311") {
 }
 
 # Scotland----
-if (0) { 
+if (1) { 
 # Zone A ----
 # all
- furrr::future_map(scottraits,
-                   ~cwm_veg_scot(selectedabun = "zoneA",
-                             selectedtrait = .,
-                             selectedtaxres = "stand.spec",
-                             selectedpolmode = allpol,
-                             selectedpft = allpft),
-                   .options = furrr_options(seed = TRUE))
- # only taxa in pollen data
- furrr::future_map(scottraits,
-                   ~cwm_veg_scot(selectedabun = "zoneA",
-                             selectedtrait = .,
-                             selectedtaxres = "stand.spec",
-                             selectedpolmode = allpol,
-                             selectedpft = allpft,
-                             notinpollen = TRUE),
-                   .options = furrr_options(seed = TRUE))
- # pollination
- furrr::future_map(scottraits,
-                   ~cwm_veg_scot(selectedabun = "zoneA",
-                             selectedtrait = .,
-                             selectedtaxres = "stand.spec",
-                             selectedpolmode = nowind,
-                             selectedpft = allpft),
-                   .options = furrr_options(seed = TRUE))
- furrr::future_map(scottraits,
-                   ~cwm_veg_scot(selectedabun = "zoneA",
-                             selectedtrait = .,
-                             selectedtaxres = "stand.spec",
-                             selectedpolmode = wind,
-                             selectedpft = allpft),
-                   .options = furrr_options(seed = TRUE))
-
- # growth form
- # barely trees in zone A -> only do herbs
- furrr::future_map(scottraits,
-                   ~cwm_veg_scot(selectedabun = "zoneA",
-                             selectedtrait = .,
-                             selectedtaxres = "stand.spec",
-                             selectedpolmode = allpol,
-                             selectedpft = herb
-                             ),
-                   .options = furrr_options(seed = TRUE))
- # Taxonomic resolution
- furrr::future_map(scottraits,
-                   ~cwm_veg_scot(selectedabun = "zoneA",
-                             selectedtrait = .,
-                             selectedtaxres = "genus",
-                             selectedpolmode = allpol,
-                             selectedpft = allpft),
[…]

### 2022-12-28 03_CWM_estimates_pollen.R

@@ -21,7 +21,7 @@ set.seed(1)
 
 ## Load packages
 if (!require(tidyverse)) install.packages("tidyverse")
-if (!require(rjags))  install.packages("rjags")
+# if (!require(rjags))  install.packages("rjags")
 if (!require(runjags)) install.packages("runjags")
 if (!require(mcmc)) install.packages("mcmc")
 if (!require(coda)) install.packages("coda")
@@ -54,22 +54,22 @@ if (Sys.info()[4] == "LUIN67311") {
 }
 
 # No subset ----
-if (1) {
+if (0) {
 
-  furrr::future_map(normaltraits,
-                     ~cwm_pol(selectedtrait = .x,
-                              selectedcountry = "Scotland",
-                              selectedabun = "percent"
-                              # taxtrait = TRUE
-                              ),
-                    .options = furrr_options(seed = TRUE))
-  furrr::future_map(normaltraits,
-                     ~cwm_pol(selectedtrait = .x,
-                              selectedcountry = "Switzerland",
-                              selectedabun = "percent",
-                              taxtrait = TRUE
-                              ),
-                    .options = furrr_options(seed = TRUE))
+furrr::future_map(normaltraits,
+                   ~cwm_pol(selectedtrait = .x,
+                            selectedcountry = "Scotland",
+                            selectedabun = "percent",
+                            taxtrait = TRUE
+                            ),
+                  .options = furrr_options(seed = TRUE))
+furrr::future_map(normaltraits,
+                   ~cwm_pol(selectedtrait = .x,
+                            selectedcountry = "Switzerland",
+                            selectedabun = "percent",
+                            taxtrait = TRUE
+                            ),
+                  .options = furrr_options(seed = TRUE))
   furrr::future_map(normaltraits,
                     ~cwm_pol(selectedtrait = .x,
                                 selectedcountry = "Scotland",
@@ -90,80 +90,84 @@ if (1) {
 trsh <- c("tree", "shrub")
 herb <- c("grass", "herb")
 
+# furrr::future_map(normaltraits, 
+#                   ~cwm_pol_pft(selectedtrait = .x, 
+#                               selectedcountry = "Scotland",
+#                               selectedpft = trsh,
+#                               selectedabun = "percent"),
+#                   .options = furrr_options(seed = TRUE))
 furrr::future_map(normaltraits, 
                   ~cwm_pol_pft(selectedtrait = .x, 
-                              selectedcountry = "Scotland",
-                              selectedpft = trsh,
[…]

### 2022-12-28 cwm_functions.R

@@ -1,0 +1,960 @@
+# CWM vegetation Scotland ----
+cwm_veg_scot <- function(selectedabun, selectedtrait, selectedtaxres, selectedpolmode, 
+                         selectedpft, notinpollen = FALSE, taxtrait = FALSE){
+  ## prepare abundance data
+  abun <- lABUN %>% 
+    # select zone 
+    pluck(selectedabun) %>% 
+    # optional filter for polmode and pft
+    filter(polmode %in% all_of(selectedpolmode)) %>% 
+    filter(growthform %in% all_of(selectedpft)) %>% 
+    # optional to filter out taxa that are not in pollen data
+    {if (notinpollen) filter(., !pollentaxon == "Not in pollen data") else . } %>% 
+    # rename column to selected taxonomic resolution
+    rename(tax = all_of(selectedtaxres)) %>% 
+    filter(!tax == "Unindentified") %>% 
+    dplyr::select(sitename, tax, abun) %>% 
+    drop_na() 
+  
+  ## prepare trait data
+  # taxa in the abundance data
+  taxa <- abun %>% 
+    pull(tax) %>% 
+    unique() %>% 
+    sort
+  
+  # missing field collected trait data
+  traittax <- lTrait %>%
+    pluck(selectedtrait) %>% 
+    pull(selectedtaxres) %>% 
+    unique
+  missingtax <- taxa[!taxa %in% traittax] 
+  
+  # filter species with too little observations (10 or less)
+  nobs <- lTrait %>%
+    pluck(selectedtrait) %>%
+    arrange(selectedtaxres) %>% 
+    group_by(across(all_of(selectedtaxres))) %>%
+    summarise(n = n()) %>%
+    filter(n <= 10) %>%
+    pull(selectedtaxres)
+  
+  # find taxa with not or too little data in TRY gapfilled data
+  TRY <- dfTRY %>% 
+    dplyr::select(tax = all_of(selectedtaxres), 
+                  all_of(selectedtrait)) %>% 
+    filter(tax %in% c(missingtax, nobs)) 
+  
+  # Create trait data df with field an try data
+  trait <- lTrait %>%
+    pluck(selectedtrait) %>% 
+    dplyr::select(tax = all_of(selectedtaxres), 
+                  all_of(selectedtrait)) %>% 
+    # join with try data for the missing traits
+    bind_rows(TRY) %>% 
+    filter(tax %in% taxa) %>%
+    # sort alphabetically
+    arrange(tax) %>%
[…]

### 2022-12-28 Manuscript.Rmd
add to intro

@@
     fig_caption: yes
+    bibliography: "Validation_study_references.bib" 
 ---
@@
 
+Because of the difference in  pollen productivity and dispersal among pollen taxa, pollen counts might not reflect absolute abundance and hence trait composition [@Sande2021].
+Alternative to forming a direct representation of past plant trait composition, pollen based trait reconstruction might be better suited to study relative changes in trait composition.
+For instance, @Sande2021 showed that pollen assemblages reflect relative trait changes in trait composition of the vegetation along an elevation gradient.
+
 The second challenge in reliably reconstructing past trait composition is to attribute functional trait data to the taxa found in pollen records, as it is rarely possible to resolve the pollen record to the species level [@Birks_2020].
@@
 In this study, we use a Bayesian approach for modelling trait composition from pollen data, thereby allowing for the propagation of uncertainty in the taxon level trait reconstruction to the community level.
-
-Alternative to forming a direct representation of past plant trait composition, pollen based trait reconstruction might be better suited to study relative changes in trait composition.
-For instance, @Sande202 showed that pollen assemblages reflect relative trait changes in trait composition of the vegetation.
 
@@
 ### Study sites
-
 We collected pollen, trait and species composition data from 16 sites in the Cairngorms National Park, Scotland, United Kingdom.
-Sites were selected to represent four vegetation types in the area: grassland, heath land, deciduous forest and coniferous forest.
+We selected sites to represent four vegetation types in the area: grassland, heath land, deciduous forest and coniferous forest.
 Sampling locations were at least 1 km from each other to avoid pollen source areas from overlapping with each other.
@@
 ### Study design
-
-The data that was collected in Scotland was used to test if pollen-based trait reconstructions represents vegetation trait composition.
+We used the Scottish data to test if pollen-based trait reconstructions represents vegetation trait composition.
 We compare the community weighted mean (CWM) trait value of the vegetation with the CWM trait values reconstructed from surface pollen samples.
-Surface pollen samples are collected from moss polsters and approximate vegetation composition in recent years[@Pardoe2010].
 Study design mostly followed the protocol of @Bunting_2013, which allows for the calculation of vegetation composition in concentric circles around the moss sampling point to test the spatial scale that is represented by the moss sample (Appendix).
@@
 The vegetation cover of the outer ring (1000 m) was estimated using QGIS and land cover maps [\@scotisch_government; \@QGIS_software].
-To test the effect of spatial scale on the representativeness of pollen assemblages for trait composition reconstruction, we test compare vegetation CWM and pollen CWM in the rings of 10 m, 100 m and 1000 m.
+To test the effect of spatial scale on the representativeness of pollen assemblages for trait composition reconstruction, we compare vegetation CWM and pollen CWM in the rings of 10 m, 100 m and 1000 m.
 Furthermore, four different treatments were assigned to test what factors could influence representation of trait composition by pollen data.
@@
 ### Vegetation survey
-
-At the Scottish sites, in the inner ring (10 m) 17 sub-plots of 0.25 m^2^ were surveyed: one in the centre, 3 in every wind direction at 3, 6 and 10 meters, and one in the NE, NW, SE and SW direction (See design in Appendix).
-In the middle ring (10-100 m) between 4 and 6 plots of 5 by 5 meter were selected on a stratified random design based on vegetation types that were identified in the fields.
+At the Scottish sites, we surveyed 17 sub-plots of 0.25 m^2^ in the inner ring (10 m): one in the center, 3 in every wind direction at 3, 6 and 10 meters, and one in the NE, NW, SE and SW direction (See design in Appendix).
+In the middle ring (10-100 m), we selected between 4 and 6 plots of 5 by 5 meter based on a stratified random design based on vegetation types that were identified in the fields.
 Vegetation abundance was estimated using percentage cover.
@@
 ### Trait data collection
-
 We used three traits for the analysis, specific leaf area, leaf area and and plant height.
 These traits reflect different components in the spectrum of plant strategies [@Diaz2016].
-In Scotland, trait data was collected in the middle ring (100 m).
+In Scotland, we collected trait data in the middle ring (100 m).
 For determining plant height, we randomly selected 10 flowering individuals of each species at every site.
@@
 Samples were dried at 70 ***°***C for 72 hours to constant dry weight before measuring dry weight.
-In Switzerland, trait data was collected for the species present at the nine points in the 3 by 3 grid, otherwise trait measurement was the same.
-
-For reconstructing trait composition from pollen data and previously gapfilled trait data from the TRY database was used [@kattge2020try] [\@Schrodt_2015].
-for classifying vegetation data in to plant growth forms and pollination mode,
+In Switzerland, trait data was collected for the species present at the nine points in the 3 by 3 grid, otherwise trait measurement was the same.For reconstructing trait composition from pollen data and previously gapfilled trait data from the TRY database was used [@kattge2020try] [@Schrodt_2015].
 
 ### Pollen data
+Surface pollen samples are collected from moss polsters and approximate vegetation composition in recent years [@Pardoe2010].
+Moss was collected from multiple location within a 1 m radius in the center of each plot.Pollen were extracted from the moss using standardized pollen preparation protocol (REF).
+Pollen were identified using light microscope and the key of (REF).
 
-Moss was collected from multiple location within a 1 m radius in the centre of each plot.
-Only living tissue was sampled.
-Pollen were extracted from the moss using standardized pollen preparation protocol (REF).
-Pollen were identified using light microscope and the key of (REF).
-To test the ability of pollen correction factors to improve pollen - vegetation relationships, pollen productivity estimates from @Githumbi_2022 were used.
-Pollen productivity estimates were not available for all taxa, taxa without pollen productivity factors were excluded.
+### 
+To test the ability of pollen correction factors to improve pollen - vegetation relationships, pollen productivity estimates from @Githumbi_2022 were used.Pollen productivity estimates were not available for all taxa, taxa without pollen productivity factors were excluded.
 On average, this led to the exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%).
@@
 the lowest taxonomic name the pollen could be identified to.
-All trait data that was available for the species in the translation table was then used in the calculation trait values on the pollen taxonomic level .
+All trait data that was available for the species in the translation table was then used in the calculation trait values on the pollen taxonomic level.
+For reconstructing trait composition from pollen data and previously gapfilled trait data from the TRY database was used [@kattge2020try] [@Schrodt_2015].
 
@@
 $Y_{elev}$ is the elevation as the linear predictor.
-We modelled this relationship for the grassland and forest plots separately.
+We modeled this relationship for the grassland and forest plots separately.
 
@@
 
+### Pollen representation by the vegetation
+- missing species in pollen 
+- 
+### Effect of spatial scale and plant functional group
 Over all treatments and spatial scales, we only found evidence of a significant positive relationship between pollen-based reconstructed CWM trait values and vegetation CWM values for plant height.

### 2022-12-28 Manuscript.Rmd
method adjustments

@@
 ### Trait data collection
-We used three traits for the analysis, specific leaf area, leaf area and and plant height.
+We used three traits for the analysis, specific leaf area (SLA), leaf area (LA) and and plant height.
 These traits reflect different components in the spectrum of plant strategies [@Diaz2016].
@@
 Non-flowering individuals were selected if flowering individuals were not available.
-For determining specific leaf area (SLA) and leaf area (LA), we collected five leaf samples per species in sealable plastic bags.
+For determining specific leaf area and leaf area, we collected five leaf samples per species in sealable plastic bags.
 Fresh weight and surface area were measured within 72 hours of collection using a carat scale and flatbed scanner.
 Samples were dried at 70 ***°***C for 72 hours to constant dry weight before measuring dry weight.
-In Switzerland, trait data was collected for the species present at the nine points in the 3 by 3 grid, otherwise trait measurement was the same.For reconstructing trait composition from pollen data and previously gapfilled trait data from the TRY database was used [@kattge2020try] [@Schrodt_2015].
+In Switzerland, trait data was collected for the species present at the nine points in the 3 by 3 grid, otherwise trait measurement was the same.
+For reconstructing trait composition from pollen data, we did not use the data collected in the field but previously gapfilled trait data from the TRY database was used [@kattge2020try] [@Schrodt_2015].
 
-### Pollen data
+### Pollen data collection
 Surface pollen samples are collected from moss polsters and approximate vegetation composition in recent years [@Pardoe2010].
@@
 
-### 
-To test the ability of pollen correction factors to improve pollen - vegetation relationships, pollen productivity estimates from @Githumbi_2022 were used.Pollen productivity estimates were not available for all taxa, taxa without pollen productivity factors were excluded.
-On average, this led to the exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%).
-Pollen were divided in to two plant functional groups according to @Reitalu_2019 consisting of woody (trees and shrubs) or non-woody species (grasses, herbs and dwarf shrubs of the Ericaceae family).
-@Reitalu_2019 was also used to classify pollen in to pollination mode, either insect-pollinated or not insect-pollinated.
-
-For reconstructing plant trait composition from pollen data, assumptions needed to be made about the species belonging to the pollen taxa in the data set.
+For reconstructing plant trait composition from pollen data, assumptions need to be made about the species belonging to the pollen taxa in the data set.
 Therefore we created a pollen to species translation table.
@@
 All trait data that was available for the species in the translation table was then used in the calculation trait values on the pollen taxonomic level.
-For reconstructing trait composition from pollen data and previously gapfilled trait data from the TRY database was used [@kattge2020try] [@Schrodt_2015].
 
 ### Calculation of community weighted mean
-
-We use the community weighted mean (CWM) to characterize functional composition, which here is specified as the trait distribution of a taxon, i.e.
+We use the community weighted mean (CWM) to characterize functional composition.
+Here CWM is specified as the trait distribution of a taxon, i.e.
 the species in the vegetation data and the pollen taxon in the pollen data, weighted by its abundance.
 First, the trait distribution on the species or the pollen taxonomic level was estimated.
-We assume that the trait data of SLA, LA and plant height follow a lognormal distribution for the likelihood, so that$$trait_{i}^{observed} \sim { lognormal(\mu_{j},\sigma_j)}$$Where *i* represents each trait observation of a taxon (*j*) from the trait data.
+We assume that the trait data of SLA, LA and plant height follow a lognormal distribution for the likelihood, so that
+$$trait_{i}^{observed} \sim { lognormal(\mu_{j},\sigma_j)}$$
+Where *i* represents each trait observation of a taxon (*j*) from the trait data.
 $\mu_{j}$ and $\sigma_{j}$ are the mean trait value and the standard deviation per taxon, respectively.
 Vague priors were used for the taxon mean and standard deviation estimation.
-We then modelled the community weighted mean (CWM) as a mixture of normal distributions that were weighted by the species abundance or pollen percentage, so that
+We then modeled the CWM as a mixture of normal distributions that were weighted by the species abundance or pollen percentage, so that
 
@@
 Where *n* is the total number of taxa, and $\pi_{j}$ the abundance of each pollen taxon or species at a given site *s*.
-The mean ($CWM_{mean}$) and the standard deviation ($CWM_{sd}$)of this modelled distribution was subsequently used in the estimation of the correlation between the trait composition of the vegetation and the pollen data.
+The mean ($CWM_{mean}$) and the standard deviation ($CWM_{sd}$) of this modeled distribution was subsequently used in the estimation of the correlation between the trait composition of the vegetation and the pollen data.
 
 ### Relationship between pollen and vegetation community weighted mean
-
 To test the relationship between the pollen based reconstruction of CWM and the vegetation CWM, we standardized the CWM values and fitted a students-t distributed likelihood for pollen CWM ($CWM_{mean}^{pol}$) with vegetation CWM ($CWM_{mean}^{veg}$) as the single predictor so that,
@@
 
-The slope $\beta_{t}$ is a correlation-like metric defined with a flat, uninformative prior. $$
+The slope $\beta_{t}$ is a correlation-like metric defined with a flat, uninformative prior. 
+$$
 \beta \sim Uniform(-1,1)
@@
 
-We ran this model for every trait and zone in Scotland, and for the following different treatments: the use of correction factors, sub-setting the data in plant functional type, sub-setting the data in pollination mode and modelling on different taxonomic resolutions
+### Treatments 
+To test the ability of pollen correction factors to improve pollen - vegetation relationships, we divided pollen counts by pollen productivity estimates.
+We used pollen productivity estimates (ppe's) from @Githumbi_2022 who provides ppe's and uncertainty estimates for 39 taxa in Europe.
+Pollen productivity estimates were not available for all taxa in the pollen data, taxa without pollen productivity factors were excluded.
+On average, this led to the exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%).
+
+To test for differences in the representation of particular plant functional groups by pollen assemblages, pollen were divided in to plant functional groups.
+We used the classification of @Reitalu_2019 who distinguished between woody (trees and shrubs) and non-woody species (grasses, herbs and dwarf shrubs of the Ericaceae family).
+@Reitalu_2019 was also used to classify pollen in to pollination mode, either insect-pollinated or not insect-pollinated.
+To classify the vegetation in to plant functional groups and pollination mode, data from the TRY database was used [@kattge2020try].
+
+To test for the effect of taxonomic resolution on the trait reconstructions, we calculated CWM trait values of the vegetation using all available trait data of the species, genus or family.
 
 ### Community weighted mean changes with elevation
-
-To test if trait compositional changes of the vegetation along an elevation gradient are reflected by pollen based trait reconstructions, we fitted a generalized linear model with elevation as the predictor, so that the likelihood of the CWM value is defined as follows,
+To test if trait composition changes of the vegetation along an elevation gradient are reflected by pollen based trait reconstructions, we fitted a generalized linear model with elevation as the predictor, so that the likelihood of the CWM value is defined as follows,
 

### 2022-12-28 Manuscript.Rmd
start results

@@
 - missing species in pollen 
-- 
-### Effect of spatial scale and plant functional group
+- spatial scale
+
+### Treatments
 Over all treatments and spatial scales, we only found evidence of a significant positive relationship between pollen-based reconstructed CWM trait values and vegetation CWM values for plant height.

### 2022-12-28 01_Swiss_data_compilation.R

@@ -13,8 +13,8 @@
 ## ---------------------------
 ##
 ## Notes:
-##   - missing plot numbers
-##   - missing trait data
+##   
+##   
 ##
 ## ---------------------------
 
@@ -113,7 +113,8 @@ bdm_sp <- bdm_sp %>%
                                                    "dead wood", "dead trunk",
                                                    "stump", "fungi", "rock",
                                                    "root", "other", "bareground") ~ "No plants",
-                           str_detect(species_code, pattern = "[:digit:]") ~ "Unindentified"
+                           (str_detect(species_code, pattern = "[:digit:]") & !is.na(binom)) ~ binom,
+                           (str_detect(species_code, pattern = "[:digit:]") & is.na(binom)) ~ "Unindentified"
                                 )
          ) %>% 
   # select relevant columns, rename

### 2023-01-24 01_PFT_table_vegetation.R
add helinger transformation

@@ -79,7 +79,9 @@ pft <- traits %>%
   filter(!(AccSpeciesName == "Dryopteris" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Empetrum nigrum" & growthform == "shrub")) %>% 
   filter(!(AccSpeciesName == "Equisetum sylvaticum" & growthform == "shrub")) %>% 
+  filter(!(AccSpeciesName == "Equisetum" & growthform == "fern")) %>% 
   filter(!(AccSpeciesName == "Galium rotundifolium" & growthform == "shrub")) %>% 
+  filter(!(AccSpeciesName == "Holcus lanatus" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Equisetum sylvaticum" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Juncus bufonius" & growthform == "tree")) %>% 
   filter(!(AccSpeciesName == "Juniperus communis" & growthform == "tree")) %>% 
@@ -88,6 +90,7 @@ pft <- traits %>%
   filter(!(str_detect(AccSpeciesName,"Luzula") & growthform == "shrub")) %>% 
   filter(!(AccSpeciesName == "Luzula multiflora" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Milium effusum" & growthform == "herb")) %>% 
+  filter(!(AccSpeciesName == "Molinia caerulea" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Nardus stricta" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Orthilia secunda" & growthform == "shrub")) %>% 
   filter(!(str_detect(AccSpeciesName, "Plantago") & growthform == "tree")) %>% 
@@ -101,6 +104,7 @@ pft <- traits %>%
   filter(!(str_detect(AccSpeciesName, "Rumex") & growthform == "tree")) %>% 
   filter(!(AccSpeciesName == "Stellaria media" & growthform == "tree")) %>% 
   filter(!(AccSpeciesName == "Stellaria media" & growthform == "shrub")) %>% 
+  filter(!(AccSpeciesName == "Salix" & growthform == "tree")) %>% 
   filter(!(AccSpeciesName == "Trichophorum cespitosum" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Trifolium repens" & growthform == "tree")) %>% 
[…]

### 2023-01-24 01_Prepare_pollen_data.R
add Hellinger transformation

@@ -19,10 +19,10 @@
 
 options(scipen = 999)
 set.seed(1)
+
 ## Load packages
 if (!require(tidyverse)) install.packages("tidyverse")
 if (!require(readxl)) install.packages("readxl")
-
 
 # Scotland ----
 pollen_raw <- read.csv("Data/Pollen_count_Scotland-checked.csv", skip = 1, sep = ";")
@@ -46,8 +46,6 @@ dfPOL <- pollen_raw %>%
          count = as.numeric(count)) %>% 
   # remove zeros
   filter(!count == 0) %>% 
-  mutate(pollentaxon = recode(pollentaxon, Tsuga = "unindentified", 
-                              Liliaceae = "unindentified", Drosera = "unindentified")) %>%  # probably misidentified
   #correct pollen taxon names
   mutate(pollentaxon = 
            # remove "-type" to facilitate GBIF search
@@ -75,7 +73,9 @@ dfPOL_nc <- dfPOL %>%
   group_by(sitename) %>% 
   mutate(percent = count/sum(count, na.rm = TRUE)) %>%
   group_by(sitename, pollentaxon) %>%
-  summarise(percent = sum(percent)) 
+  summarise(percent = sum(percent)) %>% 
+  # Helinger transformation
+  mutate(adjusted_helinger = sqrt(percent))
 
 # make function to continue draw from rnorm till RPP is positive (otherwise negative percentages)
 norm_positive <- function(RPP, RPP_sd){
@@ -97,7 +97,7 @@ dfPOL_cor <- dfPOL %>%
   mutate(mean = count/RPP) %>%  
   # correction with random draw from RPP distribution
   rowwise() %>%
-  mutate(col_rnorm = list(setNames(count/replicate(n,norm_positive(RPP, RPP_sd)), paste0("draw",1:n)))) %>%
+  mutate(col_rnorm = list(setNames(count/replicate(n, norm_positive(RPP, RPP_sd)), paste0("draw",1:n)))) %>%
   unnest_wider(col_rnorm) %>% 
   # calculate percentages
   group_by(sitename) %>% 
@@ -124,8 +124,6 @@ dfPOL <- read.csv("Data/Pollen_count_Switserland-checked.csv", skip = 1, sep = "
                values_to = "count") %>%
   mutate(count = as.numeric(count)) %>% # change variable types
[…]

### 2023-01-24 01_Prepare_trait_data_vegetation.R
remove unused traits

@@ -22,7 +22,6 @@ library(readxl)
 ## Load data ----
 plantheight_raw <- read_xlsx("Data/Plant_height.xlsx")
 sla_ldmc_raw <- read_xlsx("Data/LDMC_SLA_dataset.xlsx")
-cn_raw <- read_xlsx("Data/CN_dataset_150.xlsx")
 harm <- read_xlsx("Data/Harmonization_table_species_names.xlsx") %>% 
   rename(family = fam) %>% 
   mutate(stand.spec = str_remove(stand.spec, " sp."))
@@ -56,7 +55,7 @@ saveRDS(plantheight, "RDS_files/01_PlantHeight.rds")
 sla <- sla_ldmc_raw %>% 
   # Harmonize column names
   dplyr::select(sitename = Site, species = Species,
-          SLA = `SLA (mm^2 mg−1)`) %>% 
+          SLA = 'SLA (mm^2 mg−1)') %>% 
   # create measurement ID
   group_by(sitename, species) %>% 
   mutate(measurementID = row_number()) %>% 
@@ -66,22 +65,6 @@ sla <- sla_ldmc_raw %>%
   left_join(harm, by = "species") %>% 
   filter(!is.na(SLA)) # check notes from raw data file for reason why missing
 saveRDS(sla, "RDS_files/01_SLA.rds")
-
-## LDMC ----
-ldmc <- sla_ldmc_raw %>% 
-  dplyr::select(sitename = Site, species = Species,
-         LDMC = "LDMC (mg g–1)") %>% 
-  distinct(sitename, species, .keep_all = TRUE) %>%  # ldmc has one measurement per species per site
-  # create measurement ID
-  group_by(sitename, species) %>% 
-  mutate(measurementID = row_number(),
-         LDMC = LDMC/1000) %>% # convert from mg/g to g/g 
-  # sort
-  arrange(sitename, species, measurementID) %>% 
-  # merge with species harmonisation list
-  left_join(harm, by = "species") %>% 
-  filter(!is.na(LDMC)) # check notes from raw data file for reason why missing
-saveRDS(ldmc, "RDS_files/01_LDMC.rds")
 
 ## LA ----
 la <- sla_ldmc_raw %>% 
@@ -99,53 +82,29 @@ la <- sla_ldmc_raw %>%
   mutate(LA = LA/100)
 saveRDS(la, "RDS_files/01_LA.rds")
     
-## C ----
[…]

### 2023-01-24 01_Species_abundance.Rmd
update from macbook
remove swiss bc

@@
   # create unique id for plot in site
-  group_by(sitename, direction, distance) %>% 
-  mutate(plot = cur_group_id())
+  group_by(sitename, species) %>%
+  # average by 10 m
+  summarise(abun = mean(abun))
 
@@
 ```
-
-# Switzerland
-
-## QGIS
-
-Land cover data was downloaded from:
-<https://www.bfs.admin.ch/bfs/en/home/services/geostat/swiss-federal-statistics-geodata/land-use-cover-suitability.html>
-
-To decrease file size, I only selected relevant columns: the one
-containing coordinates and land cover data using the following code:
-
-```{r eval = FALSE}
-# ag_b_00_03_37_area_csv <- read_delim(
-#     "C:/Users/lgxgv/OneDrive - The University of Nottingham/Data analysis/PhD/Validation-study/Data/Swiss_LC/ag-b-00.03-37-area-csv.csv",+delim = ";",
-#     escape_double = FALSE,
-#     trim_ws = TRUE
-#   )
-# 
-# LC <- ag_b_00_03_37_area_csv %>%
-#   dplyr::select(RELI, E_COORD = E, N_COORD = N, LC18_6, LC18_27)
-# 
-# write.csv(LC, file = "C:/Users/lgxgv/OneDrive - The University of Nottingham/Data
-#           analysis/PhD/Validation-study/Data/Swiss_LC/Swiss_LC18.csv")
-
-```
-
-In this file I found the CRS:
-<https://www.bfs.admin.ch/bfs/en/home/services/geostat/swiss-federal-statistics-geodata/land-use-cover-suitability/swiss-land-use-statistics.assetdetail.20104775.html>\
-"Les coordonnées « E_COORD » et « N_COORD » des hectares dans le système
-de coordonnées suisse MN95 (7 chiffres). Les coordonnées représentent le
-point d'échantillonnage correspondant (intersections des coordonnées
-hectométriques)."
-
-Import Swiss land cover (100m resolution) Swiss_LC18.csv
-
--   Set CRS code to EPSG:4150 Rasterize this layer using the GDAL
-    rasterize (vector to raster) tool
--   Set burnin to LC18_27
--   Horizontal and vertical resolution to 100 m
--   export and set resolution to 25 m for merge with forest cover data
-
-Import forest cover data in ArcGIS pro reproject to EPSG:4150 and export
-raster(folder called reproj)/ Import to QGIS set layer crs to EPSG:4150
-
-Import coordinates of pollen plots (Pollen_site_coord.csv).
-
-Create buffer zone of 100 m and 1000 m
-
-Use Zonal histogram to compute pixels in zone attributed to certain land
-cover type Export as Veg_cover_zoneC.csv and Veg_cover_zoneB.csv
-
-Import coordinate of all BDM sites Use Sample raster values to compute
-landcover types for vegetation plots. Export as Veg_type_plots_csv
-
-```{r}
-zone_b_raw <- read.csv("Data/Swiss_LC/FC_LC_cover_zoneB_swiss.csv")
-zone_c_raw <- read.csv("Data/Swiss_LC/FC_LC_cover_zoneC_swiss.csv")
-LC_labels <- read_xlsx("Data/Swiss_LC/be-b-00.03-37-area-var-v01.xlsx",
-                       sheet = "Codes LC_27")
-FC_labels <- read.csv("Data/Swiss_LC/Forest_cover_labels.csv") %>% 
-  rename(CODE = category)
-LC_labels <- LC_labels %>% 
-  dplyr::select(CODE, LABEL_DE) %>% 
-  bind_rows(FC_labels)
-
-zone_b <- zone_b_raw %>% 
-  # make tidy
-  dplyr::rename(sitename = aID_STAO,lon = X, lat = Y) %>% 
-  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
-  # rename veg_types
-  mutate(veg_type = str_remove(veg_type, "HISTO_") %>% as.numeric) %>% 
-  filter(!value == 0) %>% 
-  left_join(LC_labels, by = c("veg_type" = "CODE")) %>% 
-  # filter land cover type that are not vegation
-  filter(!LABEL_DE %in% c( "Lockergestein", "Wasser","Gebäude",
-                           "Versteinte Flächen","Befestigte Flächen",
-                           "Gemischte Kleinstrukturen")) %>%
-  # calculate veg_type percentage
-  group_by(sitename) %>% 
-  # calculate percentages
-  mutate(veg_cov = value/sum(value)) %>%
-  group_by(sitename, LABEL_DE) %>%
-  summarise(veg_cov = sum(veg_cov, na.rm = TRUE))
-
-zone_c <- zone_c_raw %>% 
-  # make tidy
-  dplyr::rename(sitename = aID_STAO,lon = X, lat = Y) %>% 
-  pivot_longer(cols = contains("HISTO"), names_to = "veg_type") %>% 
-  # rename veg_types
-  mutate(veg_type = str_remove(veg_type, "HISTO_") %>% as.numeric) %>% 
-  filter(!value == 0) %>% 
-  left_join(LC_labels, by = c("veg_type" = "CODE")) %>% 
-  # filter land cover type that are not vegation
-  filter(!LABEL_DE %in% c( "Lockergestein", "Wasser","Gebäude",
-                           "Versteinte Flächen","Befestigte Flächen",
-                           "Gemischte Kleinstrukturen")) %>%
-  # calculate veg_type percentage
-  group_by(sitename) %>% 
-  # calculate percentages
-  mutate(veg_cov = value/sum(value)) %>%
-  group_by(sitename, LABEL_DE) %>%
-  summarise(veg_cov = sum(veg_cov, na.rm = TRUE))
-
-# merge zone b and c
-veg_cover <- zone_b %>% 
-  full_join(zone_c, by = c("sitename", "LABEL_DE"), 
-            suffix = c("_b","_c")) %>% 
-  replace_na(list(veg_cov_b = 0))
-```
-
-```{r join veg type data and species data}
-# Veg type of BDM sites
-veg_type <- read.csv("Data/Swiss_LC/FC_LC_type_swiss.csv")
-
-# Join with labels
-veg_type <- veg_type %>% 
-  left_join(LC_labels, by = c("SAMPLE_1" = "CODE"))
-
-# Load in species data
-bdm_sp <- readRDS("RDS_files/01_Species_abundance_Swiss.rds")
-
-bdm_sp <- bdm_sp %>% 
-  # filter stuff that is not a plant
-  filter(!stand.spec == "No plants") %>% 
-  # filter unindentified
-  filter(!stand.spec == "Unindentified") %>% 
-  # join with vegetation type data
-  left_join(veg_type, by = c("sitename" = "aID_STAO")) 
-# create table with genus and family to join after summarizing
-genus_fam <- bdm_sp %>% 
-  ungroup() %>% 
-  dplyr::select(family, genus, stand.spec) %>% 
-  distinct
-```
-
-```{r}
-spec_cover <- bdm_sp %>% 
-  # calculate average species cover per vegetation type using the plots belonging to that vegetation type
-  group_by(sitename, LABEL_DE, stand.spec) %>% 
-  summarise(abun = mean(abun)) %>% 
-  # join with vegetation cover data
-  right_join(veg_cover, by = c("sitename", "LABEL_DE")) %>% 
-  # weight species abundance by veg cover
-  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c)
-
-# print table with missing species data for vegetation types
-missing_vegtype <- spec_cover %>% 
-  filter(is.na(stand.spec)) %>% 
-  group_by(sitename, LABEL_DE) %>% 
-  summarise(cov_b = round(sum(veg_cov_b), 2), 
-            cov_c = round(sum(veg_cov_c), 2)) %>%
-  filter(!cov_b == 0 |!cov_c == 0)
-
-# for missing data, use species data from other plots
-missing <- bdm_sp %>% 
-  filter(LABEL_DE %in% missing_vegtype$LABEL_DE) %>% 
-  group_by(LABEL_DE, stand.spec) %>% 
-  summarise(abun = mean(abun))
-spec_cover_na <- spec_cover %>% 
-  filter(is.na(stand.spec)) %>%
-  dplyr::select(sitename, LABEL_DE, veg_cov_b, veg_cov_c) %>% 
-  left_join(missing, by = "LABEL_DE") %>% 
-  mutate(spec_abun_b = abun*veg_cov_b, spec_abun_c = abun*veg_cov_c) 
-spec_cover_bc <-  spec_cover %>% 
-  filter(!is.na(stand.spec)) %>%
-  bind_rows(., spec_cover_na) %>% 
-  arrange(sitename, LABEL_DE, stand.spec) %>% 
-  dplyr::select(sitename, stand.spec, spec_abun_b, spec_abun_c) %>% 
-  ungroup() %>% 
-  left_join(genus_fam, by = "stand.spec")
-
-saveRDS(spec_cover_bc, "RDS_files/01_Species_abundance_zoneBC_swiss.rds")
-```

### 2023-01-24 02_CWM_estimates_vegetation.R
remove unused traits

@@ -19,14 +19,13 @@
 
 set.seed(1)
 
-
 ## Load packages
 if (!require(tidyverse)) install.packages("tidyverse")
 if (!require(rjags))  install.packages("rjags")
 if (!require(runjags)) install.packages("runjags")
 if (!require(mcmc)) install.packages("mcmc")
 if (!require(coda)) install.packages("coda")
-if (!require(furrr))  install.packages("furrr")
+if (!require(furrr)) install.packages("furrr")
 
 ## Load functions
 source("cwm_functions.R")
@@ -36,8 +35,6 @@ source("cwm_functions.R")
 # Scotland ----
 plantheight <- readRDS("RDS_files/01_PlantHeight.rds") %>% 
   ungroup() 
-ldmc <- readRDS("RDS_files/01_LDMC.rds") %>% 
-  ungroup()
 sla <- readRDS("RDS_files/01_SLA.rds") %>% 
   ungroup()
 la <- readRDS("RDS_files/01_LA.rds") %>% 
@@ -45,18 +42,18 @@ la <- readRDS("RDS_files/01_LA.rds") %>%
 
 
 lTrait <- list(PlantHeight = plantheight,
-               LDMC = ldmc,
                SLA = sla,
                LA = la)
 
 dfTRY_raw <- readRDS("RDS_files/02_Gapfilled_traits.rds")
 
 # Pollination and PFT ----
-dfPFT <- readRDS("RDS_files/Polmode_pft_vegetation.rds")
+dfPFT <- readRDS("RDS_files/01_Polmode_pft_vegetation.rds")
 
 # Species to pollen table
 polspec <- readRDS("RDS_files/02_PollenType_species.rds") %>% 
-  dplyr::select(pollentaxon, stand.spec) %>% distinct() %>% as_tibble()
+  dplyr::select(pollentaxon, stand.spec) %>% distinct() %>% as_tibble() %>% 
+  drop_na()
 
 # Species abundance data ----
 dfABUN_a <- readRDS("RDS_files/01_Species_abundance_a.rds") 
@@ -94,10 +91,10 @@ lABUN <- list("zoneA" = zoneA, "zoneB" = zoneB, "zoneC" = zoneC)
 
 trsh <- c("tree", "shrub")
 herb <- c("herb", "grass", "fern")
-allpft  <- c("tree","herb","grass","fern","shrub", NA)
+allpft  <- c("tree","herb","grass","fern","shrub")
[…]

### 2023-01-24 01_PFT_table_vegetation.R
fix duplication

@@ -79,9 +79,7 @@ pft <- traits %>%
   filter(!(AccSpeciesName == "Dryopteris" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Empetrum nigrum" & growthform == "shrub")) %>% 
   filter(!(AccSpeciesName == "Equisetum sylvaticum" & growthform == "shrub")) %>% 
-  filter(!(AccSpeciesName == "Equisetum" & growthform == "fern")) %>% 
   filter(!(AccSpeciesName == "Galium rotundifolium" & growthform == "shrub")) %>% 
-  filter(!(AccSpeciesName == "Holcus lanatus" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Equisetum sylvaticum" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Juncus bufonius" & growthform == "tree")) %>% 
   filter(!(AccSpeciesName == "Juniperus communis" & growthform == "tree")) %>% 
@@ -90,7 +88,6 @@ pft <- traits %>%
   filter(!(str_detect(AccSpeciesName,"Luzula") & growthform == "shrub")) %>% 
   filter(!(AccSpeciesName == "Luzula multiflora" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Milium effusum" & growthform == "herb")) %>% 
-  filter(!(AccSpeciesName == "Molinia caerulea" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Nardus stricta" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Orthilia secunda" & growthform == "shrub")) %>% 
   filter(!(str_detect(AccSpeciesName, "Plantago") & growthform == "tree")) %>% 
@@ -104,7 +101,6 @@ pft <- traits %>%
   filter(!(str_detect(AccSpeciesName, "Rumex") & growthform == "tree")) %>% 
   filter(!(AccSpeciesName == "Stellaria media" & growthform == "tree")) %>% 
   filter(!(AccSpeciesName == "Stellaria media" & growthform == "shrub")) %>% 
-  filter(!(AccSpeciesName == "Salix" & growthform == "tree")) %>% 
   filter(!(AccSpeciesName == "Trichophorum cespitosum" & growthform == "herb")) %>% 
   filter(!(AccSpeciesName == "Trifolium repens" & growthform == "tree")) %>% 
[…]

### 2023-01-24 02_PollenType_Species_GBIF.R
add missing package

@@ -25,6 +25,7 @@
 ## Load packages
 if (!require(tidyverse)) install.packages("tidyverse")
 if (!require(rgbif)) install.packages("rgbif")
+if (!require(devtools)) install.packages("devtools")
 if (!require(LCVP)) devtools::install_github("idiv-biodiversity/LCVP")
 if (!require(lcvplants)) devtools::install_github("idiv-biodiversity/lcvplants")
 

### 2023-01-24 Manuscript.Rmd
resolve merge conflict

--- 2022-12-28 Manuscript.Rmd
@@
     fig_caption: yes
-    bibliography: "Validation_study_references.bib" 
+  bibliography: Validation_study_references.bib
 ---
@@
 # Libraries
+library(tidyverse)
 library(bibtex)
@@
 Functional traits provide a better insight in to the processes that drive biodiversity and how biodiversity affects ecosystem functioning [@Hevia_2017].
-While the functioning of modern ecosystems is regularly contingent on past human and environmental factors [@Levis_2018; @Jogiste_2017; @Blonder_2018], the observational duration of trait-based studies is generally at most a couple of decades [@Estes_218].
+While the functioning of modern ecosystems is regularly contingent on past human and environmental factors [@Levis_2018; @Jogiste_2017; @Blonder_2018; ], the observational duration of trait-based studies is generally at most a couple of decades [@Estes_218].
 So to extend the time scales of ecosystem functioning research, considerable interest in the trait-based approach has arisen in palaeoecology [@Brussel_2018; @Blaus_2020; @Carvalho_2019; @Reitalu_2015; @Sande_2019].
@@
 bunting 2003, bunting 2022).
-Over- or underrepresentation can be corrected for with the use of taxon-specific pollen productivity estimates, these correction factors down-weight the contribution of dominant taxa [@Githumbi_2022].
+Over- or under-representation can be corrected for with the use of taxon-specific pollen productivity estimates, these correction factors down-weight the contribution of dominant taxa [@Githumbi_2022].
 Furthermore, it has been suggested by @Blaus_2020 to treat woody and non-woody taxa separately in the reconstruction of functional composition from pollen data as their representative pollen source area differs greatly.
 
-Because of the difference in  pollen productivity and dispersal among pollen taxa, pollen counts might not reflect absolute abundance and hence trait composition [@Sande2021].
-Alternative to forming a direct representation of past plant trait composition, pollen based trait reconstruction might be better suited to study relative changes in trait composition.
-For instance, @Sande2021 showed that pollen assemblages reflect relative trait changes in trait composition of the vegetation along an elevation gradient.
-
-The second challenge in reliably reconstructing past trait composition is to attribute functional trait data to the taxa found in pollen records, as it is rarely possible to resolve the pollen record to the species level [@Birks_2020].
+Perhaps the most challenging in reliably reconstructing past trait composition is to attribute functional trait data to the taxa found in pollen records, as it is rarely possible to resolve the pollen record to the species level [@Birks_2020].
 Trait variation within taxa can be substantial depending on the trait and the diversity of the taxon [@Albert_2010; @Prinzing_2008].
@@
 @Blaus_2020 and @Carvalho_2019 proposed that the high species number included in herbaceous pollen taxon could be one of the reasons why herbaceous vegetation was less well represented by pollen based functional diversity reconstructions.
+In both studies, however, traits values of pollen taxa were calculated by averaging over the available species level trait information.
 In this study, we use a Bayesian approach for modelling trait composition from pollen data, thereby allowing for the propagation of uncertainty in the taxon level trait reconstruction to the community level.
@@
 ### Study sites
+
 We collected pollen, trait and species composition data from 16 sites in the Cairngorms National Park, Scotland, United Kingdom.
-We selected sites to represent four vegetation types in the area: grassland, heath land, deciduous forest and coniferous forest.
+Sites were selected to represent four vegetation types in the area: grassland, heath land, deciduous forest and coniferous forest.
 Sampling locations were at least 1 km from each other to avoid pollen source areas from overlapping with each other.
@@
 This is a systematic sampling grid for the long-term monitoring of plant and animal species in Switzerland [@BDM_2022].
-Trait and species composition data, but not pollen data, from an additional `r nrow(read.csv("Data/Swiss_LC/all_BDM_plots_coord.csv")) - nrow(bdm_meta)` sites that were surveyed in this network were also used in this study.
+Trait and species composition data from an additional `r nrow(read.csv("Data/Swiss_LC/all_BDM_plots_coord.csv")) - nrow(bdm_meta)` sites that were surveyed in this network were also used in this study.
 
 ### Study design
-We used the Scottish data to test if pollen-based trait reconstructions represents vegetation trait composition.
+
+The data that was collected in Scotland was used to test if pollen-based trait reconstructions represents vegetation trait composition.
 We compare the community weighted mean (CWM) trait value of the vegetation with the CWM trait values reconstructed from surface pollen samples.
+Surface pollen samples are collected from moss polsters and approximate vegetation composition in recent years [@Birks_2016] [+Other ref?].
 Study design mostly followed the protocol of @Bunting_2013, which allows for the calculation of vegetation composition in concentric circles around the moss sampling point to test the spatial scale that is represented by the moss sample (Appendix).
@@
 The vegetation cover of the outer ring (1000 m) was estimated using QGIS and land cover maps [\@scotisch_government; \@QGIS_software].
-To test the effect of spatial scale on the representativeness of pollen assemblages for trait composition reconstruction, we compare vegetation CWM and pollen CWM in the rings of 10 m, 100 m and 1000 m.
+To test the effect of spatial scale on the representativeness of pollen assemblages for trait composition reconstruction, we test compare vegetation CWM and pollen CWM in the rings of 10 m, 100 m and 1000 m.
 Furthermore, four different treatments were assigned to test what factors could influence representation of trait composition by pollen data.
@@
 ### Vegetation survey
-At the Scottish sites, we surveyed 17 sub-plots of 0.25 m^2^ in the inner ring (10 m): one in the center, 3 in every wind direction at 3, 6 and 10 meters, and one in the NE, NW, SE and SW direction (See design in Appendix).
-In the middle ring (10-100 m), we selected between 4 and 6 plots of 5 by 5 meter based on a stratified random design based on vegetation types that were identified in the fields.
+
+At the Scottish sites, in the inner ring (10 m) 17 sub-plots of 0.25 m^2^ were surveyed: one in the centre, 3 in every wind direction at 3, 6 and 10 meters, and one in the NE, NW, SE and SW direction (See design in Appendix).
+In the middle ring (10-100 m) between 4 and 6 plots of 5 by 5 meter were selected on a stratified random design based on vegetation types that were identified in the fields.
 Vegetation abundance was estimated using percentage cover.
@@
 ### Trait data collection
-We used three traits for the analysis, specific leaf area, leaf area and and plant height.
-These traits reflect different components in the spectrum of plant strategies [@Diaz2016].
-In Scotland, we collected trait data in the middle ring (100 m).
-For determining plant height, we randomly selected 10 flowering individuals of each species at every site.
-Non-flowering individuals were selected if flowering individuals were not available.
-For determining specific leaf area (SLA) and leaf area (LA), we collected five leaf samples per species in sealable plastic bags.
+
+Three traits were selected for the analysis, specific leaf area, leaf area and and plant height.
+These traits reflect different components in the spectrum of plant strategies.
+XXX.
+In Scotland, trait data was collected in the middle ring (100 m).
+For determining plant height 10 flowering of individuals were selected at every site if present, non-flowering individuals were selected if flowering individuals were not available.
+Five leaf samples per species for specific leaf area (SLA) and leaf area (LA) collected in seal-able plastic bags.
 Fresh weight and surface area were measured within 72 hours of collection using a carat scale and flatbed scanner.
 Samples were dried at 70 ***°***C for 72 hours to constant dry weight before measuring dry weight.
-In Switzerland, trait data was collected for the species present at the nine points in the 3 by 3 grid, otherwise trait measurement was the same.For reconstructing trait composition from pollen data and previously gapfilled trait data from the TRY database was used [@kattge2020try] [@Schrodt_2015].
+In Switzerland, the trait data was collected for species present at the nine points in the 3 by 3 grid.
+XXXX
+
+For reconstructing trait composition from pollen data and for classifying vegetation data in to plant growth forms and pollination mode, previously gapfilled trait data from the TRY database was used [@kattge2020try] [\@Schrodt_2015].
+XXX
 
 ### Pollen data
-Surface pollen samples are collected from moss polsters and approximate vegetation composition in recent years [@Pardoe2010].
-Moss was collected from multiple location within a 1 m radius in the center of each plot.Pollen were extracted from the moss using standardized pollen preparation protocol (REF).
+
+Moss was collected from multiple location within a 1 m radius in the centre of each plot.
+Only living tissue was sampled.
+Pollen were extracted from the moss using standardized pollen preparation protocol (REF).
 Pollen were identified using light microscope and the key of (REF).
-
-### 
-To test the ability of pollen correction factors to improve pollen - vegetation relationships, pollen productivity estimates from @Githumbi_2022 were used.Pollen productivity estimates were not available for all taxa, taxa without pollen productivity factors were excluded.
+To test the ability of pollen correction factors to improve pollen - vegetation relationships, pollen productivity estimates from @Githumbi_2022 were used.
+Pollen productivity estimates were not available for all taxa, taxa without pollen productivity factors were excluded.
 On average, this led to the exclusion of `r round(mean(missing_ppe_scot$percent_missing),2)*100`% of the Scottish pollen data (max `r round(max(missing_ppe_scot$percent_missing),2)*100`%).
@@
 the lowest taxonomic name the pollen could be identified to.
-All trait data that was available for the species in the translation table was then used in the calculation trait values on the pollen taxonomic level.
-For reconstructing trait composition from pollen data and previously gapfilled trait data from the TRY database was used [@kattge2020try] [@Schrodt_2015].
+All trait data that was available for the species in the translation table was then used in the calculation trait values on the pollen taxonomic level .
 
@@
 $Y_{elev}$ is the elevation as the linear predictor.
-We modeled this relationship for the grassland and forest plots separately.
+We modelled this relationship for the grassland and forest plots separately.
 
@@
 
-### Pollen representation by the vegetation
-- missing species in pollen 
-- 
-### Effect of spatial scale and plant functional group
 Over all treatments and spatial scales, we only found evidence of a significant positive relationship between pollen-based reconstructed CWM trait values and vegetation CWM values for plant height.
@@
 
-Pollen assemblages are the main proxy for studying past vegetation features, and are a reliable proxy for plant species richness and XX.
-However the relationship with vegetation trait composition is complex.
-Here we find little evidence for pollen being a reliable proxy for plant trait composition.but the relationship between pollen assemblages and the vegetation composition is complex Exploring the possible causes for the lack of a relationship between pollen CWM values and vegetation CWM values, we suggest that this either has to do with the representation of the vegetation by the pollen data (abundance, species richness) or the large trait variation within pollen types.
-
-vegetation structure (Switzerland) and scale (Scotland)
+Pollen are the main proxy for reconstructing past plant composition, but the relationship between pollen assemblages and the vegetation composition is complicated.
+Here the relationship is further complicated by throwing traits in the mix.
+Here we find little evidence for pollen being a reliable proxy for plant trait composition.
+Exploring the possible causes for the lack of a relationship between pollen CWM values and vegetation CWM values, we suggest that this either has to do with the representation of the vegetation by the pollen data (abundance, species richness) or the large trait variation within pollen types.
 
@@
 the traits of species that do not show up in the dat.a
-
-## Conclusion
-Characterizing the factors that influence the relationship between vegetation trait composition and , and validating the relationship between trait composition of the vegetation and is an important step towards the application of traits in palaeoecology.
 
@@
     -   Is this a pollen problem or a trait problem?
+    - ambiguous relationship between vegetation and pollen trait 
+    - suggested positive, but credibility interval intersects 0 _> interpretation with caution
+    - largerr sample size needed
 

### 2023-01-24 03_CWM_estimates_pollen.R

@@ -47,33 +47,34 @@ selectedabun <- str_subset(colnames(dfPOL), "percent")[1]
 normaltraits <- c("PlantHeight","SLA","LA")
 
 # set number of cores to use
-if (Sys.info()[4] == "LUIN67311") {
-  plan(multisession, workers = 1)
-} else {
-  plan(multisession, workers = 6)
-}
+plan(multisession, workers = 3)
 
 # No subset ----
 if (1) {
 
-  furrr::future_map(normaltraits,
-                     ~cwm_pol(selectedtrait = .x,
-                              selectedcountry = "Scotland",
-                              selectedabun = "percent"
-                              # taxtrait = TRUE
-                              ),
-                    .options = furrr_options(seed = TRUE))
-  furrr::future_map(normaltraits,
-                     ~cwm_pol(selectedtrait = .x,
-                              selectedcountry = "Switzerland",
-                              selectedabun = "percent",
-                              taxtrait = TRUE
-                              ),
-                    .options = furrr_options(seed = TRUE))
+furrr::future_map(normaltraits,
+                   ~cwm_pol(selectedtrait = .x,
+                            selectedcountry = "Scotland",
+                            selectedabun = "percent",
+                            taxtrait = TRUE
+                            ),
+                  .options = furrr_options(seed = TRUE))
+furrr::future_map(normaltraits,
+                   ~cwm_pol(selectedtrait = .x,
+                            selectedcountry = "Switzerland",
+                            selectedabun = "percent",
+                            taxtrait = TRUE
+                            ),
+                  .options = furrr_options(seed = TRUE))
   furrr::future_map(normaltraits,
                     ~cwm_pol(selectedtrait = .x,
                                 selectedcountry = "Scotland",
                                 selectedabun = "adjustedpercent_mean"),
+                    .options = furrr_options(seed = TRUE))
+  furrr::future_map(normaltraits,
+                    ~cwm_pol(selectedtrait = .x,
+                             selectedcountry = "Scotland",
+                             selectedabun = "adjusted_helinger"),
                     .options = furrr_options(seed = TRUE))
   furrr::future_map(normaltraits,
                     ~cwm_pol(selectedtrait = .x,
@@ -90,32 +91,29 @@ if (1) {
 trsh <- c("tree", "shrub")
 herb <- c("grass", "herb")
 
-furrr::future_map(normaltraits, 
-                  ~cwm_pol_pft(selectedtrait = .x, 
+furrr::future_map(normaltraits,
[…]

### 2023-01-24 04_Correlation_bayesian.R

@@ -29,13 +29,12 @@ if (!require(runjags)) install.packages("runjags")
 if (!require(furrr)) install.packages("furrr")
 
 # load PFT summary table
-# filter sites with to little vegetation cover of particular type
-sum_pft <- readRDS("RDS_files/01_Percentage_cover_pft.rds")
-dropsites <- sum_pft %>% 
-  transmute(drop_herb = case_when(`Non-Woody`<= 5 ~ paste(zone, sitename)),
-            drop_trsh = case_when(Woody <= 5 ~ paste(zone, sitename)),
-            drop_wind = case_when(wind <= 5 ~ paste(zone, sitename)),
-            drop_nowind = case_when(`not wind` <= 5 ~ paste(zone, sitename)))
+# to filter sites with to little vegetation cover of particular type
+sum_pft <- readRDS("RDS_files/01_Percentage_cover_pft.rds") %>% 
+  replace_na(list(`Non-Woody` = 0,
+                  Woody = 0,
+                  `not wind` = 0,
+                  wind = 0)) 
 
 # collate data
 traits <- c("LA","SLA","PlantHeight")
@@ -44,7 +43,7 @@ pollen_files <- list.files("RDS_files/") %>%
   str_subset(pattern = "03_zCWM_estimates_") %>% 
   str_subset(paste(traits, collapse = "|")) %>% 
   str_subset(pattern = "pollen") %>% 
-  str_subset(pattern = "notinpollen", negate = TRUE)
+  str_subset(pattern = "notinpollen", negate = TRUE) 
 
 pollen_lab <- pollen_files %>% 
   str_subset(paste(traits, collapse = "|")) %>% 
@@ -92,16 +91,19 @@ dfCWM_veg <- veg_files %>%
   bind_rows() %>% 
   as_tibble() %>% 
   mutate(country = str_extract(label, pattern = "Scotland|Switzerland"),
-         pollination = dplyr::recode(pollination, `no wind` = "not wind")) %>% 
-  dplyr::select(country, sitename, trait, veg_mean = Mean, veg_sd = SD,
-                growthform, pollination, taxres, zone) %>% 
+         pollination = dplyr::recode(pollination, `no wind` = "not wind")) %>%
+  # only Scottish data
+  filter(country == "Scotland") %>% 
   # remove sites with to little cover of pft (5% or less) 
[…]

### 2023-01-24 05_Scatterplots.R

@@ -22,9 +22,87 @@
 if (!require(tidyverse)) install.packages("tidyverse")
 
 ## Load data
-dfCWM_veg <- readRDS("RDS_files/04_CWM_estimates_vegetation_Scotland.rds") 
-dfCWM_pol <- readRDS("RDS_files/04_CWM_estimates_pollen_Scotland.rds") %>%
-  filter(!str_detect(correction, pattern = "draw")) 
+# load PFT summary table
+# to filter sites with to little vegetation cover of particular type
+sum_pft <- readRDS("RDS_files/01_Percentage_cover_pft.rds") %>% 
+  replace_na(list(`Non-Woody` = 0,
+                  Woody = 0,
+                  `not wind` = 0,
+                  wind = 0)) 
+
+# collate data
+traits <- c("LA","SLA","PlantHeight")
+
+pollen_files <- list.files("RDS_files/") %>% 
+  str_subset(pattern = "03_CWM_estimates_") %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
+  str_subset(pattern = "pollen") %>% 
+  str_subset(pattern = "notinpollen", negate = TRUE) %>% 
+  str_subset(pattern = "draw", negate = TRUE)
+
+pollen_lab <- pollen_files %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
+  str_remove(., "03_zCWM_estimates_") %>% 
+  str_remove(., ".rds")
+
+veg_files <- list.files("RDS_files/") %>% 
+  str_subset(pattern = "03_CWM_estimates_") %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
+  str_subset(pattern = "pollen", negate = TRUE) 
+
+veg_lab <- veg_files %>% 
+  str_subset(paste(traits, collapse = "|")) %>% 
+  str_remove(., "03_zCWM_estimates_") %>% 
+  str_remove(., ".rds")
+
+folderpath.fun <- function(x)
+{paste("RDS_files/", x, sep = "/")}
+
+dfCWM_pol <- pollen_files %>% 
+  folderpath.fun(.) %>% 
+  purrr::map2(pollen_lab, ~readRDS(.) %>% 
+                mutate(label = .y)) %>% 
+  bind_rows() %>% 
+  as_tibble() %>% 
+  mutate(correction = case_when(str_detect(label, pattern = "adjustedpercent_draw") ~ str_extract(label, pattern = "draw[0-9]*$"),
[…]

### 2023-01-24 SI_Pollen_plot.R
initial commit

@@ -1,0 +1,92 @@
+## ---------------------------
+##
+## Script name: SI_pollen_plot
+##
+## Purpose of script: Make pollen plot
+##
+## Author: Annegreet Veeken
+##
+## Date Created: 2022-12-19
+##
+## Email: veeken.g.a@gmail.com
+##
+## ---------------------------
+##
+## Notes:
+##   
+## References:
+##
+## ---------------------------
+
+set.seed(1)
+options(scipen = 9999)
+
+## Load packages
+if (!require(tidyverse)) install.packages("tidyverse")
+
+## Load data
+dfPOL_Scot <- readRDS("RDS_files/01_Pollen_data_Scot.rds")
+dfPOL_Swiss <- readRDS("RDS_files/01_Pollen_data_Swiss.rds") %>% 
+  filter(!sitename == "unknown")
+bdm_meta <- read_csv("Data/bdm_metadata.csv") %>% 
+  mutate(sitename = as.character(aID_STAO))
+dfPOL_Swiss <- left_join(dfPOL_Swiss, bdm_meta, by = "sitename")
+
+# Plots
+(p_scot <- dfPOL_Scot %>% 
+    # tidy to allow grouped bar plot
+    dplyr::select(sitename, pollentaxon, percent, adjustedpercent_mean) %>% 
+    pivot_longer(cols = percent:adjustedpercent_mean, names_to = "pollen_correction", 
+                 values_to = "percent") %>% 
+    mutate(percent = percent*100) %>% 
+    # only common pollen types
+    group_by(pollentaxon) %>% 
+    filter(any(percent > 5)) %>% 
+    # plot
+    ggplot(aes(x = sitename, y = percent, fill = pollen_correction)) +
+      geom_bar(position = "dodge", stat = "identity") +
+      coord_flip() +
+      # scales
+      scale_x_discrete("Site", labels = 16:1, limits = rev) +
+      scale_y_continuous("Pollen percentage (%)", limits = c(0,100)) +
+      scale_fill_manual(name = "", 
+                       values = c("darkorchid", "darkorange"),
+                       labels = c(`percent` = "No correction", adjustedpercent_mean = "Correction")) +
+      # faceting
+      facet_wrap(~pollentaxon, nrow = 1) +
+      # theme
+      theme_bw() +
+      theme(text = element_text(size = 8),
+            strip.background = element_rect(fill = "white"),
+            legend.position = "top",
+            legend.key.size = unit(0.1, "inch")))
[…]

### 2023-01-24 cwm_functions.R
initial commit

@@ -1,0 +1,960 @@
+# CWM vegetation Scotland ----
+cwm_veg_scot <- function(selectedabun, selectedtrait, selectedtaxres, selectedpolmode, 
+                         selectedpft, notinpollen = FALSE, taxtrait = FALSE){
+  ## prepare abundance data
+  abun <- lABUN %>% 
+    # select zone 
+    pluck(selectedabun) %>% 
+    # optional filter for polmode and pft
+    filter(polmode %in% selectedpolmode) %>% 
+    filter(growthform %in% selectedpft) %>% 
+    # optional to filter out taxa that are not in pollen data
+    {if (notinpollen) filter(., !pollentaxon == "Not in pollen data") else . } %>% 
+    # rename column to selected taxonomic resolution
+    rename(tax = all_of(selectedtaxres)) %>% 
+    filter(!tax == "Unindentified") %>% 
+    dplyr::select(sitename, tax, abun) %>% 
+    drop_na() 
+  
+  ## prepare trait data
+  # taxa in the abundance data
+  taxa <- abun %>% 
+    pull(tax) %>% 
+    unique() %>% 
+    sort
+  
+  # missing field collected trait data
+  traittax <- lTrait %>%
+    pluck(selectedtrait) %>% 
+    pull(selectedtaxres) %>% 
+    unique
+  missingtax <- taxa[!taxa %in% traittax] 
+  
+  # filter species with too little observations (10 or less)
+  nobs <- lTrait %>%
+    pluck(selectedtrait) %>%
+    arrange(selectedtaxres) %>% 
+    group_by(across(all_of(selectedtaxres))) %>%
+    summarise(n = n()) %>%
+    filter(n <= 10) %>%
+    pull(selectedtaxres)
+  
+  # find taxa with not or too little data in TRY gapfilled data
+  TRY <- dfTRY %>% 
+    dplyr::select(tax = all_of(selectedtaxres), 
+                  all_of(selectedtrait)) %>% 
+    filter(tax %in% c(missingtax, nobs)) 
+  
+  # Create trait data df with field an try data
+  trait <- lTrait %>%
+    pluck(selectedtrait) %>% 
+    dplyr::select(tax = all_of(selectedtaxres), 
+                  all_of(selectedtrait)) %>% 
+    # join with try data for the missing traits
+    bind_rows(TRY) %>% 
+    filter(tax %in% taxa) %>%
+    # sort alphabetically
+    arrange(tax) %>%
[…]
